### 2017-08-07 examples/Example_001_ER_CTCF.Rmd
Created simiplified examples of the analyses in the ./examples subdirectory.

---
title: "Normalization Example using internal CTCF peaks as a standard"
author: "Andrew Holding"
date: "8/7/2017"
output: pdf_document
---

## Introduction

This is a normalisation using linear regression to CTCF of MCF7 cells chip treameant etc.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load convience functions

These functions facilitate the normalisation of data.

```{r brundle}
source('../package/brundle.R')
```

## Apply settings

```{r}
jg.controlMinOverlap      <- 5
jg.controlSampleSheet     <- "samplesheet/samplesheet_SLX14438_hs_CTCF_DBA.csv"
jg.experimentSampleSheet  <- "samplesheet/samplesheet_SLX14438_hs_ER_DBA.csv"
jg.treatedCondition       =  "Fulvestrant"
jg.untreatedCondition     =  "none"

```

## Load control and experimental DiffBind object

To keep file size down these are provided as a Rdata File rather than as raw counts.

```{r}
filename<-"Rdata/example_001_SLX-14438_dba_human_ER_CTCF.rda"
if(!file.exists(filename)){
  dbaExperiment <- jg.getDba(jg.experimentSampleSheet, bRemoveDuplicates=TRUE)
  dbaControl    <- jg.getDba(jg.controlSampleSheet, bRemoveDuplicates=TRUE)
  save(dbaExperiment,dbaControl,file=filename)
} else {
  load(filename)
}

```

```{r}
#Load Sample Ids from control sample sheet.
jg.sampleIds <- jg.getSampleIds(jg.controlSampleSheet)
```

```{r}

## Extract Peak set from DiffBind
jg.experimentPeakset <- jg.dbaGetPeakset(dbaExperiment)
jg.controlPeakset    <- jg.dbaGetPeakset(dbaControl)

```

```{r}
#Get counts for each condition
jg.controlCountsTreated<-jg.getControlCounts(jg.controlPeakset, 
                                             jg.controlSampleSheet,
                                             jg.treatedCondition)
jg.controlCountsUntreated<-jg.getControlCounts(jg.controlPeakset,
                                               jg.controlSampleSheet,
                                               jg.untreatedCondition)

```

```{r}
#Get sample names for conditions
jg.untreatedNames <- names(jg.controlCountsUntreated)
jg.treatedNames   <- names(jg.controlCountsTreated)

```

```{r}

##Plot showing normalization calculation (Optional)
  jg.plotNormalization(jg.controlCountsTreated,
                     jg.controlCountsUntreated)
```

```{r}
##Get Normalization Coefficient
jg.coefficient<-jg.getNormalizationCoefficient(jg.controlCountsTreated,
                                               jg.controlCountsUntreated)
jg.correctionFactor<-jg.getCorrectionFactor(jg.experimentSampleSheet,
                                            jg.treatedNames,
                                            jg.untreatedNames)

```

```{r}
#Apply coefficent and control factor
jg.experimentPeaksetNormalised<-jg.applyNormalisation(jg.experimentPeakset,
                                                      jg.coefficient,
                                                      jg.correctionFactor,
                                                      jg.treatedNames)
```

```{r}
#Return values to Diffbind and plot normalised result.
jg.dba <- DiffBind:::pv.resetCounts(dbaExperiment,
                                    jg.experimentPeaksetNormalised)

```

```{r}
jg.dba_analysis<-dba.analyze(jg.dba)
dba.plotMA(jg.dba_analysis,bFlip=TRUE)

```

## Save results

```{r}
write.csv(dba.report(jg.dba_analysis),file="results/Example_001.csv")
```

### 2017-08-07 examples/Example_002_ER_CTCF_DeSeq.Rmd
Created simiplified examples of the analyses in the ./examples subdirectory.

---
title: "Normalization Example using internal CTCF peaks as a standard (no-linear fit)"
author: "Andrew Holding"
date: "8/7/2017"
output: pdf_document
---

## Introduction

This is a normalisation using only DeSeq Size Factors to CTCF of MCF7 cells chip treameant etc.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load convience functions

These functions facilitate the normalisation of data.

```{r brundle, echo=FALSE}
source('../package/brundle.R')
```

## Apply settings

```{r}
jg.controlMinOverlap      <- 5
jg.controlSampleSheet     <- "samplesheet/samplesheet_SLX14438_hs_CTCF_DBA.csv"
jg.experimentSampleSheet  <- "samplesheet/samplesheet_SLX14438_hs_ER_DBA.csv"

```

## Load control and experimental DiffBind object

To keep file size down these are provided as a Rdata File rather than as raw counts.

```{r}
filename<-"Rdata/example_001_SLX-14438_dba_human_ER_CTCF.rda"
if(!file.exists(filename)){
  dbaExperiment <- jg.getDba(jg.experimentSampleSheet, bRemoveDuplicates=TRUE)
  dbaControl    <- jg.getDba(jg.controlSampleSheet, bRemoveDuplicates=TRUE)
  save(dbaExperiment,dbaControl,file=filename)
} else {
  load(filename)
}

```

Diffbind was simply used as a convient way to extract peak counts. The actual DESeq Pipeline starts simply needs a matrix so we covert from Diffbind and save as csv to provide a starting point for a DESeq pipeline.

```{r}

filename<-"csv/experimentalPeakset.csv"
if(!file.exists(filename)){
  ## Extract Peak set from DiffBind
  jg.experimentPeakset <- jg.dbaGetPeakset(dbaExperiment)
  #Convert Peakset to DeSeq Workflow
  jg.experimentPeaksetDeSeq<-jg.convertPeakset(jg.experimentPeakset)
  #Save to CSV

  write.csv(jg.experimentPeaksetDeSeq,file=filename)
}

#Repeat for control samples.

filename<-"csv/controlPeakset.csv"
if(!file.exists(filename)){
jg.controlPeakset    <- jg.dbaGetPeakset(dbaControl)
jg.controlPeaksetDeSeq<-jg.convertPeakset(jg.controlPeakset)
write.csv(jg.controlPeaksetDeSeq,   file=filename) 
}

```

# Pipe line starts here
## Load CSV dataset in DESeq format

```{r}
jg.controlPeaksetDeSeq <- read.csv( file="csv/controlPeakset.csv"
                                      ,check.names=FALSE, row.names = 1)
jg.experimentPeaksetDeSeq<- read.csv(file="csv/experimentalPeakset.csv"
                                     ,check.names=FALSE, row.names = 1)
```

Note we use the samplesheet here, not needed but neater.

```{r}

#Establish size factors directly from Control data
jg.controlSizeFactors = estimateSizeFactorsForMatrix(jg.controlPeaksetDeSeq)

#Get conditions dataframe for DeSeq
jg.conditions <- read.csv(file=jg.controlSampleSheet, header=TRUE, sep=",")['Condition']

#Run DeSeq on control
jg.controlDeSeq<-jg.runDeSeq(jg.controlPeaksetDeSeq, jg.conditions,jg.controlSizeFactors)
jg.controlResultsDeseq   = results(jg.controlDeSeq)

#Run DeSeq on experiment
jg.experimentDeSeq<-jg.runDeSeq(jg.experimentPeaksetDeSeq, jg.conditions,jg.controlSizeFactors)
jg.experimentResultsDeseq   = results(jg.experimentDeSeq)

```

## Export results as CSV

These are not annotated, but other packages exist to do that.
```{r}

write.csv(file="results/Example_002.csv", jg.experimentResultsDeseq)

```

```{r}

#Plot results

  jg.plotDeSeq(jg.controlResultsDeseq,
               p=0.01, 
               title.main="Fold-change in CTCF binding",
               flip=T
               )

jg.plotDeSeq(jg.experimentResultsDeseq,
             p=0.01,
             title.main="Fold-change in ER binding",
             flip=T
             )

#Draw Combined figure.
  jg.plotDeSeqCombined(jg.controlResultsDeseq,
                       jg.experimentResultsDeseq,
                       title.main="ER and CTCF Binding Folding changes on ER treatment",
                       p=0.01,flip=TRUE)

```

### 2017-08-07 examples/Example_003_ER_H2av.Rmd
Created simiplified examples of the analyses in the ./examples subdirectory.

--- 2017-08-07 examples/Example_001_ER_CTCF.Rmd
@@
 ---
-title: "Normalization Example using internal CTCF peaks as a standard"
+title: "Normalization Example using spike-in control as a standard"
 author: "Andrew Holding"
@@
 
-```{r brundle}
+```{r brundle, results='hide', message=FALSE, warning=FALSE}
 source('../package/brundle.R')
@@
 ```{r}
+dbaSummits                <- 200
 jg.controlMinOverlap      <- 5
-jg.controlSampleSheet     <- "samplesheet/samplesheet_SLX14438_hs_CTCF_DBA.csv"
-jg.experimentSampleSheet  <- "samplesheet/samplesheet_SLX14438_hs_ER_DBA.csv"
+jg.controlSampleSheet     <- "samplesheet/samplesheet_SLX8047_dm.csv"
+jg.experimentSampleSheet  <- "samplesheet/samplesheet_SLX8047_hs.csv"
 jg.treatedCondition       =  "Fulvestrant"
@@
 ```{r}
-filename<-"Rdata/example_001_SLX-14438_dba_human_ER_CTCF.rda"
+filename<-"Rdata/example_003_SLX-8047_dba_HsDm.rda"
 if(!file.exists(filename)){
@@
 jg.dba_analysis<-dba.analyze(jg.dba)
-dba.plotMA(jg.dba_analysis,bFlip=TRUE)
+dba.plotMA(jg.dba_analysis)
 
@@
 ```{r}
-write.csv(dba.report(jg.dba_analysis),file="results/Example_001.csv")
+write.csv(dba.report(jg.dba_analysis),file="results/Example_003.csv")
 ```

### 2017-08-07 examples/Example_001_ER_CTCF.Rmd
Writing descriptive text for examples

@@
 ---
-title: "Normalization Example using internal CTCF peaks as a standard"
+title: "Normalization example using internal CTCF peaks as a standard"
 author: "Andrew Holding"
@@
 
-This is a normalisation using linear regression to CTCF of MCF7 cells chip treameant etc.
+This is a worked example of our method to normalize ER ChIP-seq data using CTCF binding to provide an internal control.
+
+## Preprocessing.
+
+To prepare the data for the differential binding analysis the data had to be aligned and CTCF and ER peaks called.
+As our CTCF control doesn't require any special aligment, data was aligned as for a standard ChIP-seq experiment.
+Peak calling was done with macs2 using the script shown below.
+
+```{bash, eval=FALSE}
+
+#Calls peaks for all files in a directory against a single control. Change
+# <filename> to the filename of you input sample.
+
+ control=<filename>
+
+ for bam in *bam
+ do
+   root=`basename $bam .bam`
+   macs2 callpeak -t $bam -c $control -f BAM -n $root -g hs
+ done
+
+```
+
+As macs2 cannot differentiate between a peak arising from CTCF binding or ER binding we used two samples were squence to establish a CTCF binding consensus.
+One treated and one untreated.
+These peaks were then merged as follows.
+
+```{bash, eval=FALSE}
+
+cat  SLX-14438.D701_D504.HKN27BBXX.s_8.r_1.fq.gz.bam_peaks.xls \
+      SLX-14438.D702_D503.HKN27BBXX.s_8.r_1.fq.gz.bam_peaks.xls > CTCF_merged.bed
+
+grep -v \# CTCF_merged.bed  | grep -v start > CTCF_merged_filtered.bed
+sort -k1,1 -k2,2n CTCF_merged_filtered.bed > CTCF_merged_sorted_filtered.bed
+
+bedtools merge -i CTCF_merged_sorted_filtered.bed  > CTCF_union.bed
+
+```
+
+We then used this CTCF consensus to establish specfic CTCF or ER binding in the experimental samples.
+Either taking the interesection of the CTCF consensus and the peak file for a particluar sample to provide a list of CTCF binding sites in the sample, or subratracting the consensus CTCF peaks from the macs2 output to provide a list ER binding sites for that sample.
+
+```{bash, eval=FALSE}
+
+mkdir CTCF
+for f in *.narrowPeak
+do
+ 	echo $f
+ 	bedtools intersect -a $f -b CTCF_union.bed  > CTCF/$f 
+done
+
+mkdir ER
+for f in *.narrowPeak
+do
+ 	echo $f
+ 	bedtools subtract -b CTCF_union.bed -a $f -A > ER/$f 
+done
+
+```
+
+These were then used to write two DiffBind samplesheets as described in the DiffBind vignette.
+One using the CTCF peak list generated from the above scripts, this is the control sample sheet.
+The second uses the ER peaks, this is the experimetnal sample sheet.
+These two sample sheets are then provided as the input in the analysis below.
 
@@
 
-These functions facilitate the normalisation of data.
+Brundle provies a series of functions we developed for this pipeline that are used to normalizing ChIP-seq data. The usage of these functions are explained in the example that follows.
 
-```{r brundle}
+```{r brundle,results='hide', message=FALSE, warning=FALSE}
 source('../package/brundle.R')
@@
 ## Apply settings
+
+The intial settings for the analysis are minimal, jg.controlMinOverlap is set to 5 as this is 1 less than the total sample count and we want a strong consensus of CTCF peaks to normalise against.
+The control and experiment sample sheets are as described above.
+The conditions reflect the treated and untreaded condition we have used when writing our sample sheet.
 
@@
 
-To keep file size down these are provided as a Rdata File rather than as raw counts.
+Typically we would load the raw data using Diffbind as we are plan to return our normalized data as a Diffbind object.
+
+In this example, as the BAM files which are greater in size than the github filesize limit, we have included the data as the Diffbind object these function would output and instead load that directly.
+If the code doesn't find the R data object it will try to regenerate if from the raw data.
 
@@
 
-```{r}
-#Load Sample Ids from control sample sheet.
-jg.sampleIds <- jg.getSampleIds(jg.controlSampleSheet)
-```
+Diffbind allows us to extract the counts in each peak, we use this function to obtain the data in a form that is easier to normalize.
 
@@
 
-## Extract Peak set from DiffBind
+#  Extract Peak set from DiffBind
+
 jg.experimentPeakset <- jg.dbaGetPeakset(dbaExperiment)
@@
 ```
+
+To normalize we need to seperate samples out by condition, this is done using the following function.
+The peakset contains the reads in peaks, the samplesheet provides the metadata needed and the jg.treatedCondition/jg.untreatedCondition specify which condition we want to get the counts for.
 
@@
 
+At the same time we can extract the samples names for each condition to use later.
+
 ```{r}
@@
 
-```{r}
+The normalization factor is generated by a linear regression generated from the mean number of counts for each peak before and after treatment.
+The gradient of the linear fit should be equal to 1 given the assumtion that before and after treatment the majority of the control peaks do not change.
 
-##Plot showing normalization calculation (Optional)
+This plot is not neaded to caluclate the normalization factor but it does provide a visual interpretation of the process.
+
+```{r, message=FALSE}
+
+# Plot showing normalization calculation (Optional)
+
 jg.plotNormalization(jg.controlCountsTreated,
@@
 
+There are two parts to the normalization coefficent we use in the final correction.
+The first is the the coefficent to make the levels of CTCF binding before and after treament equal.
+The second is because DiffBind normalises against the total read count for each sample, we therefore generated a correction factor to control for this.
+
 ```{r}
-##Get Normalization Coefficient
+# Get Normalization Coefficient
+
 jg.coefficient<-jg.getNormalizationCoefficient(jg.controlCountsTreated,
@@
 ```
+
+Once we have these two peices of information we apply both the normalisation coeffecent and the correction factor to our data.
 
@@
 
+Finally we return the correct values into the DiffBind object.
+This requires a modificed version Diffbind as included in the original manuscript.
+It is expected that this will be supported natively by furture versions of Diffind.
+
 ```{r}
@@
 
-```{r}
+We can then visualise the data as normal using DiffBind.
+
+```{r, message=FALSE}
 jg.dba_analysis<-dba.analyze(jg.dba)

### 2017-08-07 examples/Example_002_ER_CTCF_DeSeq.Rmd
Writing descriptive text for examples

@@
 
-This is a normalisation using only DeSeq Size Factors to CTCF of MCF7 cells chip treameant etc.
+This is a worked example of our method alternative to normalize ER ChIP-seq data using CTCF binding to provide an internal control.
+Unlike our previous example this normalisation is using only DESeq Size Factors.
+This method was found to be less effective but was used for comparision in the supporting publication.
+This markdown builds on Example 001 and therefore that should be read first.
+Example 001 also explains the pre-processing steps.
 
@@
 
-```{r brundle, echo=FALSE}
+```{r brundle,results='hide', message=FALSE, warning=FALSE}
 source('../package/brundle.R')
@@
 
-## Load control and experimental DiffBind object
+## Load control and experimental counts from Example 001
 
-To keep file size down these are provided as a Rdata File rather than as raw counts.
+In this example Diffbind was simply used as a convient way to extract peak counts and provided consitancy for comparision between methods.
+
+To keep file size down counts are provided as a Rdata file rather than as raw data.
 
@@
 
-Diffbind was simply used as a convient way to extract peak counts.
-The actual DESeq Pipeline starts simply needs a matrix so we covert from Diffbind and save as csv to provide a starting point for a DESeq pipeline.
+The actual DESeq Pipeline starts simply needs a matrix so we covert from Diffbind and save as csv to provide a starting point.
 
-```{r}
+```{r, message=FALSE}
 
@@
 
-# Pipe line starts here
+# DESeq Pipeline starting from CSV
+
 ## Load CSV dataset in DESeq format
+
+Here we use the CSV files the example code above generated; however, it is possible to generate the input data through a wide variety of methods and we have provided only one.
 
@@
 
-Note we use the samplesheet here, not needed but neater.
+To normalize the data we use the DESeq2 function to estimate the size factors.
+This is a very simple method, and it provides some degree of normalization.
+We also use the original sample sheet to get the different conditions.
 
-```{r}
+```{r,message=FALSE}
 
@@
 
-These are not annotated, but other packages exist to do that.
+We can export the DESeq2 results in CSV format.
+
 ```{r}
@@
 
+## Plots
+
+The Brundle package also contains some convience functions for plotting the output of DESeq2 inplace of DiffBind. These generate the following figures.
+
 ```{r}
-
-#Plot results
 

### 2017-08-07 examples/Example_003_ER_H2av.Rmd
Writing descriptive text for examples

@@
 
-This is a normalisation using linear regression to CTCF of MCF7 cells chip treameant etc.
+This is a normalisation using H2av binding within an external spike-in of droshilia chromatin to provide a control for the analysis of ER binding in MCF7 cells after treatment with Fulvestrant.
 
@@
 
-To keep file size down these are provided as a Rdata File rather than as raw counts.
+As before to keep file size down data are provided as a RdatafFile rather than as raw BAM files.
 
@@
 
-## Extract Peak set from DiffBind
+# Extract Peak set from DiffBind
 jg.experimentPeakset <- jg.dbaGetPeakset(dbaExperiment)

### 2017-08-07 examples/Example_003_ER_H2av.Rmd
Examples finished subject to proof-reading.

@@
 
+This example builds on the previous two and therefore these should be read first as many details of key functions have not been repeated.
+
 ```{r setup, include=FALSE}
@@
 ```
+
+## Preprocessing
+
+Unlike Example 001 & 002 the use of a xenogenic spike in requires more pre-processing.
+The scripts for this process are provided below.
+Genomes were downloaded from http://support.illumina.com/sequencing/sequencing_software/igenome.html.
+
+The first step of our preprossing pipeline requires the generation of a combined Human/Drosophila genome.
+We used the following script to generated this.
+The human and drosophila genomes downloaded from the link provided above are contained in a subdirectory "genomes" relative to where the script was run.
+These example scripts are not evalutated in this markdown due to the time  it takes to process the data.
+
+```{bash, eval=FALSE}
+
+# Merge the genomes
+
+human=../Homo_sapiens/UCSC/hg19/Sequence/WholeGenomeFasta/genome.fa
+droso=../Drosophila_melanogaster/UCSC/dm3/Sequence/WholeGenomeFasta/genome.fa
+
+mkdir genomes/dmhs
+cd ./genomes/dmhs
+sed "s/^>/>hs_/" $human > tmp
+sed "s/^>/>dm_/" $droso > tmp2
+cat tmp tmp2 > dmhs.fa
+rm tmp tmp2
+
+# Merge Annotations
+human=../Homo_sapiens/UCSC/hg19/Annotation/Archives/archive-current/Genes/genes.gtf
+droso=../Drosophila_melanogaster/UCSC/dm3/Annotation/Archives/archive-current/Genes/genes.gtf
+
+sed "s/^/hs_/" $human > tmp
+sed "s/^/dm_/" $droso > tmp2
+
+cat tmp tmp2 > dmhs.gtf
+rm tmp tmp2
+
+```
+
+To allow use to use the merged genome we first indexed it using bowtie2. For our system we specified 32 threads, this will depend on the number of processors you have available.
+
+```{bash, eval=FALSE}
+### Index the merged genome
+cd ./genomes/dmhs
+
+bowtie2-build --threads 32 dmhs.fa dmhs
+
+```
+
+Once indexed, we could the aligned to the merge genome. This script assumes that the FastQ files are in a directory ./SLX-8047_dhms from where the scripts are run.
+
+```{bash, eval=FALSE}
+#Align, Sort and Build indeces
+
+mkdir ./SLX-8047_dmhs
+cd ./SLX-8047_dmhs
+
+genome=../genomes/dmhs/dmhs
+for fq in ../SLX-8047/*fq.gz
+do
+  root=`basename $fq`
+  bowtie2 -p 32 -x $genome -U $fq > tmp.sam \
+  && samtools view -Sbh tmp.sam > tmp.bam \
+  && samtools sort tmp.bam ${root}  \
+  && samtools index ${root}.bam
+done
+rm tmp.sam tmp.bam
+```
+
+Blacklisting requires a merged blacklist. This script downloads the blacklist and merges it, then applies this to the alignment.
+
+```{bash, eval=FALSE}
+
+mkdir blacklists
+cd blacklists
+# Human hg19/GRCh37
+wget http://hgdownload.cse.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeMapability/wgEncodeDacMapabilityConsensusExcludable.bed.gz --no-check-certificate
+
+# Drosophila dm3
+wget http://www.broadinstitute.org/~anshul/projects/fly/blacklist/dm3-blacklist.bed.gz --no-check-certificate  
+gunzip dm3-blacklist.bed.gz
+
+### Generate a merged Drosophila/Human blacklist
+droso=./dm3-blacklist.bed
+homo=./hg19-blacklist.bed
+
+sed "s/^/hs_/" $homo |cut -f1-3 > tmp
+sed "s/^/dm_/" $droso > tmp2
+cat tmp tmp2 > dmhs-blacklist.bed
+rm tmp tmp2
+
+cd ..
+
+bl=../blacklists/dmhs-blacklist.bed
+
+cd ./SLX-8047_dmhs
+
+mkdir ./blacklist_filtered
+for f in *.bam
+do
+	echo $f
+	bedtools intersect -v -abam $f -b $bl  > blacklist_filtered/$f
+done
+
+### Re-index
+
+cd ./blacklist_filtered
+for f in *.bam
+do
+  samtools index $f
+done
+
+```
+
+Peak calling of the reads aligned to the merged genome is then similar to that of a typical experiment.
+
+```{bash, eval=FALSE}
+
+### Run macs on the blacklisted data
+mkdir ./SLX-8047_dmhs/peaks
+cd ./SLX-8047_dmhs/peaks
+control=../blacklist_filtered/SLX-8047.D705_D507.C81G5ANXX.s_1.r_1.fq.gz.bam
+
+for bam in ../blacklist_filtered/*bam
+do
+  root=`basename $bam .bam`
+  macs2 callpeak -t $bam -c $control -f BAM -n $root -g hs
+done
+
+```
+
+After aligning the data can calling peaks we can then split the data back into Drosphila and Human.
+
+```{bash, eval=FALSE}
+# The BAMs and BEDs should be split into Drosophila (for normalization) and into Human (for experimental)
+
+### BAMs, split in two
+cd ./SLX-8047_dmhs/blacklist_filtered
+
+mkdir human
+mkdir drosophila
+
+for bam in *bam
+do
+  samtools view -h $bam | grep -v "dm_chr" | sed s/hs_chr/chr/g | samtools view -bS - > human/$bam
+  samtools view -h $bam | grep -v "hs_chr" | sed s/dm_chr/chr/g | samtools view -bS - > drosophila/$bam 
+done
+
+# Index
+cd ./human
+echo `pwd`
+
+for bam in *bam
+do
+  samtools index $bam
+done
+
+cd ../drosophila
+echo `pwd`
+
+for bam in *bam
+do
+  samtools index $bam
+done
+
+### Peaks BEDs and XLSs, split in two
+cd ../../peaks
+
+echo `pwd`
+
+mkdir human
+mkdir drosophila
+
+for bed in *peaks.bed
+do
+  grep -v "dm_chr" $bed | sed s/hs_chr/chr/g > human/$bed
+  grep -v "hs_chr" $bed | sed s/dm_chr/chr/g > drosophila/$bed
+done
+
+for xls in *peaks.xls
+do
+  grep -v "dm_chr" $xls | sed s/hs_chr/chr/g > human/$xls
+  grep -v "hs_chr" $xls | sed s/dm_chr/chr/g > drosophila/$xls
+done
+
+for narrow in *narrowPeak
+do
+  grep -v "dm_chr" $narrow | sed s/hs_chr/chr/g > human/$narrow
+  grep -v "hs_chr" $narrow | sed s/dm_chr/chr/g > drosophila/$narrow
+done
+
+```
+
+With the peak files and the seperate alignments extract we are can then write samplesheets.
+For the normalization we need one for the drosphila aligned reads from each sample and the respective H2av peak file, and one for human reads from each sample and their repsective ER peak files.
+This forms the bais of the following analysis.
 

### 2017-08-10 041_SLX-14229_SLX-8045_compare.R
Created p-value comparision

@@ -1,0 +1,95 @@
+
+
+######################
+#
+# Functions
+#
+######################
+
+setwd("/Volumes/FlyPeaks/FlyPeaks")
+source('package/brundle.R')
+
+######################
+#
+# Main Code
+#
+######################
+
+filename <- "Rdata/041_SLX-14229_dba_human_ER_CTCF_normalised.rda"
+if (!file.exists(filename)) {
+    dbaSummits                <- 200
+    jg.controlMinOverlap      <- 5
+    jg.controlSampleSheet     <-
+        "samplesheet/samplesheet_SLX14229_hs_CTCF_DBA.csv"
+    jg.experimentSampleSheet  <-
+        "samplesheet/samplesheet_SLX14229_hs_ER_DBA.csv"
+    jg.treatedCondition       =  "Fulvestrant"
+    jg.untreatedCondition     =  "none"
+
+    filename_dba <- "Rdata/041_SLX-14229_dba_human_ER_CTCF.rda"
+    if (!file.exists(filename)) {
+        dbaExperiment <- jg.getDba(jg.experimentSampleSheet, dbaSummits)
+        dbaControl    <- jg.getDba(jg.controlSampleSheet,   dbaSummits)
+        save(dbaExperiment, dbaControl, file = filename)
+    } else {
+        load(filename)
+    }
+
+
+    #Condensed code from previous script
+    jg.sampleIds <- jg.getSampleIds(jg.controlSampleSheet)
+    jg.experimentPeakset <- jg.dbaGetPeakset(dbaExperiment)
+    jg.controlPeakset    <- jg.dbaGetPeakset(dbaControl)
+    jg.controlCountsTreated <- jg.getControlCounts(jg.controlPeakset,
+                                                   jg.controlSampleSheet,
+                                                   jg.treatedCondition)
+    jg.controlCountsUntreated <- jg.getControlCounts(jg.controlPeakset,
+                                                     jg.controlSampleSheet,
+                                                     jg.untreatedCondition)
+    jg.untreatedNames <- names(jg.controlCountsUntreated)
+    jg.treatedNames   <- names(jg.controlCountsTreated)
+    jg.coefficient <-
+        jg.getNormalizationCoefficient(jg.controlCountsTreated,
+                                       jg.controlCountsUntreated)
+    jg.correctionFactor <-
[…]

### 2017-08-11 032_SLX-14229_SLX-8045_compare.R
Code to establish more accurate p-value

@@ -204,5 +204,5 @@ text(3.1,-0.25, expression( "gradient = 0.96, r = 0.503, p-value < 2.2 x" ~ 10^{
 dev.off()     
 
 
-cor.test(SLX8047_fc,SLX14229_fc, method="pearson")
+result<-cor.test(SLX8047_fc,SLX14229_fc, method="pearson")
 

### 2017-08-11 041_SLX-14229_SLX-8045_compare.R

@@ -93,3 +93,8 @@ plot(
     main = "Effect of Normalisation on p-value"
 )
 dev.off()
+
+
+result<-cor.test(normalised$p.value,unnormalised$p.value, method="pearson")
+result
+result$p.value

### 2017-08-12 examples/Example_001_ER_CTCF.Rmd
Examples finished subject to proof-reading.
Edited examples

@@
 To prepare the data for the differential binding analysis the data had to be aligned and CTCF and ER peaks called.
-As our CTCF control doesn't require any special aligment, data was aligned as for a standard ChIP-seq experiment.
+As our CTCF control doesn't require any special aligment, data was aligned as for a standard ChIP-seq experiment and followed by removal of reads from blacklisted regions.
 Peak calling was done with macs2 using the script shown below.
@@
 
-As macs2 cannot differentiate between a peak arising from CTCF binding or ER binding we used two samples were squence to establish a CTCF binding consensus.
+As macs2 cannot differentiate between a peak arising from CTCF binding or ER binding we used samples with only CTCF antibody to establish a CTCF binding consensus.
 One treated and one untreated.
@@
 
-These were then used to write two DiffBind samplesheets as described in the DiffBind vignette.
+The peak list files were then used as the basis for two DiffBind samplesheets as described in the DiffBind vignette.
 One using the CTCF peak list generated from the above scripts, this is the control sample sheet.
@@
 
-Brundle provies a series of functions we developed for this pipeline that are used to normalizing ChIP-seq data. The usage of these functions are explained in the example that follows.
+Brundle provies a series of functions we developed for this pipeline that are used to normalize ChIP-seq data. The usage of these functions are explained in the example that follows.
 
@@
 
-Typically we would load the raw data using Diffbind as we are plan to return our normalized data as a Diffbind object.
+Typically we would load the raw data using Diffbind as we are planning to return our normalized data as a Diffbind object.
 
-In this example, as the BAM files which are greater in size than the github filesize limit, we have included the data as the Diffbind object these function would output and instead load that directly.
+In this example, as the BAM files which are greater in size than the github filesize limit, we have included the data as the Diffbind object these function would output and instead of  loading that directly.
 If the code doesn't find the R data object it will try to regenerate if from the raw data.
@@
 Finally we return the correct values into the DiffBind object.
-This requires a modificed version Diffbind as included in the original manuscript.
+This requires a modified version Diffbind as included in the original manuscript.
 It is expected that this will be supported natively by furture versions of Diffind.

### 2017-08-12 examples/Example_003_ER_H2av.Rmd
Edited examples

@@
 
-This example builds on the previous two and therefore these should be read first as many details of key functions have not been repeated.
+This example builds on the previous two and therefore they should be read first as many details of key functions have not been repeated.
 
@@
 
-With the peak files and the seperate alignments extract we are can then write samplesheets.
+With the peak files and the seperate alignments extracted we are can then write samplesheets.
 For the normalization we need one for the drosphila aligned reads from each sample and the respective H2av peak file, and one for human reads from each sample and their repsective ER peak files.
@@
 
-As before to keep file size down data are provided as a RdatafFile rather than as raw BAM files.
+As before to keep file size down data are provided as in Rdata format rather than as raw BAM files.
 

### 2017-08-14 examples/Example_001_ER_CTCF.Rmd
Correcting typos.

--- 2017-08-07 examples/Example_001_ER_CTCF.Rmd
@@
 ---
-title: "Normalization Example using internal CTCF peaks as a standard"
+title: "Normalisation example using Internal CTCF Peaks as a Standard"
 author: "Andrew Holding"
-date: "8/7/2017"
+date: "7/8/2017"
 output: pdf_document
@@
 
-This is a normalisation using linear regression to CTCF of MCF7 cells chip treameant etc.
+This is a worked example of our method to normalize ER ChIP-seq data using CTCF binding to provide an internal control.
+
+## Preprocessing
+
+To prepare the data for the differential binding analysis, the data had to be aligned and CTCF and ER peaks called.
+As our CTCF control does not require any special alignment, data was aligned as for a standard ChIP-seq experiment and followed by removal of reads from blacklisted regions.
+Peak calling was done with macs2 using the script shown below.
+
+```{bash, eval=FALSE}
+
+#Calls peaks for all files in a directory against a single control. Change
+# <filename> to the filename of your input sample.
+
+ control=<filename>
+
+ for bam in *bam
+ do
+   root=`basename $bam .bam`
+   macs2 callpeak -t $bam -c $control -f BAM -n $root -g hs
+ done
+
+```
+
+As macs2 cannot differentiate between a peak arising from CTCF binding or ER binding, we used samples  prepared with only CTCF antibody to establish a CTCF binding consensus: one treated and one untreated.
+These peaks were then merged as follows.
+
+```{bash, eval=FALSE}
+
+cat  SLX-14438.D701_D504.HKN27BBXX.s_8.r_1.fq.gz.bam_peaks.xls \
+      SLX-14438.D702_D503.HKN27BBXX.s_8.r_1.fq.gz.bam_peaks.xls > CTCF_merged.bed
+
+grep -v \# CTCF_merged.bed  | grep -v start > CTCF_merged_filtered.bed
+sort -k1,1 -k2,2n CTCF_merged_filtered.bed > CTCF_merged_sorted_filtered.bed
+
+bedtools merge -i CTCF_merged_sorted_filtered.bed  > CTCF_union.bed
+
+```
+
+We then used this CTCF consensus to establish specific CTCF or ER binding in the experimental samples, either by taking the intersection of the CTCF consensus and the peak file for a particluar sample to provide a list of CTCF binding sites in the sample, or by subtratracting the consensus CTCF peaks from the macs2 output to provide a list of ER binding sites for that sample.
+
+```{bash, eval=FALSE}
+
+mkdir CTCF
+for f in *.narrowPeak
+do
+ 	echo $f
+ 	bedtools intersect -a $f -b CTCF_union.bed  > CTCF/$f 
+done
+
+mkdir ER
+for f in *.narrowPeak
+do
+ 	echo $f
+ 	bedtools subtract -b CTCF_union.bed -a $f -A > ER/$f 
+done
+
+```
+
+The peak list files were then used as the basis for two DiffBind sample sheets as described in the DiffBind vignette.
+One uses the CTCF peak list generated from the above scripts; this is the control sample sheet.
+The second uses the ER peaks; this is the experimental sample sheet.
+These two sample sheets are then provided as the input in the analysis below.
 
@@
 
-## Load convience functions
+## Load convenience functions
 
-These functions facilitate the normalisation of data.
+Brundle provies a series of functions we developed for this pipeline that are used to normalise ChIP-seq data. The usage of these functions are explained in the example that follows.
 
-```{r brundle}
+```{r brundle,results='hide', message=FALSE, warning=FALSE}
 source('../package/brundle.R')
@@
 ## Apply settings
+
+The intial settings for the analysis are minimal.
+The variable jg.controlMinOverlap is set to 5 as this is 1 less than the total sample count and we want a strong consensus of CTCF peaks to normalise against.
+The control and experimental sample sheets are as described above.
+The conditions reflect the treated and untreated conditions we have used when writing our sample sheet.
 
@@
 
-To keep file size down these are provided as a Rdata File rather than as raw counts.
+Typically we would load the raw data using DiffBind as we are planning to return our normalised data as a DiffBind object.
+
+In this example, as the BAM files are greater in size than the GitHub file size limit, we have included the data as the DiffBind object that these functions would output and instead of loading from the raw data directly.
+If the code does not find the R data object, it will try to regenerate it from the raw data.
 
@@
 
-```{r}
-#Load Sample Ids from control sample sheet.
-jg.sampleIds <- jg.getSampleIds(jg.controlSampleSheet)
-```
+DiffBind allows us to extract the counts in each peak. We use this function to obtain the data in a form that is easier to normalise.
 
@@
 ```
+
+To normalise, we need to separate samples out by condition.
+This is done using the following function.
+The peakset contains the reads in peaks; the sample sheet provides the metadata needed and the jg.treatedCondition/jg.untreatedCondition specifies which condition we want to get the counts for.
 
@@
 
-```{r}
+The normalisation factor is generated by a linear regression from the mean number of counts for each peak before and after treatment.
+The gradient of the linear fit should be equal to 1 given the assumption that, before and after treatment, the majority of the control peaks do not change.
 
-##Plot showing normalization calculation (Optional)
+This plot is not needed to calculate the normalisation factor but it does provide a visual interpretation of the process.
+
+```{r, message=FALSE}
+
+# Plot showing normalisation calculation (Optional)
+
 jg.plotNormalization(jg.controlCountsTreated,
@@
 
+There are two parts to the normalisation coefficent we use in the final correction.
+The first is the the coefficent to make the levels of CTCF binding before and after treament equal.
+The second is because DiffBind normalises against the total read count for each sample, we therefore generated a correction factor to control for this.
+
 ```{r}
-##Get Normalization Coefficient
+# Get Normalisation Coefficient
+
 jg.coefficient<-jg.getNormalizationCoefficient(jg.controlCountsTreated,
@@
 ```
+
+Once we have these two peices of information, we apply both the normalisation coeffficient and the correction factor to our data.
 
@@
 
+Finally we return the correct values into the DiffBind object.
+This requires a modified version DiffBind as included in the original manuscript.
+It is expected that this will be supported natively by future versions of DiffBind.
+
 ```{r}
-#Return values to Diffbind and plot normalised result.
+#Return values to DiffBind and plot normalised result.
 jg.dba <- DiffBind:::pv.resetCounts(dbaExperiment,

### 2017-08-14 examples/Example_002_ER_CTCF_DeSeq.Rmd
Edited examples
Correcting typos.

@@
 ---
-title: "Normalization Example using internal CTCF peaks as a standard (no-linear fit)"
+title: "Normalisation Example using Internal CTCF Peaks as a Standard (No-linear Fit)"
 author: "Andrew Holding"
-date: "8/7/2017"
+date: "7/8/2017"
 output: pdf_document
@@
 
-This is a worked example of our method alternative to normalize ER ChIP-seq data using CTCF binding to provide an internal control.
-Unlike our previous example this normalisation is using only DESeq Size Factors.
+This is a worked example of our alternative method to normalise ER ChIP-seq data using CTCF binding to provide an internal control.
+Unlike our previous example, this normalisation is using only DESeq Size Factors.
 This method was found to be less effective but was used for comparision in the supporting publication.
@@
 
-## Load convience functions
+## Load convenience functions
 
@@
 
-In this example Diffbind was simply used as a convient way to extract peak counts and provided consitancy for comparision between methods.
+In this example DiffBind was simply used as a convenient way to extract peak counts and provided consisteancy for comparision between methods.
 
-To keep file size down counts are provided as a Rdata file rather than as raw data.
+To keep file sizes down, counts are provided as a Rdata file rather than as raw data.
 
@@
 
-The actual DESeq Pipeline starts simply needs a matrix so we covert from Diffbind and save as csv to provide a starting point.
+The actual DESeq Pipeline starts simply needs a matrix so we convert from DiffBind and save as a CSV file to provide a starting point.
 
@@
   jg.experimentPeakset <- jg.dbaGetPeakset(dbaExperiment)
-  #Convert Peakset to DeSeq Workflow
+  #Convert Peakset to DESeq Workflow
   jg.experimentPeaksetDeSeq<-jg.convertPeakset(jg.experimentPeakset)
@@
 
-Here we use the CSV files the example code above generated; however, it is possible to generate the input data through a wide variety of methods and we have provided only one.
+Here we use the CSV files that the example code above generated; however, it is possible to generate the input data through a wide variety of methods and we have provided only one.
 
@@
 
-To normalize the data we use the DESeq2 function to estimate the size factors.
+To normalize the data, we use the DESeq2 function to estimate the size factors.
 This is a very simple method, and it provides some degree of normalization.
@@
 
-#Establish size factors directly from Control data
+#Establish size factors directly from control data
 jg.controlSizeFactors = estimateSizeFactorsForMatrix(jg.controlPeaksetDeSeq)
@@
 
-The Brundle package also contains some convience functions for plotting the output of DESeq2 inplace of DiffBind. These generate the following figures.
+The Brundle package also contains some convenience functions for plotting the output of DESeq2 in place of DiffBind. These generate the following figures.
 

### 2017-08-14 examples/Example_003_ER_H2av.Rmd
Correcting typos.

--- 2017-08-07 examples/Example_003_ER_H2av.Rmd
@@
 ---
-title: "Normalization Example using spike-in control as a standard"
+title: "Normalisation Example using Spike-in Control as a Standard"
 author: "Andrew Holding"
-date: "8/7/2017"
+date: "7/8/2017"
 output: pdf_document
@@
 
-This is a normalisation using H2av binding within an external spike-in of droshilia chromatin to provide a control for the analysis of ER binding in MCF7 cells after treatment with Fulvestrant.
+This is a normalisation using H2av binding within an external spike-in of drosophilia chromatin to provide a control for the analysis of ER binding in MCF7 cells after treatment with Fulvestrant.
 
-This example builds on the previous two and therefore these should be read first as many details of key functions have not been repeated.
+This example builds on the previous two examples and therefore they should be read first as many details of key functions have not been repeated here.
 
@@
 
-Unlike Example 001 & 002 the use of a xenogenic spike in requires more pre-processing.
+Unlike Examples 001 and 002 the use of a xenogenic spike-in requires more preprocessing.
 The scripts for this process are provided below.
@@
 
-The first step of our preprossing pipeline requires the generation of a combined Human/Drosophila genome.
-We used the following script to generated this.
+The first step of our preprossing pipeline requires the generation of a combined human/drosophila genome.
+We used the following script to generate this.
 The human and drosophila genomes downloaded from the link provided above are contained in a subdirectory "genomes" relative to where the script was run.
@@
 
-To allow use to use the merged genome we first indexed it using bowtie2. For our system we specified 32 threads, this will depend on the number of processors you have available.
+To allow us to use the merged genome we first indexed it using bowtie2. For our system, we specified 32 threads; the number of threads will depend on the number of processors you have available.
 
@@
 
-Once indexed, we could the aligned to the merge genome. This script assumes that the FastQ files are in a directory ./SLX-8047_dhms from where the scripts are run.
+Once indexed, we could then align to the merged genome. This script assumes that the FastQ files are in a directory ./SLX-8047_dhms from where the scripts are run.
 
 ```{bash, eval=FALSE}
-#Align, Sort and Build indeces
+#Align, Sort and Build indexes
 
@@
 
-After aligning the data can calling peaks we can then split the data back into Drosphila and Human.
+After aligning the data and calling peaks, we can then split the data back into human and drosophila.
 
@@
 
-With the peak files and the seperate alignments extract we are can then write samplesheets.
-For the normalization we need one for the drosphila aligned reads from each sample and the respective H2av peak file, and one for human reads from each sample and their repsective ER peak files.
+With the peak files and the separate alignments extracted, we can then write sample sheets.
+For the normalization we need one for the drosophila aligned reads from each sample and the respective H2av peak file, and one for human reads from each sample and their respsective ER peak files.
 This forms the bais of the following analysis.
 
-## Load convience functions
+## Load convenience functions
 
@@
 
-As before to keep file size down data are provided as a RdatafFile rather than as raw BAM files.
+As before, to keep file size down, data are provided in Rdata format rather than as raw BAM files.
 
@@
 
-##Plot showing normalization calculation (Optional)
+##Plot showing normalisation calculation (Optional)
   jg.plotNormalization(jg.controlCountsTreated,
@@
 ```{r}
-##Get Normalization Coefficient
+##Get Normalisation Coefficient
 jg.coefficient<-jg.getNormalizationCoefficient(jg.controlCountsTreated,
@@
 ```{r}
-#Return values to Diffbind and plot normalised result.
+#Return values to DiffBind and plot normalised result.
 jg.dba <- DiffBind:::pv.resetCounts(dbaExperiment,

### 2017-08-16 016_SLX-12998-MAplots.R
Figure 1 updated

@@ -39,7 +39,7 @@ A<-apply(hsrpm,1,function(x){
   return(log10(sum(x)))
 })
 png("plots/016_SLX-12998_MA_RPMpeaks.png")
-plot(A,M,pch=20,ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"),main="RPM reads in peaks")
+plot(A,M,pch=20,ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"),main="RPM reads in peaks",ylim=c(-6.25,4.25))
 abline(h=0)
 dev.off()
 
@@ -69,7 +69,7 @@ A<-apply(hsrpm,1,function(x){
   return(log10(sum(x)))
 })
 png("plots/016_SLX-12998_MA_RPMaligned.png")
-plot(A,M,pch=20,ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"),main="RPM aligned reads")
+plot(A,M,pch=20,ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"),main="RPM aligned reads",ylim=c(-6.25,4.25))
 abline(h=0)
 dev.off()
 
@@ -98,7 +98,7 @@ A<-apply(hsrpm,1,function(x){
   return(log10(sum(x)))
 })
 png("plots/016_SLX-12998_MA_RPMtotal.png")
-plot(A,M,pch=20,ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"),main="RPM total reads")
+plot(A,M,pch=20,ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"),main="RPM total reads",ylim=c(-6.25,4.25))
 abline(h=0)
 dev.off()
 
@@ -117,7 +117,7 @@ A<-apply(hscounts,1,function(x){
   return(log10(sum(x)))
 })
 png("plots/016_SLX-12998_MA_counts.png")
-plot(A,M,pch=20,ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"),main="Raw counts in peaks")
[…]

### 2017-08-16 029_SLX-14229_DeSeq_SizeFactors.R
Exported Control Peaks from SLX-14229 for motif analysi

@@ -158,4 +158,4 @@ nrow(jg.controlResultsDeseq[jg.DeSeqSignificant,][jg.DeSeqSignificantUp,])
 jg.DeSeqSignificantDown <- jg.controlResultsDeseq[jg.DeSeqSignificant,]$"log2FoldChange" > 0
 nrow(jg.controlResultsDeseq[jg.DeSeqSignificant,][jg.DeSeqSignificantDown,])
 
-
+write.csv(jg.controlResultsDeseq, file="csv/02_SLX14229_controlResultsDeseq.csv")

### 2017-08-16 029_SLX-8047_SLX-14229_Compare_DeSeq_SizeFactors.R
Quick plot tot comapre shape of CTCF and H2av normalizationw

@@ -1,0 +1,152 @@
+
+######################
+#
+# Functions 
+#
+######################
+
+source('package/brundle.R')
+
+######################
+#
+# Settings
+#
+######################
+
+setwd("/Volumes/FlyPeaks/FlyPeaks")
+dbaSummits                <- 200
+jg.controlMinOverlap      <- 5
+jg.controlSampleSheet     <- "samplesheet/samplesheet_SLX8047_dm.csv"
+jg.experimentSampleSheet  <- "samplesheet/samplesheet_SLX8047_hs.csv"
+
+
+######################
+#
+# Main Code
+#
+######################
+
+filename<-"Rdata/029_SLX-8047_dba_human_drosophila.rda"
+if(!file.exists(filename)){
+  dbaExperiment <- jg.getDba(jg.experimentSampleSheet,   dbaSummits)
+  dbaControl    <- jg.getDba(jg.controlSampleSheet,   dbaSummits)
+  save(dbaExperiment,dbaControl,file=filename)
+} else {
+  load(filename)
+}
+
+
+## Extract Peak set from DiffBind
+jg.experimentPeakset <- jg.dbaGetPeakset(dbaExperiment)
+jg.controlPeakset    <- jg.dbaGetPeakset(dbaControl)
+
+#Convert Peakset to DeSeq Workflow
+jg.controlPeaksetDeSeq<-jg.convertPeakset(jg.controlPeakset)
+
+#Establish size factors directly from Control data
+jg.controlSizeFactors = estimateSizeFactorsForMatrix(jg.controlPeaksetDeSeq)
+
+#Get conditions dataframe for DeSeq
+jg.conditions <- read.csv(file=jg.controlSampleSheet, header=TRUE, sep=",")['Condition']
+
+#Run DeSeq on control
+jg.controlDeSeq<-jg.runDeSeq(jg.controlPeaksetDeSeq, jg.conditions,jg.controlSizeFactors)
+jg.controlResultsDeseq   = results(jg.controlDeSeq)
+
+#Repeat for experimental conditions
+
+#Convert experiment Peakset to DeSeq Workflow
+jg.experimentPeaksetDeSeq<-jg.convertPeakset(jg.experimentPeakset)
+
+#Run DeSeq on experiment
+jg.experimentDeSeq<-jg.runDeSeq(jg.experimentPeaksetDeSeq, jg.conditions,jg.controlSizeFactors)
[…]

### 2017-08-17 022_RARA_gene.R
Scaled RARA locus figure to CTCF height

@@ -98,3 +98,12 @@ dev.off()
 pdf("plots/022_RARA_locus_min.pdf",pointsize=3)
 plotTracks(list(ideoTrack, axisTrack,dTrack1a,dTrack1b,CTCFa,  txTranscripts_v1), from =  windowStart, to = windowEnd)
 dev.off()
+
+
+dTrack1a_scaled <- DataTrack(range = bamFile, genome = "hg19", type = "l", name = "ICI1", window = -1, chromosome = chromosome, col="red", ylim=c(0,96),showAxis=F)
+CTCFa_scaled <- DataTrack(range = bamFile, genome = "hg19", type = "l", name = "CTCF", window = -1, chromosome = chromosome, col="purple",ylim=c(0,96),showAxis=F)
+pdf("plots/022_RARA_locus_min_scaled.pdf",pointsize=3)
+plotTracks(list(ideoTrack, axisTrack,dTrack1a_scaled,dTrack1b,CTCFa_scaled,  txTranscripts_v1), from =  windowStart, to = windowEnd)
+dev.off()
+
+

### 2017-08-17 022_RARA_gene.R
Scaled RARA locus figure to CTCF height

@@ -99,9 +99,9 @@ pdf("plots/022_RARA_locus_min.pdf",pointsize=3)
 plotTracks(list(ideoTrack, axisTrack,dTrack1a,dTrack1b,CTCFa,  txTranscripts_v1), from =  windowStart, to = windowEnd)
 dev.off()
 
-
-dTrack1a_scaled <- DataTrack(range = bamFile, genome = "hg19", type = "l", name = "ICI1", window = -1, chromosome = chromosome, col="red", ylim=c(0,96),showAxis=F)
 CTCFa_scaled <- DataTrack(range = bamFile, genome = "hg19", type = "l", name = "CTCF", window = -1, chromosome = chromosome, col="purple",ylim=c(0,96),showAxis=F)
+bamFile <- "SLX-14229_mmhs/blacklist_filtered/human/SLX-14229.D701_D503.HJJL7BBXX.s_8.r_1.fq.gz.bam"
+dTrack1a_scaled <- DataTrack(range = bamFile, genome = "hg19", type = "l", name = "ICI1", window = -1, chromosome = chromosome, col="red", ylim=c(0,96),showAxis=F)
 pdf("plots/022_RARA_locus_min_scaled.pdf",pointsize=3)
 plotTracks(list(ideoTrack, axisTrack,dTrack1a_scaled,dTrack1b,CTCFa_scaled,  txTranscripts_v1), from =  windowStart, to = windowEnd)
 dev.off()

### 2017-08-17 032_SLX-14229_SLX-8045_compare.R
Made fold change comparision include both all points.

@@ -186,9 +186,9 @@ plot(
 dba.report(jg.dba_analysis_SLX8047)
 dba.report(jg.dba_analysis_SLX14229)
 
-nearestPeak <- nearest(dba.report(jg.dba_analysis_SLX8047),dba.report(jg.dba_analysis_SLX14229))
-SLX8047_fc<-dba.report(jg.dba_analysis_SLX8047,bFlip=TRUE)$Fold
-SLX14229_fc<-dba.report(jg.dba_analysis_SLX14229)[nearestPeak]$Fold
+nearestPeak <- nearest(dba.report(jg.dba_analysis_SLX8047, th=1),dba.report(jg.dba_analysis_SLX14229, th=1))
+SLX8047_fc<-dba.report(jg.dba_analysis_SLX8047,bFlip=TRUE, th=1)$Fold
+SLX14229_fc<-dba.report(jg.dba_analysis_SLX14229,th=1)[nearestPeak]$Fold
 
 
 png("plots/032_normalization_comparision.png")
@@ -200,9 +200,26 @@ abline(h=0, col="grey",lty=2)
 abline(v=0, col="grey",lty=2)
 coef<-lm( SLX8047_fc ~ SLX14229_fc + 0 )
 abline(coef, col="navyblue")
-text(3.1,-0.25, expression( "gradient = 0.96, r = 0.503, p-value < 2.2 x" ~ 10^{-16}))
+text(3.1,-0.25, expression( "gradient = 0.96, r = 0.557, p-value < 2.2 x" ~ 10^{-16}))
 dev.off()     
 
 
 result<-cor.test(SLX8047_fc,SLX14229_fc, method="pearson")
+nearestPeak <- nearest(dba.report(jg.dba_analysis_SLX8047,th=1),dba.report(jg.dba_analysis_SLX14229,th=1))
+SLX8047_pv<-dba.report(jg.dba_analysis_SLX8047,bFlip=TRUE,th=1)$"p-value"
+SLX14229_pv<-dba.report(jg.dba_analysis_SLX14229, th=1)[nearestPeak]$"p-value"
+
+
+png("plots/032_normalization_pvalue_comparision.png")
+smoothScatter(-log(SLX8047_pv),-log(SLX14229_pv),pch=20, col="#00000055",
+              xlab="-log(p-value) after normalization using Drosophila/H2av spike-in",
+              ylab="-log(p-value) after normalization using CTCF control peaks",
[…]

### 2017-08-17 ENCODE/000_download_encode.sh
Added code to download ENCODE data for figure

@@ -1,0 +1,2 @@
+wget http://www.encodeproject.org/files/ENCFF591QUF/@@download/ENCFF591QUF.bam --no-check-certificate
+wget http://www.encodeproject.org/files/ENCFF783NXW/@@download/ENCFF783NXW.bam --no-check-certificate

### 2017-11-06 054_SLX-14438_peakscatter.R
Scatter plot for IDR

@@ -7,7 +7,7 @@
 setwd("/Volumes/FlyPeaks/FlyPeaks")
 source('package/brundle.R')
 
-jg.plotScatter<-function(peaks,samples,yaxis,samplesCombinations)
+jg.plotScatter<-function(peaks,samples,yaxis,samplesCombinations,noYLabel=FALSE)
 {
     peaks$count <- 1
     peaks_aggregate <- aggregate(count ~ ., peaks, FUN = sum)
@@ -16,11 +16,11 @@ jg.plotScatter<-function(peaks,samples,yaxis,samplesCombinations)
     peaks_aggregate$Col <- colormap[as.numeric(cut(log(peaks_aggregate$count,2),breaks=255))]
     peaks_aggregate<-peaks_aggregate[order(peaks_aggregate$count),]
     plot(log(peaks_aggregate[,c(1,2)]), col=peaks_aggregate$Col,pch=20, cex=0.3, xlim=c(0,7),ylim=c(0,7),axes=FALSE)
-    if (samples > 1) {Axis(side=2, labels=FALSE)} else {
+    if (samples > 1|| noYLabel==TRUE) {Axis(side=2, labels=FALSE)} else {
         Axis(side=2, labels=TRUE)
         mtext("log(Counts)", side=2, line=3)
-        }
-    if (yaxis !=TRUE) {Axis(side=1, labels=FALSE)} else {
+    }
+    if (yaxis !=TRUE ) {Axis(side=1, labels=FALSE)} else {
         Axis(side=1, labels=TRUE,ylab="Log(Counts)")
         mtext("log(Counts)", side=1, line=3)
     }
@@ -29,6 +29,45 @@ jg.plotScatter<-function(peaks,samples,yaxis,samplesCombinations)
     text(2,6.0,paste0("Correlation = ",signif(sampleCor,3)),cex=1.0)
         abline(a=0,b=1,col="grey")
     return(cor(log(peaks_aggregate[,c(1,2)])))
+}
+
+
+jg.plotMargins<-function(a,b)
+{
+    par(mar=c(0.5, 0.5, 0.2, 0.2), mfrow=c(a,b),
+        oma = c(4, 4, 0.2, 0.2))
+}
+jg.plotPanelsPlot<-function(treatmentNames,controlNames,dbaExperiment,noYLabel=FALSE)
+{
+    
+    jg.experimentPeakset <- jg.dbaGetPeakset(dbaExperiment)
+    relativeLibrarySize<-jg.getLibrarySize(dbaExperiment,
[…]

### 2017-11-07 Manuscript_Figures.Rmd
Removal of ER proximal CTCF sites

---
title: "Brundle Development"
author: "Andrew Holding"
date: "11/7/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, messages=FALSE)
```

## Comparison of simple normalisation methods

```{r}
library(DiffBind)
#setwd("/Volumes/FlyPeaks/FlyPeaks")

hsconsensus<-readRDS("Rdata/015_SLX-12998_hsconsensus.rds")
mmconsensus<-readRDS("Rdata/015_SLX-12998_mmconsensus.rds")
hscounts<-hsconsensus[,-c(1:3)]
mmcounts<-mmconsensus[,-c(1:3)]
### MA RPM in peaks
hsrpm<-apply(hscounts,2,function(x){
  1E6*x/sum(x)
})
mmrpm<-apply(mmcounts,2,function(x){
  1E6*x/sum(x)
})
M<-apply(hsrpm,1,function(x){
  fulvestrant<-mean(x[c(1,2,4,5)])
  untreated<-mean(x[c(3,6,7,8)])
  fc<-mean(fulvestrant)/mean(untreated)
  log2fc<-log2(fc)
  return(log2fc)
})
A<-apply(hsrpm,1,function(x){
  return(log10(sum(x)))
})

plot(A,M,cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5,pch=20,ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"),main="RPM reads in peaks",ylim=c(-6.25,4.25))
abline(h=0)

```

```{r}

load("Rdata/012_SLX-12998_aligned.rda")
aligned<-aligned[grep("SLX-12998.D",names(aligned))]
aligned<-sapply(aligned,sum)/1E6
aligned<-aligned[c(4,7,8,9,2,3,6,5)]

hsrpm<-hscounts
for(i in 1:length(hsrpm)){
  hsrpm[i]<-hscounts[i]/aligned[i]
}
mmrpm<-mmcounts
for(i in 1:length(mmrpm)){
  mmrpm[i]<-mmcounts[i]/aligned[i]
}
M<-apply(hsrpm,1,function(x){
  fulvestrant<-mean(x[c(1,2,4,5)])
  untreated<-mean(x[c(3,6,7,8)])
  fc<-mean(fulvestrant)/mean(untreated)
  log2fc<-log2(fc)
  return(log2fc)
})
A<-apply(hsrpm,1,function(x){
  return(log10(sum(x)))
})

plot(A,M,cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5,cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5,pch=20,ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"),main="RPM aligned reads",ylim=c(-6.25,4.25))
abline(h=0)

```

```{r}

load("Rdata/012_SLX-12998_nrreads.rda")
nrreads<-nrreads[grep("SLX-12998.D",names(nrreads))]
nrreads<-nrreads/1E6
nrreads<-nrreads[c(4,7,8,9,2,3,6,5)]
hsrpm<-hscounts
for(i in 1:length(hsrpm)){
  hsrpm[i]<-hscounts[i]/nrreads[i]
}
mmrpm<-mmcounts
for(i in 1:length(mmrpm)){
  mmrpm[i]<-mmcounts[i]/nrreads[i]
}
M<-apply(hsrpm,1,function(x){
  fulvestrant<-mean(x[c(1,2,4,5)])
  untreated<-mean(x[c(3,6,7,8)])
  fc<-mean(fulvestrant)/mean(untreated)
  log2fc<-log2(fc)
  return(log2fc)
})
A<-apply(hsrpm,1,function(x){
  return(log10(sum(x)))
})

plot(A,M,cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5,pch=20,ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"),main="RPM total reads",ylim=c(-6.25,4.25))
abline(h=0)

```

```{r}
### MA Counts human
# a is treated, b is not treated
M<-apply(hscounts,1,function(x){
  fulvestrant<-mean(x[c(1,2,4,5)])
  untreated<-mean(x[c(3,6,7,8)])
  fc<-mean(fulvestrant)/mean(untreated)
  log2fc<-log2(fc)
  return(log2fc)
})
A<-apply(hscounts,1,function(x){
  return(log10(sum(x)))
})

par(mar=c(5.1,5.1,4.1,2.1))
plot(A,M,cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5,pch=20,ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"),main="Raw counts in peaks", ylim=c(-6.25,4.25))
abline(h=0)

```

## MA Plot of H2AV normalization
```{r}
# Read consensus peaks
hsconsensus<-readRDS("Rdata/015_SLX-8047_hsconsensus.rds")
dmconsensus<-readRDS("Rdata/015_SLX-8047_dmconsensus.rds")

# Get counts
hscounts<-hsconsensus[,-c(1:3)]
dmcounts<-dmconsensus[,-c(1:3)]

### MA for Drosophila + Human Counts
# a is treated, b is not treated
Mhs<-apply(hscounts,1,function(x){
  fulvestrant<-mean(x[c(1,3,5,7)])
  untreated<-mean(x[c(2,4,6,8)])
  fc<-mean(fulvestrant)/mean(untreated)
  log2fc<-log2(fc)
  return(log2fc)
})
Ahs<-apply(hscounts,1,function(x){
  return(log10(sum(x)))
})
Mdm<-apply(dmcounts,1,function(x){
  fulvestrant<-mean(x[c(1,3,5,7)])
  untreated<-mean(x[c(2,4,6,8)])
  fc<-mean(fulvestrant)/mean(untreated)
  log2fc<-log2(fc)
  return(log2fc)
})
Adm<-apply(dmcounts,1,function(x){
  return(log10(sum(x)))
})

plot(Ahs,Mhs,cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5,pch=20,ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"),main="RPM aligned reads",ylim=c(-6.25,2))
points(Adm,Mdm,col="cornflowerblue",pch=20)
abline(h=0)
legend("topright",legend=c("Drosophila","Human"),pch=20,col=c("cornflowerblue","black"))
lm1<-lm(Mdm~Adm)
abline(lm1$coef,col="red")
```

```{r}
### 20th century normalization
# Residuals of the Drosophila fit
lm1<-lm(Mdm~Adm)
intercept<-lm1$coef[1]
angularcoeff<-lm1$coef[2]
MhsFit<-Mhs-(Ahs*angularcoeff)-intercept
MdmFit<-Mdm-(Adm*angularcoeff)-intercept

plot(Ahs,MhsFit,cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5,pch=20,ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"),main="Counts normalized by H2av Distribution",ylim=c(-6.25,2))
points(Adm,MdmFit,pch=20,col="cornflowerblue")
abline(h=0)
legend("topright",legend=c("Drosophila","Human"),pch=20,col=c("cornflowerblue","black"))
lm1<-lm(MdmFit~Adm)
abline(lm1$coef,col="red")

```

## DeSeq Results

```{r}

source('package/brundle.R')
## Changing font sizes for manuscript

jg.plotDeSeq<-function(ma.df, filename = 'file.name', p = 0.01, title.main = "Differential ChIP",log2fold =0.5, flip=FALSE)
{;

    if (flip == TRUE)
    {
        ma.df$log2FoldChange <- -ma.df$log2FoldChange
    }
    par(mar=c(5.1,5.1,4.1,2.1))
    xyplot(ma.df$log2FoldChange ~ log(ma.df$baseMean, base=10),#xlim=c(0,4),ylim=c(-3,1.25),
           groups=(ma.df$padj < p & abs(ma.df$log2FoldChange) > log2fold & !is.na(ma.df$padj)),
           col=c("black","red"), main=title.main, scales="free", aspect=1, pch=20, cex=0.5,
           ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"),
           par.settings=list(axis.text=list(cex=1.5,font=1),par.main.text=list(cex=1.5,font=2),par.xlab.text=list(cex=1.5,font=2), par.ylab.text=list(cex=1.5,font=2)
           )
    );

}

jg.plotDeSeqCombined <- function(jg.controlResultsDeseq,jg.experimentResultsDeseq,title.main,padjX,flip=FALSE)
{
    jg.controlResultsDeseq$group = 'a'
    jg.experimentResultsDeseq$group = 'b'

    if (flip == TRUE)
    {
        jg.controlResultsDeseq$log2FoldChange <- -jg.controlResultsDeseq$log2FoldChange
        jg.experimentResultsDeseq$log2FoldChange <- -jg.experimentResultsDeseq$log2FoldChange
    }

    for (i in 1:length(jg.experimentResultsDeseq$group)) {
        if (!is.na(jg.experimentResultsDeseq$padj[i]) & !is.na(jg.experimentResultsDeseq$log2FoldChange[i]) & jg.experimentResultsDeseq$padj[i] < padjX & jg.experimentResultsDeseq$log2FoldChange[i] < 0) {
            jg.experimentResultsDeseq$group[i] <- 'd'
        }
        else if (!is.na(jg.experimentResultsDeseq$padj[i]) & !is.na(jg.experimentResultsDeseq$log2FoldChange[i]) & jg.experimentResultsDeseq$padj[i] < padjX & jg.experimentResultsDeseq$log2FoldChange[i] > 0) {
            jg.experimentResultsDeseq$group[i] <- 'c'
        }
    }

    for (i in 1:length(jg.controlResultsDeseq$group)) {
        if (!is.na(jg.controlResultsDeseq$padj[i]) & !is.na(jg.controlResultsDeseq$log2FoldChange[i]) & jg.controlResultsDeseq$padj[i] < padjX & jg.controlResultsDeseq$log2FoldChange[i] < 0) {
            jg.controlResultsDeseq$group[i] <- 'f'
        }
        else if (!is.na(jg.controlResultsDeseq$padj[i]) & !is.na(jg.controlResultsDeseq$log2FoldChange[i]) & jg.controlResultsDeseq$padj[i] < padjX & jg.controlResultsDeseq$log2FoldChange[i] > 0) {
            jg.controlResultsDeseq$group[i] <- 'e'
        }
    }

    full.res = rbind(jg.controlResultsDeseq, jg.experimentResultsDeseq)
    par(mar=c(5.1,5.1,4.1,2.1))
    xyplot(full.res$log2FoldChange ~ log(full.res$baseMean, base=10), data = full.res,
           groups=full.res$group,
           col=c("grey40","grey80",  "#5480ff", "#ff5454", "#08298a","#750505"),
           ylab = expression('log'[2]*' Differential ChIP'),
           xlab = expression("log"[10]~"Mean of Normalized Counts"),
           aspect=1.0,
           pch=16,
           cex=0.5,
           main=title.main,
           scales=list(x=list(cex=1.5, relation = "free"), y =list(cex=1.5, relation="free")),
           between=list(y=0.5, x=0.5),
           auto.key = TRUE,
           par.settings=list(axis.text=list(cex=1.5,font=1),par.main.text=list(cex=1.5,font=2),par.xlab.text=list(cex=1.5,font=2), par.ylab.text=list(cex=1.5,font=2)),
           key=list(corner=c(1,0),
                    cex=1.0,
                    points=list(col=c( "gray80","gray40", "#ff5454", "#5480ff", "#750505", "#08298a","white"), pch=20),
                    text=list(c("Target Binding","Control Binding", "Target Decreased","Target Increased","Control Decreased","Control Increased", " "))
           ))

}

######################
#
# Settings
#
######################

setwd("/Volumes/FlyPeaks/FlyPeaks")
dbaSummits                <- 200
jg.controlMinOverlap      <- 5
jg.controlSampleSheet     <- "samplesheet/samplesheet_SLX14229_hs_CTCF_DBA.csv"
jg.experimentSampleSheet  <- "samplesheet/samplesheet_SLX14229_hs_ER_DBA.csv"

filename<-"Rdata/028_SLX-14229_dba_human_ER_CTCF.rda"
load(filename)

## Extract Peak set from DiffBind
jg.experimentPeakset <- jg.dbaGetPeakset(dbaExperiment)
jg.controlPeakset    <- jg.dbaGetPeakset(dbaControl)

#Convert Peakset to DeSeq Workflow
jg.controlPeaksetDeSeq<-jg.convertPeakset(jg.controlPeakset)

#Establish size factors directly from Control data
jg.controlSizeFactors = estimateSizeFactorsForMatrix(jg.controlPeaksetDeSeq)

#Get conditions dataframe for DeSeq
jg.conditions <- read.csv(file=jg.controlSampleSheet, header=TRUE, sep=",")['Condition']

#Run DeSeq on control
jg.controlDeSeq<-jg.runDeSeq(jg.controlPeaksetDeSeq, jg.conditions,jg.controlSizeFactors)
jg.controlResultsDeseq   = results(jg.controlDeSeq)

#Repeat for experimental conditions

#Convert experiment Peakset to DeSeq Workflow
jg.experimentPeaksetDeSeq<-jg.convertPeakset(jg.experimentPeakset)

#Run DeSeq on experiment
jg.experimentDeSeq<-jg.runDeSeq(jg.experimentPeaksetDeSeq, jg.conditions,jg.controlSizeFactors)
jg.experimentResultsDeseq   = results(jg.experimentDeSeq)

  jg.plotDeSeq(jg.controlResultsDeseq,
               p=0.01, 
               title.main="Fold-change in CTCF binding",
               flip=T
               )

```

```{r}
jg.plotDeSeq(jg.experimentResultsDeseq,
             p=0.01,
             title.main="Fold-change in ER binding",
             flip=T
             )
```

#CTCF Binding

```{r}

jg.experimentDeSeqInternal<-jg.runDeSeq(jg.experimentPeaksetDeSeq, jg.conditions, NULL)
jg.experimentResultsDeseqInternal   = results(jg.experimentDeSeqInternal)

 jg.plotDeSeq(jg.experimentResultsDeseqInternal,
               title.main="Fold-change in ER binding (no correction)",
               p=0.01,
               flip=T
               )
```

#H2av With Diff Bind
```{r}
jg.controlSampleSheet     <- "samplesheet/samplesheet_SLX8047_dm.csv"
jg.experimentSampleSheet  <- "samplesheet/samplesheet_SLX8047_hs.csv"
filename<-"Rdata/029_SLX-8047_dba_human_drosophila.rda"

load(filename)

## Extract Peak set from DiffBind
jg.experimentPeakset <- jg.dbaGetPeakset(dbaExperiment)
jg.controlPeakset    <- jg.dbaGetPeakset(dbaControl)

#Convert Peakset to DeSeq Workflow
jg.controlPeaksetDeSeq<-jg.convertPeakset(jg.controlPeakset)

#Establish size factors directly from Control data
jg.controlSizeFactors = estimateSizeFactorsForMatrix(jg.controlPeaksetDeSeq)

#Get conditions dataframe for DeSeq
jg.conditions <- read.csv(file=jg.controlSampleSheet, header=TRUE, sep=",")['Condition']

#Run DeSeq on control
jg.controlDeSeq<-jg.runDeSeq(jg.controlPeaksetDeSeq, jg.conditions,jg.controlSizeFactors)
jg.controlResultsDeseq   = results(jg.controlDeSeq)

#Repeat for experimental conditions

#Convert experiment Peakset to DeSeq Workflow
jg.experimentPeaksetDeSeq<-jg.convertPeakset(jg.experimentPeakset)

#Run DeSeq on experiment
jg.experimentDeSeq<-jg.runDeSeq(jg.experimentPeaksetDeSeq, jg.conditions,jg.controlSizeFactors)
jg.experimentResultsDeseq   = results(jg.experimentDeSeq)

jg.experimentDeSeqInternal<-jg.runDeSeq(jg.experimentPeaksetDeSeq, jg.conditions, NULL)
jg.experimentResultsDeseqInternal   = results(jg.experimentDeSeqInternal)

dbaExperimentCorrected<-jg.correctDBASizeFactors(dbaExperiment,
                                                 jg.controlSizeFactors
                                                 )

dbaExperimentAnalysis<-dba.analyze(dbaExperiment)
  dba.plotMA(dbaExperimentAnalysis)

```

```{r}
  dbaExperimentAnalysisCorrected<-dba.analyze(dbaExperimentCorrected)
  dba.plotMA(dbaExperimentAnalysisCorrected)
```
```{r}

filename<-"Rdata/040_SLX-14438_dba_human_ER_CTCF.rda"
  load(filename)
  jg.controlSampleSheet     <- "samplesheet/samplesheet_SLX14438_hs_CTCF_DBA.csv"
jg.experimentSampleSheet  <- "samplesheet/samplesheet_SLX14438_hs_ER_DBA.csv"
jg.treatedCondition       =  "Fulvestrant"
jg.untreatedCondition     =  "none"

#Get counts for each condition
jg.controlCountsTreated<-jg.getControlCounts(jg.controlPeakset, 
                                             jg.controlSampleSheet,
                                             jg.treatedCondition)
jg.controlCountsUntreated<-jg.getControlCounts(jg.controlPeakset,
                                               jg.controlSampleSheet,
                                               jg.untreatedCondition)

#Get sample names for conditions
jg.untreatedNames <- names(jg.controlCountsUntreated)
jg.treatedNames   <- names(jg.controlCountsTreated)

jg.plotNormalization(jg.controlCountsTreated, jg.controlCountsUntreated)
```
```{r}

jg.controlSampleSheet     <- "samplesheet/samplesheet_SLX8047_dm.csv"
jg.experimentSampleSheet  <- "samplesheet/samplesheet_SLX8047_consensus.csv"

filename<-"Rdata/029_SLX-8047_SLX14229_dba_human_drosophila.rda"
load(filename)

## Extract Peak set from DiffBind
jg.experimentPeakset <- jg.dbaGetPeakset(dbaExperiment)
jg.controlPeakset    <- jg.dbaGetPeakset(dbaControl)

#Convert Peakset to DeSeq Workflow
jg.controlPeaksetDeSeq<-jg.convertPeakset(jg.controlPeakset)

#Establish size factors directly from Control data
jg.controlSizeFactors = estimateSizeFactorsForMatrix(jg.controlPeaksetDeSeq)

#Get conditions dataframe for DeSeq
jg.conditions <- read.csv(file=jg.controlSampleSheet, header=TRUE, sep=",")['Condition']

#Run DeSeq on control
jg.controlDeSeq<-jg.runDeSeq(jg.controlPeaksetDeSeq, jg.conditions,jg.controlSizeFactors)
jg.controlResultsDeseq   = results(jg.controlDeSeq)

#Repeat for experimental conditions

#Convert experiment Peakset to DeSeq Workflow
jg.experimentPeaksetDeSeq<-jg.convertPeakset(jg.experimentPeakset)

#Run DeSeq on experiment
jg.experimentDeSeq<-jg.runDeSeq(jg.experimentPeaksetDeSeq, jg.conditions,jg.controlSizeFactors)
jg.experimentResultsDeseq_SLX8047   = results(jg.experimentDeSeq)

#Now run for SLX-14229

jg.controlSampleSheet     <- "samplesheet/samplesheet_SLX14229_hs_CTCF_DBA.csv"
#Using the same consensus peak set from above 
jg.experimentSampleSheet  <- "samplesheet_SLX14229_hs_ER_consensus.csv"

filename<-"Rdata/029_SLX-8047_SLX-14229_dba_human_ER_CTCF_consensus.rda"
load(filename)

## Extract Peak set from DiffBind
jg.experimentPeakset <- jg.dbaGetPeakset(dbaExperiment)
jg.controlPeakset    <- jg.dbaGetPeakset(dbaControl)

#Convert Peakset to DeSeq Workflow
jg.controlPeaksetDeSeq<-jg.convertPeakset(jg.controlPeakset)

#Establish size factors directly from Control data
jg.controlSizeFactors = estimateSizeFactorsForMatrix(jg.controlPeaksetDeSeq)

#Get conditions dataframe for DeSeq
jg.conditions <- read.csv(file=jg.controlSampleSheet, header=TRUE, sep=",")['Condition']

#Run DeSeq on control
jg.controlDeSeq<-jg.runDeSeq(jg.controlPeaksetDeSeq, jg.conditions,jg.controlSizeFactors)
jg.controlResultsDeseq   = results(jg.controlDeSeq)

#Repeat for experimental conditions

#Convert experiment Peakset to DeSeq Workflow
jg.experimentPeaksetDeSeq<-jg.convertPeakset(jg.experimentPeakset)

#Run DeSeq on experiment
jg.experimentDeSeq<-jg.runDeSeq(jg.experimentPeaksetDeSeq, jg.conditions,jg.controlSizeFactors)
jg.experimentResultsDeseq_SLX14229   = results(jg.experimentDeSeq)

library(latticeExtra)

ma.df<-jg.experimentResultsDeseq_SLX14229

a<-    xyplot((-ma.df$log2FoldChange+0.2) ~ log(ma.df$baseMean, base=10),
              panel=function(...) {
                  panel.xyplot(...)
                  panel.abline(h=0, lty = "dotted", col = "black")
                  panel.segments(3.0,0,1,1.6, col="red",lwd=2)
                  panel.segments(3.0,0,1,-1.6, col="red",lwd=2)
                  panel.segments(1,1.6,1,-1.6, col="red",lwd=2)
              },
           col=c("deepskyblue4"), ylim=c(-6,2), main="Comparision between CTCF and H2av", scales="free", aspect=1, pch=20, cex=0.5,
           ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"),
           par.settings=list(par.main.text=list(cex=1.57,font=2),axis.text=list(cex=1.57,font=1),par.xlab.text=list(cex=1.57,font=2), par.ylab.text=list(cex=1.57,font=2)));

ma.df_2<-jg.experimentResultsDeseq_SLX8047
#Note this has been manually addjusted to be more illustrative.
b<-    xyplot(-(ma.df_2$log2FoldChange*1.1-0.1) ~ (log(ma.df_2$baseMean, base=10)*1.1+0.2),
              panel=function(...) {
                  panel.xyplot(...)
                  panel.abline(h=0, lty = "dotted", col = "black")
                  panel.segments(3.0,0,1,1.6, col="red",lwd=2)
                  panel.segments(3.0,0,1,-1.6, col="red",lwd=2)
                  panel.segments(1,1.6,1,-1.6, col="red",lwd=2)
              },
              col=c("palegreen3"), main="Comparision between CTCF and H2av", scales="free", aspect=1, pch=20, cex=0.5,
              ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"),
              par.settings=list(par.xlab.text=list(cex=1.1,font=2), par.ylab.text=list(cex=1.1,font=2)));

a + as.layer(b)

```
```{r}

plot(jg.experimentResultsDeseq_SLX14229$log2FoldChange,
     jg.experimentResultsDeseq_SLX8047$log2FoldChange,
     pch=20,
     cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5,
     main="Comparision of fold change between datasets",
     xlab=expression("log"[2]~" ER ChIP fold change CTCF normalised"),
     ylab=expression("log"[2]~" ER ChIP fold change H2av normalised"))

```

```{r}

jg.plotScatter<-function(peaks,samples,yaxis,samplesCombinations,noYLabel=FALSE)
{
    peaks$count <- 1
    peaks_aggregate <- aggregate(count ~ ., peaks, FUN = sum)
    colormap<-rainbow(255,start=0,end=4/6)
    colormap<-rev(colormap)
    peaks_aggregate$Col <- colormap[as.numeric(cut(log(peaks_aggregate$count,2),breaks=255))]
    peaks_aggregate<-peaks_aggregate[order(peaks_aggregate$count),]
    plot(log(peaks_aggregate[,c(1,2)]), col=peaks_aggregate$Col,pch=20, cex=0.3, xlim=c(0,7),ylim=c(0,7),axes=FALSE)
    if (samples > 1|| noYLabel==TRUE) {Axis(side=2, labels=FALSE)} else {
        Axis(side=2, labels=TRUE)
        mtext("log(Counts)", side=2, line=3)
    }
    if (yaxis !=TRUE ) {Axis(side=1, labels=FALSE)} else {
        Axis(side=1, labels=TRUE,ylab="Log(Counts)")
        mtext("log(Counts)", side=1, line=3)
    }
    text(1.5,6.5,paste0(samplesCombinations[,samples][1],"-",samplesCombinations[,samples][2]),cex=1.5)
    sampleCor<-cor(log(peaks[1]),log(peaks[2]))
    text(2,6.0,paste0("Correlation = ",signif(sampleCor,3)),cex=1.0)
        abline(a=0,b=1,col="grey")
    return(cor(log(peaks_aggregate[,c(1,2)])))
}

jg.plotMargins<-function(a,b)
{
    par(mar=c(0.5, 0.5, 0.2, 0.2), mfrow=c(a,b),
        oma = c(4, 4, 0.2, 0.2))
}
jg.plotPanelsPlot<-function(treatmentNames,controlNames,dbaExperiment,noYLabel=FALSE)
{

    jg.experimentPeakset <- jg.dbaGetPeakset(dbaExperiment)
    relativeLibrarySize<-jg.getLibrarySize(dbaExperiment,
                                            jg.experimentPeakset)

    samplesCombinations<-combn(controlNames,2)
    for (samples in seq(1,length(samplesCombinations)/2))  {
        peaksLibrarySize<-relativeLibrarySize[samplesCombinations[,samples]]
        peaks<-jg.experimentPeakset[samplesCombinations[,samples]]
        peaks[1]<-round(peaks[1]/peaksLibrarySize[1])
        peaks[2]<-round(peaks[2]/peaksLibrarySize[2])
        jg.plotScatter(peaks,samples, F,samplesCombinations,noYLabel)
    }
    samplesCombinations<-combn(treatmentNames,2)
    for (samples in seq(1,length(samplesCombinations)/2))  {
        peaksLibrarySize<-relativeLibrarySize[samplesCombinations[,samples]]
        peaks<-jg.experimentPeakset[samplesCombinations[,samples]]
        peaks[1]<-round(peaks[1]/peaksLibrarySize[1])
        peaks[2]<-round(peaks[2]/peaksLibrarySize[2])
        jg.plotScatter(peaks,samples,T,samplesCombinations,noYLabel)
    }
 }

jg.getLibrarySize<-function(dbaExperiment,jg.experimentPeakset) {
    librarySize<-as.numeric(dbaExperiment$class["Reads",])
    relativeLibrarySize<-librarySize/max(librarySize)
    names(relativeLibrarySize)<-colnames(jg.experimentPeakset[c(-1:-3)])
    return(relativeLibrarySize)
}

jg.controlSampleSheet     <- "samplesheet/samplesheet_SLX14438_hs_CTCF_DBA.csv"
jg.experimentSampleSheet  <- "samplesheet/samplesheet_SLX14438_hs_ER_DBA.csv"

filename<-"Rdata/048_SLX-14438_dba_human_ER_CTCF.rda"
load(filename)

controlNames<-c("1a","2a","3a")
treatmentNames<-c("1b","2b","3b")

filename<-"Rdata/026_SLX-8047_dba.counts_human.rda"
load(filename)
dbaExperiment_SLX8047<-dba

controlNames_SLX8047<-c("1b","2b","3b","4b")
treatmentNames_SLX8047<-c("1a","2a","3a","4a")

filename<-"Rdata/047_SLX-12998_dba_human.rda"
load(filename)
dbaExperiment_SLX12998<-dbaFullExperiment

controlNames_SLX12998<-c("1b","2b","3b","4b")
treatmentNames_SLX12998<-c("1a","2a","3a","4a")

jg.plotMargins(2,3)
par(mfcol=c(2,3))

#SLX-14438
#2a-3a and 1b-3b
controlNames<-c("2a","3a")
treatmentNames<-c("1b","3b")
jg.plotPanelsPlot(treatmentNames,
                  controlNames,
                  dbaExperiment)

#SLX-8047
#3b-4b and 1a-4a
controlNames<-c("3b","4b")
treatmentNames<-c("1a","4a")
jg.plotPanelsPlot(treatmentNames,
                  controlNames,
                  dbaExperiment_SLX8047, TRUE)

#SLX-12998
#3b-4b and 1a-4a
controlNames<-c("3b","4b")
treatmentNames<-c("1a","4a")
jg.plotPanelsPlot(treatmentNames,
                  controlNames,
                  dbaExperiment_SLX12998,TRUE)
```

```{r}
filename<-"Rdata/040_SLX-14438_dba_human_ER_CTCF.rda"
load(filename)

dba.plotVenn(dbaControl,dbaControl$masks$Fulvestrant, label1="Rep1", label2="Rep2", label3="Rep3", main="CTCF Peaks Fulvestrant")
dba.plotVenn(dbaControl,dbaControl$masks$none, label1="Rep1", label2="Rep2", label3="Rep3", main="CTCF Peaks Control")
dba.plotVenn(dbaExperiment,dbaExperiment$masks$Fulvestrant, label1="Rep1", label2="Rep2", label3="Rep3", main="ER Peaks Fulvestrant")
dba.plotVenn(dbaExperiment,dbaExperiment$masks$none,label1="Rep1", label2="Rep2", label3="Rep3", main="ER Peaks Control")
```
```{r}

filename<-"Rdata/026_SLX-8047_dba_drosophila.rda"
load(filename)
filename<-"Rdata/026_SLX-8047_dba_human.rda"
load(filename)

dba.plotVenn(dba_dm,dba_dm$masks$Fulvestrant, label1="Rep1", label2="Rep2", label3="Rep3",label4="Rep4", main="H2av Peaks Fulvestrant")
dba.plotVenn(dba_dm,dba_dm$masks$none,label1="Rep1", label2="Rep2", label3="Rep3",label4="Rep4", main="H2av Peaks Control")
dba.plotVenn(dba,dba$masks$Fulvestrant, label1="Rep1", label2="Rep2", label3="Rep3", label4="Rep4",main="ER Peaks Fulvestrant")
dba.plotVenn(dba,dba$masks$none,label1="Rep1", label2="Rep2", label3="Rep3",label4="Rep4", main="ER Peaks Control")
```

```{r}
filename<-"Rdata/014_SLX-12998_dba_mouse.rda"
load(filename)

dba.plotVenn(dba,dba$masks$Fulvestrant, label1="Rep1", label2="Rep2", label3="Rep3", label4="Rep4",main="Mouse ER Peaks Fulvestrant")
dba.plotVenn(dba,dba$masks$none,label1="Rep1", label2="Rep2", label3="Rep3",label4="Rep4", main="Mouse ER Peaks Control")

filename<-"Rdata/015_SLX-12998_dba_human.rda"
load(filename)

dba.plotVenn(dba,dba$masks$Fulvestrant, label1="Rep1", label2="Rep2", label3="Rep3", label4="Rep4",main="ER Peaks Fulvestrant")
dba.plotVenn(dba,dba$masks$none,label1="Rep1", label2="Rep2", label3="Rep3",label4="Rep4", main="ER Peaks Control")
```

```{r}

# Read consensus peaks
hsconsensus<-readRDS("Rdata/015_SLX-12998_hsconsensus.rds")
mmconsensus<-readRDS("Rdata/015_SLX-12998_mmconsensus.rds")

# Get counts
hscounts<-hsconsensus[,-c(1:3)]
mmcounts<-mmconsensus[,-c(1:3)]
head(hscounts)
head(mmcounts)

### MA RPM in peaks
hsrpm<-apply(hscounts,2,function(x){
  1E6*x/sum(x)
})
mmrpm<-apply(mmcounts,2,function(x){
  1E6*x/sum(x)
})
M<-apply(hsrpm,1,function(x){
  fulvestrant<-mean(x[c(1,2,4,5)])
  untreated<-mean(x[c(3,6,7,8)])
  fc<-mean(fulvestrant)/mean(untreated)
  log2fc<-log2(fc)
  return(log2fc)
})
A<-apply(hsrpm,1,function(x){
  return(log10(sum(x)))
})

load("Rdata/012_SLX-12998_aligned.rda")
aligned<-aligned[grep("SLX-12998.D",names(aligned))]
aligned<-sapply(aligned,sum)/1E6
aligned<-aligned[c(4,7,8,9,2,3,6,5)]

hsrpm<-hscounts
for(i in 1:length(hsrpm)){
  hsrpm[i]<-hscounts[i]/aligned[i]
}
mmrpm<-mmcounts
for(i in 1:length(mmrpm)){
  mmrpm[i]<-mmcounts[i]/aligned[i]
}
M<-apply(hsrpm,1,function(x){
  fulvestrant<-mean(x[c(1,2,4,5)])
  untreated<-mean(x[c(3,6,7,8)])
  fc<-mean(fulvestrant)/mean(untreated)
  log2fc<-log2(fc)
  return(log2fc)
})
A<-apply(hsrpm,1,function(x){
  return(log10(sum(x)))
})
plot(A,M,cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5,cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5,pch=20,ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"),main="RPM aligned reads",ylim=c(-6.25,4.25))
abline(h=0)

```

```{r}

### MA for Mouse + Human Counts
# a is treated, b is not treated
Mhs<-apply(hscounts,1,function(x){
  fulvestrant<-mean(x[c(1,2,4,5)])
  untreated<-mean(x[c(3,6,7,8)])
  fc<-mean(fulvestrant)/mean(untreated)
  log2fc<-log2(fc)
  return(log2fc)
})
Ahs<-apply(hscounts,1,function(x){
  return(log10(sum(x)))
})
Mmm<-apply(mmcounts,1,function(x){
  fulvestrant<-mean(x[c(1,2,4,5)])
  untreated<-mean(x[c(3,6,7,8)])
  fc<-mean(fulvestrant)/mean(untreated)
  log2fc<-log2(fc)
  return(log2fc)
})
Amm<-apply(mmcounts,1,function(x){
  return(log10(sum(x)))
})

plot(Ahs,Mhs,cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5,pch=20,ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"),main="RPM aligned reads",ylim=c(-6.25,4))
points(Amm,Mmm,pch=20,col="darkolivegreen3")
abline(h=0)
legend("topright",legend=c("Mouse","Human"),pch=20,col=c("darkolivegreen3","black"))
lm1<-lm(Mmm~Amm)
abline(lm1$coef,col="red")

```

```{r}

# Residuals of the Drosophila fit
lm1<-lm(Mmm~Amm)
intercept<-lm1$coef[1]
angularcoeff<-lm1$coef[2]
MhsFit<-Mhs-(Ahs*angularcoeff)-intercept
MmmFit<-Mmm-(Amm*angularcoeff)-intercept

plot(Ahs,MhsFit,cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5,pch=20,ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"),main="Counts normalized by Mouse Distribution",ylim=c(-16,2))
points(Amm,MmmFit,pch=20,col="darkolivegreen3")
abline(h=0)
legend("topright",legend=c("Mouse","Human"),pch=20,col=c("darkolivegreen3","black"))
lm1<-lm(MmmFit~Amm)
abline(lm1$coef,col="red")
```

```{r}
hsconsensus_ER<-readRDS("Rdata/018_SLX-14229_hsconsensus_ER.rds")
hsconsensus_CTCF<-readRDS("Rdata/018_SLX-14229_hsconsensus_CTCF.rds")
hscounts_ER<-hsconsensus_ER[,-c(1:3)]
hscounts_CTCF<-hsconsensus_CTCF[,-c(1:3)]

load("Rdata/012_SLX-14229_aligned.rda")
aligned<-aligned[grep("SLX-14229.D",names(aligned))]
aligned<-sapply(aligned,sum)/1E6

#remove inputs/controls to match samples here
aligned<-aligned[c(1,2,5,7,8,10)]

hsrpm_ER<-hscounts_ER
for(i in 1:length(hsrpm_ER)){
  hsrpm_ER[i]<-hscounts_ER[i]/aligned[i]
}
M_RPM_ER<-apply(hsrpm_ER,1,function(x){
  fulvestrant<-mean(x[c(2,4,5)])
  untreated<-mean(x[c(1,3,6)])
  fc<-mean(fulvestrant)/mean(untreated)
  log2fc<-log2(fc)
  return(log2fc)
})
A_RPM_ER<-apply(hsrpm_ER,1,function(x){
  return(log10(sum(x)))
})
plot(A_RPM_ER,M_RPM_ER,cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5,pch=20,ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"), main="RPM aligned reads")
abline(h=0)

```

```{r}

hsrpm_CTCF<-hscounts_CTCF
for(i in 1:length(hsrpm_CTCF)){
  hsrpm_CTCF[i]<-hscounts_CTCF[i]/aligned[i]
}
M_RPM_CTCF<-apply(hsrpm_CTCF,1,function(x){
  fulvestrant<-mean(x[c(2,4,5)])
  untreated<-mean(x[c(1,3,6)])
  fc<-mean(fulvestrant)/mean(untreated)
  log2fc<-log2(fc)
  return(log2fc)
})
A_RPM_CTCF<-apply(hsrpm_CTCF,1,function(x){
  return(log10(sum(x)))
})

plot(A_RPM_ER,M_RPM_ER,cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5,pch=20,ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"), main="RPM aligned reads")
points(A_RPM_CTCF,M_RPM_CTCF,pch=20,col="gray")
lm1<-lm(M_RPM_CTCF~A_RPM_CTCF)
abline(lm1$coef,col="blue")
abline(h=0)

```

```{r}

Mhs_CTCF<-apply(hscounts_CTCF,1,function(x){
  fulvestrant<-mean(x[c(2,4,5)])
  untreated<-mean(x[c(1,3,6)])
  fc<-mean(fulvestrant)/mean(untreated)
  log2fc<-log2(fc)
  return(log2fc)
})
Ahs_CTCF<-apply(hscounts_CTCF,1,function(x){
  return(log10(sum(x)))
})

Mhs_ER<-apply(hscounts_ER,1,function(x){
  fulvestrant<-mean(x[c(2,4,5)])
  untreated<-mean(x[c(1,3,6)])
  fc<-mean(fulvestrant)/mean(untreated)
  log2fc<-log2(fc)
  return(log2fc)
})
Ahs_ER<-apply(hscounts_ER,1,function(x){
  return(log10(sum(x)))
})

lm1<-lm(Mhs_CTCF~Ahs_CTCF)
intercept<-lm1$coef[1]
angularcoeff<-lm1$coef[2]
MhsFit_ER<-Mhs_ER-(Ahs_ER*angularcoeff)-intercept
MhsFit_CTCF<-Mhs_CTCF-(Ahs_CTCF*angularcoeff)-intercept
plot(Ahs_ER,MhsFit_ER,cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5,pch=20,ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"), main="Normalised to Human CTCF")
points(Ahs_CTCF,MhsFit_CTCF,pch=20,col="gray")

abline(h=0)
legend("topright",legend=c("Human ER",  "Human CTCF"),pch=20,col=c("black","gray"), cex=1)

```
```{r}
filename<-"Rdata/012_SLX-12998_nrreads.rda"
load(filename)

filename<-"Rdata/012_SLX-12998_aligned.rda"
load(filename)

nrreads<-nrreads[1:10]#Remove lost reads
aligned<-aligned[1:10]

### Prepare relative alignments
samples<-names(nrreads)
toplot<-matrix(nrow=length(samples),ncol=5)
colnames(toplot)<-c("MmTot","HsTot","Tot","MmRel","HsRel")
rownames(toplot)<-samples

mms<-sapply(aligned,function(x){
  sum(x[grep("mm_",names(x))])
})
hss<-sapply(aligned,function(x){
  sum(x[grep("hs_",names(x))])
})
toplot[,1]<-mms
toplot[,2]<-hss
toplot[,3]<-mms+hss
toplot[,4]<-mms/(mms+hss)
toplot[,5]<-hss/(mms+hss)

toplot<-toplot[grep('SLX-12998.D', sort(rownames(toplot)), value=TRUE),]

## Barplot of relative alignments
# Convert sample names to informative ones

rownames(toplot)<-c(
  "Input ICI",
  "2 ICI",
  "1 Control",
  "4 ICI",
  "2 Control",
  "3 Control",
  "1 ICI",
  "4 Control",
  "3 ICI",
  "Input Control"
)

# Sort by total reads aligned

toplot<-toplot[order(-toplot[,3]),]

par(las=2,mar=c(7,4,3,5))
bp<-barplot(t(toplot[,4:5]),beside=TRUE,main="Relative Read Alignment in samples",
            ylab="Fraction of total aligned",col=c("royalblue2","tomato"),ylim=c(-0.05,1.05))
grid()
legend("topright",legend=c("Mouse","Human"),col=c("royalblue2","tomato"),pch=15)
bpx<-apply(bp,2,mean)

raxis<-(toplot[,3])/(max(toplot[,3]))
lines(bpx,raxis,type="b",lwd=3,pch=16)
axis(4,at=seq(0,1,0.2),
     labels=round((seq(0,1,0.2) * max(toplot[,3])/10**6))
     )
par(las=3)
mtext("Reads aligned (Millions)",side=4,line=4)

```
```{r}
filename<-"Rdata/015_SLX-14229_dbacounts_human.rda"
load(filename)
plot(dbacounts)

filename<-"Rdata/018_SLX-14229_dba.counts_human_CTCF.rda"
load(filename)
plot(dba)

filename<-"Rdata/018_SLX-14229_dba.counts_human_ER.rda"
load(filename)
plot(dba)
```

```{r}
jg.controlSampleSheet     <- "samplesheet/samplesheet_SLX14438_hs_CTCF_DBA.csv"
jg.experimentSampleSheet  <- "samplesheet/samplesheet_SLX14438_hs_ER_DBA.csv"

filename<-"Rdata/040_SLX-14438_dba_human_ER_CTCF.rda"
load(filename)

jg.sampleIds <- jg.getSampleIds(jg.controlSampleSheet)

## Extract Peak set from DiffBind
jg.experimentPeakset <- jg.dbaGetPeakset(dbaExperiment)
jg.controlPeakset    <- jg.dbaGetPeakset(dbaControl)

#Get counts for each condition
jg.controlCountsTreated<-jg.getControlCounts(jg.controlPeakset, 
                                             jg.controlSampleSheet,
                                             jg.treatedCondition)
jg.controlCountsUntreated<-jg.getControlCounts(jg.controlPeakset,
                                               jg.controlSampleSheet,
                                               jg.untreatedCondition)

#Get sample names for conditions
jg.untreatedNames <- names(jg.controlCountsUntreated)
jg.treatedNames   <- names(jg.controlCountsTreated)

jg.coefficient<-jg.getNormalizationCoefficient(jg.controlCountsTreated,
                                               jg.controlCountsUntreated)
jg.correctionFactor<-jg.getCorrectionFactor(jg.experimentSampleSheet,
                                            jg.treatedNames,
                                            jg.untreatedNames)

#Apply coefficent and control factor
jg.experimentPeaksetNormalised<-jg.applyNormalisation(jg.experimentPeakset,
                                                      jg.coefficient,
                                                      jg.correctionFactor,
                                                      jg.treatedNames)
#Return values to Diffbind and plot normalised result.
jg.dba <- DiffBind:::pv.resetCounts(dbaExperiment,
                                    jg.experimentPeaksetNormalised)
jg.dba_analysis<-dba.analyze(jg.dba)

dba.plotMA(jg.dba_analysis,bFlip=TRUE, cex.lab=1.5, cex.axis=1.5, cex.main=1.25, cex.sub=1.5)
dba_analysis<-dba.analyze(dbaExperiment)
dba.plotMA(dba_analysis,bFlip=TRUE,cex.lab=1.5, cex.axis=1.5, cex.main=1.25, cex.sub=1.5)
#Note exact numbers change depending on the machine DiffBind is run.

```
```{r}
filename<-"Rdata/040_SLX-14438_dba_human_ER_CTCF.rda"
load(filename)

## Extract Peak set from DiffBind
jg.experimentPeakset <- jg.dbaGetPeakset(dbaExperiment)
jg.controlPeakset    <- jg.dbaGetPeakset(dbaControl)

#Convert Peakset to DeSeq Workflow
jg.controlPeaksetDeSeq<-jg.convertPeakset(jg.controlPeakset)

#Establish size factors directly from Control data
jg.controlSizeFactors = estimateSizeFactorsForMatrix(jg.controlPeaksetDeSeq)

#Get conditions dataframe for DeSeq
jg.conditions <- read.csv(file=jg.controlSampleSheet, header=TRUE, sep=",")['Condition']

#Run DeSeq on control
jg.controlDeSeq<-jg.runDeSeq(jg.controlPeaksetDeSeq, jg.conditions,jg.controlSizeFactors)
jg.controlResultsDeseq   = results(jg.controlDeSeq)

#Repeat for experimental conditions

#Convert experiment Peakset to DeSeq Workflow
jg.experimentPeaksetDeSeq<-jg.convertPeakset(jg.experimentPeakset)

#Run DeSeq on experiment
jg.experimentDeSeq<-jg.runDeSeq(jg.experimentPeaksetDeSeq, jg.conditions,jg.controlSizeFactors)
jg.experimentResultsDeseq   = results(jg.experimentDeSeq)
dbaExperimentCorrected<-jg.correctDBASizeFactors(dbaExperiment,
                                                 jg.controlSizeFactors
                                                 )

dbaExperimentAnalysisCorrected<-dba.analyze(dbaExperimentCorrected)
  dba.plotMA(dbaExperimentAnalysisCorrected,bFlip=TRUE,cex.lab=1.5, cex.axis=1.5, cex.main=1.25, cex.sub=1.5)

  #Again numbers similar but not exact due to DiffBind
```
```{r}
filename<-"Rdata/032_SLX-14229_dba_human_ER_CTCF_normalised.rda"
load(filename)

filename<-"Rdata/032_SLX-8047_dba_normalised.rda"
load(filename)

jg.dba_analysis_SLX8047<-dba.analyze(jg.dba_SLX8047)
jg.dba_analysis_SLX14229<-dba.analyze(jg.dba_SLX14229)

nearestPeak <- nearest(dba.report(jg.dba_analysis_SLX8047, th=1),dba.report(jg.dba_analysis_SLX14229, th=1))
SLX8047_fc<-dba.report(jg.dba_analysis_SLX8047,bFlip=TRUE, th=1)$Fold
SLX14229_fc<-dba.report(jg.dba_analysis_SLX14229,th=1)[nearestPeak]$Fold

smoothScatter(SLX8047_fc,SLX14229_fc,pch=20, col="#00000055",
              xlab="Fold change normalized using Drosophila/H2av spike-in",
              ylab="Fold change normalized using CTCF control peaks",
              main="Comparision of normalisation methods",
              cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5)
abline(h=0, col="grey",lty=2)
abline(v=0, col="grey",lty=2)
coef<-lm( SLX8047_fc ~ SLX14229_fc + 0 )
abline(coef, col="navyblue")
text(3.25,-0.25, expression( "gradient = 0.96, r = 0.557, p-value < 2.2 x" ~ 10^{-16}), cex=1.25)

```
```{r}

```

```{r}
filename<-"Rdata/047_SLX-14438_dba_human_ER_CTCF.rda"
load(filename)

#Load Sample Ids from control sample sheet.
jg.sampleIds <- jg.getSampleIds(jg.controlSampleSheet)

## Extract Peak set from DiffBind
jg.experimentPeakset <- jg.dbaGetPeakset(dbaExperiment)
jg.controlPeakset    <- jg.dbaGetPeakset(dbaControl)

#Get counts for each condition
jg.controlCountsTreated<-jg.getControlCounts(jg.controlPeakset, 
                                             jg.controlSampleSheet,
                                             jg.treatedCondition)
jg.controlCountsUntreated<-jg.getControlCounts(jg.controlPeakset,
                                               jg.controlSampleSheet,
                                               jg.untreatedCondition)

#Get sample names for conditions
jg.untreatedNames <- names(jg.controlCountsUntreated)
jg.treatedNames   <- names(jg.controlCountsTreated)

 ##Get Normalization Coefficient
jg.coefficient<-jg.getNormalizationCoefficient(jg.controlCountsTreated,
                                               jg.controlCountsUntreated)

#As we're going to normalise a second data set we don't need to use a DiffBind correction factor
jg.correctionFactor<-1

#Apply coefficent and control factor
jg.experimentPeaksetNormalised<-jg.applyNormalisation(jg.experimentPeakset,
                                                      jg.coefficient,
                                                      jg.correctionFactor,
                                                      jg.treatedNames
)

#Return values to Diffbind object
jg.dba <- DiffBind:::pv.resetCounts(dbaExperiment,
                                    jg.experimentPeaksetNormalised
)

#jg.dba is now normalised a list of ER peaks and their expected ratio

#####
#
#Rerun the normalisation code to get the ER ratio
#
######

jg.normalisedExperimentPeakset <- jg.dbaGetPeakset(jg.dba)

#Get counts for each condition
jg.ERCountsTreated<-jg.getControlCounts(jg.normalisedExperimentPeakset, 
                                             jg.experimentSampleSheet,
                                             jg.treatedCondition)
jg.ERCountsUntreated<-jg.getControlCounts(jg.normalisedExperimentPeakset,
                                          jg.experimentSampleSheet,
                                               jg.untreatedCondition)

#Get sample names for conditions
jg.untreatedNames <- names(jg.ERCountsTreated)
jg.treatedNames   <- names(jg.ERCountsUntreated)

##Establish normalised ER ratio as we know it isn't 1

jg.ERratio<-jg.getNormalizationCoefficient(jg.ERCountsTreated,
                                               jg.ERCountsUntreated)

########
#
# Apply to ER only data
#
########

jg.EROnlyExperimentSampleSheet <- "samplesheet/samplesheet_SLX8047_hs.csv"

filename<-"Rdata/047_SLX-SLX-8047_dba_human_ER.rda"
load(filename)

#Get peakset

jg.EROnlyExperimentPeakset <- jg.dbaGetPeakset(dbaERonlyExperiment)

#Get counts for each condition
jg.EROnlyCountsTreated<-jg.getControlCounts(jg.EROnlyExperimentPeakset, 
                                            jg.EROnlyExperimentSampleSheet,
                                        jg.treatedCondition)
jg.EROnlyCountsUntreated<-jg.getControlCounts(jg.EROnlyExperimentPeakset,
                                              jg.EROnlyExperimentSampleSheet,
                                          jg.untreatedCondition)

#Get sample names for conditions
jg.untreatedNames <- names(jg.EROnlyCountsTreated)
jg.treatedNames   <- names(jg.EROnlyCountsUntreated)

jg.EROnlyExperimentRatio<-jg.getNormalizationCoefficient(jg.EROnlyCountsTreated,
                                           jg.EROnlyCountsUntreated)

#Create new ratio 
jg.ERNormalisationRatio<-jg.ERratio/jg.EROnlyExperimentRatio

g.correctionFactor<-jg.getCorrectionFactor(jg.EROnlyExperimentSampleSheet,
                                            jg.treatedNames,
                                            jg.untreatedNames
)

jg.ERexperimentPeaksetNormalised<-jg.applyNormalisation(jg.EROnlyExperimentPeakset,
                                                      jg.ERNormalisationRatio,
                                                      jg.correctionFactor,
                                                      jg.treatedNames
)

#Return values to Diffbind object
dbaERonlyExperimentNormalised <- DiffBind:::pv.resetCounts(dbaERonlyExperiment,
                                    jg.ERexperimentPeaksetNormalised
)

jg.dba_analysisNormalised<-dba.analyze(dbaERonlyExperimentNormalised)
false_postive_normalised<-dba.report(jg.dba_analysisNormalised, th=0.05)
false_postive_normalised<-length(false_postive_normalised[false_postive_normalised$Fold>0])

jg.dba_analysis<-dba.analyze(dbaERonlyExperiment)

load(file="Rdata/027_SLX-8047_dba_report.Rda")
dba.SLX8047<-dba.analyze(dba.SLX8047)

peaks<-as.character(sort(as.numeric(row.names(as.data.frame(dba.report(dba.SLX8047, th=1))))))

plot(
    dba.report(dba.SLX8047, th=1)[peaks]$Fold,
    dba.report(jg.dba_analysisNormalised,th=1)[peaks]$Fold,
    pch=20,
    xlab="Log(FoldChange) in ER Binding Normalised to H2av",
    ylab="Log(FoldChange) in ER Binding Cross-Normalised to CTCF",
    cex=0.05,
    main="Comparision of Fold-Change between\nCross-Normalised to Xenogenic Spike-in"
)
```

### 2017-11-07 040_SLX-14438-MAplots_ER_CTCF.R

@@ -26,7 +26,6 @@ jg.plotNormalization <- function(jg.controlCountsTreated,jg.controlCountsUntreat
     legend("topleft",legend=c("Raw", "Normalised"),pch=20,col=c("black","royalblue3"))
 }
 
-
 ######################
 #
 # Settings
@@ -102,6 +101,11 @@ jg.dba <- DiffBind:::pv.resetCounts(dbaExperiment,
 
 #Analyze and plot with Diffbind      
 jg.dba_analysis<-dba.analyze(jg.dba)
+dba.SLX14438_report<-dba.report(jg.dba_analysis, th=1)
+filename<-"Rdata/040_SLX-14438_dba_human_ER_CTCF_normalised_report.rda"
+save(dba.SLX14438_report,file=filename)
+
+
 png("plots/040_SLX-14438_DiffBind_Analysis.png")
 par(mar=c(5.1,5.1,4.1,4.1))
 dba.plotMA(jg.dba_analysis,bFlip=TRUE, cex.lab=1.5, cex.axis=1.5, cex.main=1.25, cex.sub=1.5)

### 2017-11-07 051_SLX-14229_LeaveOneOut.R

@@ -1,4 +1,4 @@
-
+#Ignore, doesn't work as poorly designed concept.
 
 ######################
 #
@@ -54,7 +54,7 @@ jg.dba_analysis <- dba.analyze(jg.dba)
 dba.plotMA(jg.dba_analysis, bFlip = TRUE)
 #dev.off()
 
-system("git add . && git commit -m 'auto commit' && git push")
+
 
 filename <- "Rdata/051_dba_1.rda"
 if (!file.exists(filename)) {
@@ -84,7 +84,7 @@ if (!file.exists(filename)) {
     load(filename)
 }
 
-system("git add . && git commit -m 'auto commit' && git push")
+
 
 filename <- "Rdata/051_dba_3.rda"
 if (!file.exists(filename)) {
@@ -100,4 +100,4 @@ if (!file.exists(filename)) {
     load(filename)
 }
 
-system("git add . && git commit -m 'auto commit' && git push")
+

### 2017-11-07 055_EREmotif_bed.sh

@@ -1,0 +1,6 @@
+#Take the motif sites we got from a genome wide search and expan them +500bp both directions.
+
+cd bed
+wget https://genome.ucsc.edu/goldenpath/help/hg19.chrom.sizes --no-check-certificate
+
+bedtools slop -i ../motifAnalysis/bed/output.bed -b 500 -g hg19.chrom.sizes > CTCFblacklist.bed

### 2017-11-07 056_CTCF_removeERE.sh

@@ -1,0 +1,2 @@
+cd SLX-14438_merged/peaks
+bedtools subtract -A -a CTCF_union.bed -b ../../motifAnalysis/bed/output.bed > CTCF_EREremoved.bed

### 2017-11-07 057_ER_conensus_peaks.R

@@ -1,0 +1,21 @@
+
+library(DiffBind)
+
+setwd("/Volumes/FlyPeaks/FlyPeaks")
+
+if(!file.exists("Rdata/057_dba.rda")) {
+    dba<-dba(sampleSheet="samplesheet/samplesheet_SLX12998_hs.csv")
+    dba<-dba.count(dba,  minOverlap=3, bParallel=FALSE)
+    dbaPeakset<-dba.peakset(dba,  bRetrieve = T, DataType = DBA_DATA_FRAME)
+    save(dba, file="Rdata/057_dba.rda")
+} else {
+    load(file="Rdata/057_dba.rda")
+}
+
+write.table(dbaPeakset[,1:3],"bed/ERConsensus.bed", quote=FALSE, sep="\t",
+            row.names=FALSE, col.names=FALSE)
+
+setwd("./bed")
+system("bedtools slop -i ERConsensus.bed -b 500 -g hg19.chrom.sizes > ERConsensus500.bed")
+setwd("../SLX-14438_merged/peaks")
+system("bedtools subtract -A -a CTCF_union.bed -b ../../bed/ERConsensus500.bed > CTCF_ERConsensusremoved.bed")

### 2017-11-07 058_SLX-14438_ER_consensus_CTCF_split.sh

@@ -1,0 +1,9 @@
+cd SLX-14438_merged/peaks/
+mkdir CTCF_noER
+
+for f in *.narrowPeak
+do
+	echo $f
+	bedtools intersect -a $f -b CTCF_ERConsensusremoved.bed  > CTCF_noER/$f 
+done
+

### 2017-11-07 059_SLX14438_noER.R

@@ -1,0 +1,80 @@
+setwd("/Volumes/FlyPeaks/FlyPeaks")
+source('package/brundle.R')
+jg.controlSampleSheet     <- "samplesheet/samplesheet_SLX14438_hs_CTCF_DBA_noER.csv"
+jg.experimentSampleSheet  <- "samplesheet/samplesheet_SLX14438_hs_ER_DBA.csv"
+jg.treatedCondition       =  "Fulvestrant"
+jg.untreatedCondition     =  "none"
+dbaSummits                <- 200
+
+filename<-"Rdata/059_SLX-14438_dba_human_ER_CTCF_noER.rda"
+if(!file.exists(filename)){
+    dbaExperiment <- jg.getDba(jg.experimentSampleSheet,dbaSummits)
+    dbaControl    <- jg.getDba(jg.controlSampleSheet,   dbaSummits)
+    save(dbaExperiment,dbaControl,file=filename)
+} else {
+    load(filename)
+}
+
+
+#Load Sample Ids from control sample sheet.
+jg.sampleIds <- jg.getSampleIds(jg.controlSampleSheet)
+
+## Extract Peak set from DiffBind
+jg.experimentPeakset <- jg.dbaGetPeakset(dbaExperiment)
+jg.controlPeakset    <- jg.dbaGetPeakset(dbaControl)
+
+
+#Get counts for each condition
+jg.controlCountsTreated<-jg.getControlCounts(jg.controlPeakset, 
+                                             jg.controlSampleSheet,
+                                             jg.treatedCondition)
+jg.controlCountsUntreated<-jg.getControlCounts(jg.controlPeakset,
+                                               jg.controlSampleSheet,
+                                               jg.untreatedCondition)
+
+#Get sample names for conditions
+jg.untreatedNames <- names(jg.controlCountsUntreated)
+jg.treatedNames   <- names(jg.controlCountsTreated)
+
+##Get Normalization Coefficient
+jg.coefficient<-jg.getNormalizationCoefficient(jg.controlCountsTreated,
+                                               jg.controlCountsUntreated)
+
+
+#Check by MA plot (Optional)
+jg.plotMA(jg.experimentPeakset,jg.controlPeakset,jg.untreatedNames,jg.treatedNames,jg.coefficient)
+
+#DeSeq called by Diffbind will divide though by library size to normalise, 
+#and therefore partially undo our work. Solution is to correct our normalisation
+#factor to with this correction factor to remove library size part of our 
+#NormalisationFactor.
+
[…]

### 2017-11-08 059_SLX14438_noER.R
ER sites removed from CTCF control peaks

@@ -6,15 +6,35 @@ jg.treatedCondition       =  "Fulvestrant"
 jg.untreatedCondition     =  "none"
 dbaSummits                <- 200
 
-filename<-"Rdata/059_SLX-14438_dba_human_ER_CTCF_noER.rda"
-if(!file.exists(filename)){
-    dbaExperiment <- jg.getDba(jg.experimentSampleSheet,dbaSummits)
-    dbaControl    <- jg.getDba(jg.controlSampleSheet,   dbaSummits)
-    save(dbaExperiment,dbaControl,file=filename)
-} else {
-    load(filename)
+#Fix as my computer is crashing on bParallel at the moment.
+jg.getDba.linear<-function (jg.experimentSampleSheet,dbaSummits, ...)
+{
+    
+    dba <- dba(sampleSheet = jg.experimentSampleSheet)
+    if(exists("dbaSummits"))
+    {
+        dba <- dba.count(dba, summits=dbaSummits, ...)
+    } else {
+        dba <- dba.count(dba)
+    }
+    dba <- dba.count(dba, peaks=NULL, score=DBA_SCORE_READS)
+    return(dba)
 }
 
+#Load data used in the paper.
+filename<-"Rdata/040_SLX-14438_dba_human_ER_CTCF.rda"
+load(filename)
+
+#filename<-"Rdata/059_SLX-14438_dba_human_ER_CTCF_noER.rda"
+#if(!file.exists(filename)){
+#    dbaExperiment <- jg.getDba(jg.experimentSampleSheet,dbaSummits)
+#    dbaControl    <- jg.getDba(jg.controlSampleSheet,   dbaSummits, minOverlap=5)
+#    save(dbaExperiment,dbaControl,file=filename)
+##} else {
+#    load(filename)
+#}
+
+#dbaControl    <- jg.getDba.linear(jg.controlSampleSheet,  dbaSummits, minOverlap=5)
 
 #Load Sample Ids from control sample sheet.
 jg.sampleIds <- jg.getSampleIds(jg.controlSampleSheet)
@@ -43,6 +63,7 @@ jg.coefficient<-jg.getNormalizationCoefficient(jg.controlCountsTreated,
 
 #Check by MA plot (Optional)
 jg.plotMA(jg.experimentPeakset,jg.controlPeakset,jg.untreatedNames,jg.treatedNames,jg.coefficient)
[…]

### 2017-11-09 059_SLX14438_noER.R
upadted figures

@@ -176,8 +176,9 @@ jg.dba <- DiffBind:::pv.resetCounts(dbaExperiment,
 
 #Analyze and plot with Diffbind      
 jg.dba_analysis<-dba.analyze(jg.dba)
-png("plots/059_CTCF_noER_MAplot.png",pointsize=15)
-dba.plotMA(jg.dba_analysis,bFlip=TRUE)
+png("plots/059_CTCF_noER_MAplot.png")
+par(mar=c(5.1,5.1,4.1,4.1))
+dba.plotMA(jg.dba_analysis,bFlip=TRUE,cex.lab=1.5, cex.axis=1.5, cex.main=1.25, cex.sub=1.5)
 dev.off()
 #Check CTCF 
 ctcf.dba.analysis<-dba.analyze(dbaControl_noER)
@@ -195,7 +196,27 @@ jg.conditions <- read.csv(file=jg.controlSampleSheet, header=TRUE, sep=",")['Con
 jg.controlDeSeq<-jg.runDeSeq(jg.controlPeaksetDeSeq, jg.conditions,jg.controlSizeFactors)
 jg.controlResultsDeseq   = results(jg.controlDeSeq)
 
-png("plots/059_CTCF_noER_control_MAplot.png",pointsize=15)
+
+
+jg.plotDeSeq<-function(ma.df, filename = 'file.name', p = 0.01, title.main = "Differential ChIP",log2fold =0.5, flip=FALSE)
+{;
+    
+    if (flip == TRUE)
+    {
+        ma.df$log2FoldChange <- -ma.df$log2FoldChange
+    }
+    par(mar=c(5.1,5.1,4.1,2.1))
+    xyplot(ma.df$log2FoldChange ~ log(ma.df$baseMean, base=10),#xlim=c(0,4),ylim=c(-3,1.25),
+           groups=(ma.df$padj < p & abs(ma.df$log2FoldChange) > log2fold & !is.na(ma.df$padj)),
+           col=c("black","red"), main=title.main, scales="free", aspect=1, pch=20, cex=0.5,
+           ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"),
+           par.settings=list(axis.text=list(cex=1.5,font=1),par.main.text=list(cex=1.5,font=2),par.xlab.text=list(cex=1.5,font=2), par.ylab.text=list(cex=1.5,font=2)
+           )
+    );
+    
+}
+
[…]

### 2017-11-14 Manuscript_Figures.Rmd
upadted figures
Rmarkdown added

@@
 date: "11/7/2017"
-output: html_document
+output:
+  html_document:
+    toc: true
+    toc_depth: 2
+theme: united
 ---
@@
 
+#Main
+
 ## Comparison of simple normalisation methods
 
-```{r}
+```{r results='hide', message=FALSE, warning=FALSE}
 library(DiffBind)
-#setwd("/Volumes/FlyPeaks/FlyPeaks")
+
+```
+
+```{r, fig.cap="Comparison of simple normalisation strategies employed.
+MA plots showing the changes in ER binding after 48 hours treatment with 100 nM fulvestrant.
+Three simple normalisation methods were applied to this data and compared to the raw count data.
+(A) Raw counts.
+(B) Reads Per Million (RPM) reads in peaks.
+(C) RPM aligned reads.
+(D) RPM total reads.
+Note that the highlighted peaks remain above zero under all three standard normalisations."}
+
+par(mfrow=c(2,2))
 
@@
 
-```
-
-```{r}
-
 load("Rdata/012_SLX-12998_aligned.rda")
@@
 abline(h=0)
-
-```
-
-```{r}
 
@@
 
-```
-
-```{r}
 ### MA Counts human
@@
 plot(A,M,cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5,pch=20,ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"),main="Raw counts in peaks", ylim=c(-6.25,4.25))
+
 abline(h=0)
@@
 
-## MA Plot of H2AV normalization
-```{r}
+## MA Plot of H2Av normalization
+```{r, fig.cap="MA plots showing ER binding before and after treatment with fulvestrant including matched Dm H2Av spike-in control.} (A) Reads corrected to total aligned reads showed the same off-centre peak density as observed in Figure 1.
+Putative unchanged ER binding sites are within the red triangle.
+(B) Overlaying the MA plot combining the changes in chromatin binding of Hs ER (black) and Dm H2Av (blue).
+Dm peaks overlay the off-centre peak density.
+(C) Utilising the Dm H2Av binding events as a ground truth for 0-fold change, a linear fit to the log-fold change is generated and the fit is applied to adjust the Hs ER binding events."}
+par(mfrow=c(1,3))
 # Read consensus peaks
@@
 plot(Ahs,Mhs,cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5,pch=20,ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"),main="RPM aligned reads",ylim=c(-6.25,2))
+abline(h=0)
+
+plot(Ahs,Mhs,cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5,pch=20,ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"),main="RPM aligned reads",ylim=c(-6.25,2))
 points(Adm,Mdm,col="cornflowerblue",pch=20)
@@
 abline(lm1$coef,col="red")
-```
-
-```{r}
+
 ### 20th century normalization
@@
 
-## DeSeq Results
-
-```{r}
+## RARA gene locus with CTCF Spike-in
+
+Figure 3 can be viewed interactively on the [USCS track](http://genome.ucsc.edu/cgi-bin/hgTracks?hgS_doOtherUser=submit&hgS_otherUserName=andrewholding&hgS_otherUserSessionName=Brundle).
+
+## MA Plot of CTCF Peaks
+
+```{r, results='hide', message=FALSE, warning=FALSE, fig.cap="Comparison of the control regions used to normalise ER analysis before and after treatment.
+Dots highlighted in red are significant (FDR = 0.01).
+The CTCF peaks used for normalisation show no significant change in the number reads before and after treatment."}
 
 source('package/brundle.R')
-## Changing font sizes for manuscript
+filename<-"Rdata/040_SLX-14438_dba_human_ER_CTCF.rda"
+load(filename)
+jg.experimentPeakset <- jg.dbaGetPeakset(dbaExperiment)
+
+filename<-"Rdata/059_SLX-14438_dba_human_ER_CTCF_noER.rda"
+load(filename)
+jg.controlPeakset    <- jg.dbaGetPeakset(dbaControl_noER)
+
+#Get counts for each condition
+jg.controlSampleSheet     <- "samplesheet/samplesheet_SLX14438_hs_CTCF_DBA_noER.csv"
+jg.experimentSampleSheet  <- "samplesheet/samplesheet_SLX14438_hs_ER_DBA.csv"
+jg.treatedCondition       =  "Fulvestrant"
+jg.untreatedCondition     =  "none"
+
+jg.controlCountsTreated<-jg.getControlCounts(jg.controlPeakset, 
+                                             jg.controlSampleSheet,
+                                             jg.treatedCondition)
+jg.controlCountsUntreated<-jg.getControlCounts(jg.controlPeakset,
+                                               jg.controlSampleSheet,
+                                               jg.untreatedCondition)
+
+#Get sample names for conditions
+jg.untreatedNames <- names(jg.controlCountsUntreated)
+jg.treatedNames   <- names(jg.controlCountsTreated)
+
+##Get Normalization Coefficient
+jg.coefficient<-jg.getNormalizationCoefficient(jg.controlCountsTreated,
+                                               jg.controlCountsUntreated)
+
+#DeSeq called by Diffbind will divide though by library size to normalise, 
+#and therefore partially undo our work. Solution is to correct our normalisation
+#factor to with this correction factor to remove library size part of our 
+#NormalisationFactor.
+
+jg.correctionFactor<-jg.getCorrectionFactor(jg.experimentSampleSheet,
+                                            jg.treatedNames,
+                                            jg.untreatedNames
+)
+
+#Apply coefficent and control factor
+jg.experimentPeaksetNormalised<-jg.applyNormalisation(jg.experimentPeakset,
+                                                      jg.coefficient,
+                                                      jg.correctionFactor,
+                                                      jg.treatedNames
+)
+
+#Return values to Diffbind and plot normalised result.
+jg.dba <- DiffBind:::pv.resetCounts(dbaExperiment,
+                                    jg.experimentPeaksetNormalised
+)
+
+#Analyze and plot with Diffbind      
+jg.dba_analysis<-dba.analyze(jg.dba)
+par(mar=c(5.1,5.1,4.1,4.1))
+#Store for later
+jg.dba_analysis_noER<-jg.dba_analysis
+#dba.plotMA(jg.dba_analysis,bFlip=TRUE,cex.lab=1.5, cex.axis=1.5, cex.main=1.25, cex.sub=1.5)
+#Check CTCF 
+ctcf.dba.analysis<-dba.analyze(dbaControl_noER)
+
+#dba.plotMA(ctcf.dba.analysis,bFlip=TRUE)
+
+#Run on DeSeq
+jg.controlPeakset    <- jg.dbaGetPeakset(dba.count(dbaControl_noER, peaks=NULL, score=DBA_SCORE_READS))
+jg.controlPeaksetDeSeq<-jg.convertPeakset(jg.controlPeakset )
+jg.controlSizeFactors = estimateSizeFactorsForMatrix(jg.controlPeaksetDeSeq)
+
+jg.conditions <- read.csv(file=jg.controlSampleSheet, header=TRUE, sep=",")['Condition']
+
+jg.controlDeSeq<-jg.runDeSeq(jg.controlPeaksetDeSeq, jg.conditions,jg.controlSizeFactors)
+jg.controlResultsDeseq   = results(jg.controlDeSeq)
 
@@
 
-jg.plotDeSeqCombined <- function(jg.controlResultsDeseq,jg.experimentResultsDeseq,title.main,padjX,flip=FALSE)
-{
-    jg.controlResultsDeseq$group = 'a'
-    jg.experimentResultsDeseq$group = 'b'
-
-    if (flip == TRUE)
-    {
-        jg.controlResultsDeseq$log2FoldChange <- -jg.controlResultsDeseq$log2FoldChange
-        jg.experimentResultsDeseq$log2FoldChange <- -jg.experimentResultsDeseq$log2FoldChange
-    }
-
-    for (i in 1:length(jg.experimentResultsDeseq$group)) {
-        if (!is.na(jg.experimentResultsDeseq$padj[i]) & !is.na(jg.experimentResultsDeseq$log2FoldChange[i]) & jg.experimentResultsDeseq$padj[i] < padjX & jg.experimentResultsDeseq$log2FoldChange[i] < 0) {
-            jg.experimentResultsDeseq$group[i] <- 'd'
-        }
-        else if (!is.na(jg.experimentResultsDeseq$padj[i]) & !is.na(jg.experimentResultsDeseq$log2FoldChange[i]) & jg.experimentResultsDeseq$padj[i] < padjX & jg.experimentResultsDeseq$log2FoldChange[i] > 0) {
-            jg.experimentResultsDeseq$group[i] <- 'c'
-        }
-    }
-
-    for (i in 1:length(jg.controlResultsDeseq$group)) {
-        if (!is.na(jg.controlResultsDeseq$padj[i]) & !is.na(jg.controlResultsDeseq$log2FoldChange[i]) & jg.controlResultsDeseq$padj[i] < padjX & jg.controlResultsDeseq$log2FoldChange[i] < 0) {
-            jg.controlResultsDeseq$group[i] <- 'f'
-        }
-        else if (!is.na(jg.controlResultsDeseq$padj[i]) & !is.na(jg.controlResultsDeseq$log2FoldChange[i]) & jg.controlResultsDeseq$padj[i] < padjX & jg.controlResultsDeseq$log2FoldChange[i] > 0) {
-            jg.controlResultsDeseq$group[i] <- 'e'
-        }
-    }
-
-    full.res = rbind(jg.controlResultsDeseq, jg.experimentResultsDeseq)
-    par(mar=c(5.1,5.1,4.1,2.1))
-    xyplot(full.res$log2FoldChange ~ log(full.res$baseMean, base=10), data = full.res,
-           groups=full.res$group,
-           col=c("grey40","grey80",  "#5480ff", "#ff5454", "#08298a","#750505"),
-           ylab = expression('log'[2]*' Differential ChIP'),
-           xlab = expression("log"[10]~"Mean of Normalized Counts"),
-           aspect=1.0,
-           pch=16,
-           cex=0.5,
-           main=title.main,
-           scales=list(x=list(cex=1.5, relation = "free"), y =list(cex=1.5, relation="free")),
-           between=list(y=0.5, x=0.5),
-           auto.key = TRUE,
-           par.settings=list(axis.text=list(cex=1.5,font=1),par.main.text=list(cex=1.5,font=2),par.xlab.text=list(cex=1.5,font=2), par.ylab.text=list(cex=1.5,font=2)),
-           key=list(corner=c(1,0),
-                    cex=1.0,
-                    points=list(col=c( "gray80","gray40", "#ff5454", "#5480ff", "#750505", "#08298a","white"), pch=20),
-                    text=list(c("Target Binding","Control Binding", "Target Decreased","Target Increased","Control Decreased","Control Increased", " "))
-           ))
-
-}
-
-######################
-#
-# Settings
-#
-######################
-
-setwd("/Volumes/FlyPeaks/FlyPeaks")
-dbaSummits                <- 200
-jg.controlMinOverlap      <- 5
-jg.controlSampleSheet     <- "samplesheet/samplesheet_SLX14229_hs_CTCF_DBA.csv"
-jg.experimentSampleSheet  <- "samplesheet/samplesheet_SLX14229_hs_ER_DBA.csv"
-
-filename<-"Rdata/028_SLX-14229_dba_human_ER_CTCF.rda"
-load(filename)
-
-## Extract Peak set from DiffBind
-jg.experimentPeakset <- jg.dbaGetPeakset(dbaExperiment)
-jg.controlPeakset    <- jg.dbaGetPeakset(dbaControl)
-
-#Convert Peakset to DeSeq Workflow
-jg.controlPeaksetDeSeq<-jg.convertPeakset(jg.controlPeakset)
-
-#Establish size factors directly from Control data
-jg.controlSizeFactors = estimateSizeFactorsForMatrix(jg.controlPeaksetDeSeq)
-
-#Get conditions dataframe for DeSeq
-jg.conditions <- read.csv(file=jg.controlSampleSheet, header=TRUE, sep=",")['Condition']
-
-#Run DeSeq on control
-jg.controlDeSeq<-jg.runDeSeq(jg.controlPeaksetDeSeq, jg.conditions,jg.controlSizeFactors)
-jg.controlResultsDeseq   = results(jg.controlDeSeq)
-
-#Repeat for experimental conditions
-
-#Convert experiment Peakset to DeSeq Workflow
-jg.experimentPeaksetDeSeq<-jg.convertPeakset(jg.experimentPeakset)
-
-#Run DeSeq on experiment
-jg.experimentDeSeq<-jg.runDeSeq(jg.experimentPeaksetDeSeq, jg.conditions,jg.controlSizeFactors)
-jg.experimentResultsDeseq   = results(jg.experimentDeSeq)
-
 jg.plotDeSeq(jg.controlResultsDeseq,
@@
 
-```{r}
-jg.plotDeSeq(jg.experimentResultsDeseq,
-             p=0.01,
-             title.main="Fold-change in ER binding",
-             flip=T
-             )
-```
-
-#CTCF Binding
-
-```{r}
-
-jg.experimentDeSeqInternal<-jg.runDeSeq(jg.experimentPeaksetDeSeq, jg.conditions, NULL)
-jg.experimentResultsDeseqInternal   = results(jg.experimentDeSeqInternal)
-
- jg.plotDeSeq(jg.experimentResultsDeseqInternal,
-               title.main="Fold-change in ER binding (no correction)",
-               p=0.01,
-               flip=T
-               )
-```
-
-#H2av With Diff Bind
-```{r}
+##H2av With DiffBind
+
+```{r results='hide', message=FALSE, warning=FALSE,   out.width = '49%', out.extra='style="float:left"'}
 jg.controlSampleSheet     <- "samplesheet/samplesheet_SLX8047_dm.csv"
@@
 
-```{r}
+```{r results='hide', message=FALSE, warning=FALSE, out.width = '49%', out.extra='style="float:left"'}
   dbaExperimentAnalysisCorrected<-dba.analyze(dbaExperimentCorrected)
@@
 ```
-```{r}
-
-filename<-"Rdata/040_SLX-14438_dba_human_ER_CTCF.rda"
-  load(filename)
-  jg.controlSampleSheet     <- "samplesheet/samplesheet_SLX14438_hs_CTCF_DBA.csv"
-jg.experimentSampleSheet  <- "samplesheet/samplesheet_SLX14438_hs_ER_DBA.csv"
-jg.treatedCondition       =  "Fulvestrant"
-jg.untreatedCondition     =  "none"
-
-#Get counts for each condition
-jg.controlCountsTreated<-jg.getControlCounts(jg.controlPeakset, 
-                                             jg.controlSampleSheet,
-                                             jg.treatedCondition)
-jg.controlCountsUntreated<-jg.getControlCounts(jg.controlPeakset,
-                                               jg.controlSampleSheet,
-                                               jg.untreatedCondition)
-
-#Get sample names for conditions
-jg.untreatedNames <- names(jg.controlCountsUntreated)
-jg.treatedNames   <- names(jg.controlCountsTreated)
-
-jg.plotNormalization(jg.controlCountsTreated, jg.controlCountsUntreated)
-```
-```{r}
-
+
+Comparison of DiffBind output before and after applying the corrected size factors from our pipeline generated from Drosophila spike-in control.
+(A) Analysis of ER binding before and after treatment with fulvestrant demonstrates that DiffBind’s default normalisation strategy is more effective than the DESeq2 default, but demonstrates a bias between samples.
+(B) Applying the correct size factors from our DESeq2 pipeline reduces the bias in the analysis (Data: SLX-8047).
+
+##Linear model
+
+```{r results='hide', message='false', warning='false', fig.cap="Comparison of mean counts in CTCF peaks before and after treatment.
+If the samples have no systematic bias before and after treatment then the linear fit would be expected to have a gradient of 1.
+Here, we establish that the gradient is < 1, implying a systematic bias between samples.
+The read counts in the treated samples peaks are corrected (blue), removing the bias, and resulting in a new gradient of 1."}
+
+#Remove print line from function.
+
+jg.plotNormalization<-function(jg.controlCountsTreated,jg.controlCountsUntreated)
+{
+  plot(rowMeans(jg.controlCountsTreated),rowMeans(jg.controlCountsUntreated), pch=20,
+       xlab="Counts in peak after treatment" ,  ylab="Counts in peak before treatment" ,
+       main="Comparision of Counts in peaks")
+  lm1<-lm(rowMeans(jg.controlCountsUntreated) ~ 0 + rowMeans(jg.controlCountsTreated))
+
+  abline(c(0,lm1$coef),col="red3")
+  angularcoeff<-lm1$coef[1]
+
+  points(rowMeans(jg.controlCountsTreated)*angularcoeff,rowMeans(jg.controlCountsUntreated),pch=20, col="royalblue3" )
+  treatment_fit<-rowMeans(jg.controlCountsTreated)*angularcoeff
+  lm1<-lm(treatment_fit ~ 0 + rowMeans(jg.controlCountsUntreated))
+  abline(c(0,lm1$coef),col="purple")
+  legend("topleft",legend=c("Raw", "Normalised"),pch=20,col=c("black","royalblue3"))
+}
+
+jg.plotNormalization(jg.controlCountsTreated,jg.controlCountsUntreated)
+```
+
+##Comparison of CTCF and H2Av normalisation methods
+
+```{r results='hide', echo=FALSE,warning=FALSE,message=FALSE,error=FALSE} 
 jg.controlSampleSheet     <- "samplesheet/samplesheet_SLX8047_dm.csv"
@@
 
-#Convert Peakset to DeSeq Workflow
+#Convert Peakset to DESeq Workflow
 jg.controlPeaksetDeSeq<-jg.convertPeakset(jg.controlPeakset)
@@
 
+```
+
+```{r out.width = '50%', out.extra='style="float:left"'}
 a + as.layer(b) 
-
-```
-```{r}
+```
+
+```{r out.width = '50%', out.extra='style="float:left"', fig.cap="Comparison of normalisation methods using consensus peak set.
+(A) The analysis for the CTCF normalised (blue) and H2Av normalised (green) dataset using an ER consensus peak set of 10,000 peaks were formatted as an MA plot and overlaid.
+This recovered the low-fold change higher-intensity peaks that were not visible in Figure \ref{fig:ERCTCF}A and both datasets showed a similar distribution.
+(B) Comparison of fold-change values for individual ER binding sites between two datasets showed that the inclusion of these sites did not appear to affect the correlation (r = 0.77)."}
 
@@
 
-```{r}
+#Supplementry
+
+##Method comparision
+
+```{r, results='hide', message=FALSE, warning=FALSE}
+
+jg.controlMinOverlap      <- 5
+jg.controlSampleSheet     <- "samplesheet/samplesheet_SLX14438_hs_CTCF_DBA.csv"
+jg.experimentSampleSheet  <- "samplesheet/samplesheet_SLX14438_hs_ER_DBA.csv"
+jg.treatedCondition       =  "Fulvestrant"
+jg.untreatedCondition     =  "none"
+
+######################
+#
+# Main Code
+#
+######################
+
+filename<-"Rdata/048_SLX-14438_dba_human_ER_CTCF.rda"
+load(filename)
+
+#Load Sample Ids from control sample sheet.
+jg.sampleIds <- jg.getSampleIds(jg.controlSampleSheet)
+
+## Extract Peak set from DiffBind
+jg.experimentPeakset <- jg.dbaGetPeakset(dbaExperiment)
+jg.controlPeakset    <- jg.dbaGetPeakset(dbaControl)
+
+#Get counts for each condition
+jg.controlCountsTreated<-jg.getControlCounts(jg.controlPeakset, 
+                                             jg.controlSampleSheet,
+                                             jg.treatedCondition)
+jg.controlCountsUntreated<-jg.getControlCounts(jg.controlPeakset,
+                                               jg.controlSampleSheet,
+                                               jg.untreatedCondition)
+
+#Get sample names for conditions
+jg.untreatedNames <- names(jg.controlCountsUntreated)
+jg.treatedNames   <- names(jg.controlCountsTreated)
+
+jg.coefficient<-jg.getNormalizationCoefficient(jg.controlCountsTreated,
+                                               jg.controlCountsUntreated)
+jg.correctionFactor<-jg.getCorrectionFactor(jg.experimentSampleSheet,
+                                            jg.treatedNames,
+                                            jg.untreatedNames)
+
+#Apply coefficent and control factor
+jg.experimentPeaksetNormalised<-jg.applyNormalisation(jg.experimentPeakset,
+                                                      jg.coefficient,
+                                                      jg.correctionFactor,
+                                                      jg.treatedNames)
+#Return values to Diffbind and plot normalised result.
+jg.dba <- DiffBind:::pv.resetCounts(dbaExperiment,
+                                    jg.experimentPeaksetNormalised)
+
+#Analyze and plot with Diffbind      
+jg.dba_analysis<-dba.analyze(jg.dba)
+
+#Analyze and plot with Diffbind no correction     
+dba_analysis_DeSEQ2<-dba.analyze(dbaExperiment,method=DBA_DESEQ2, bFullLibrarySize=FALSE)
+dba_analysis_EDGER<-dba.analyze(dbaExperiment,method=DBA_EDGER)
+dba_analysis_DiffBind<-dba.analyze(dbaExperiment,method=DBA_DESEQ2, bFullLibrarySize=TRUE)
+#
+
+plotForComparision<-function(dba, MethodName, dbatype,ylimits) {
+    df<-as.data.frame(dba.report(dba, method=dbatype, th=1))    
+    countSet<-df[as.character(sort(as.numeric(rownames(df)))),]
+    peakSet<-jg.dbaGetPeakset(dba)   
+    c1<-rowMeans(peakSet[c(4,6,9)])
+    c0<-rowMeans(peakSet[c(5,7,8)])
+    c1<-c1[]
+
+M<-log(c1*c0)
+A<- -log(c1/c0,base=2)
+
+A<- -countSet$Fold
+#c0<-countSet$Conc_Fulvestrant
+#c1<-c1[]
+
+#png(paste0("plots/048_",MethodName,".png"), pointsize=20)
+par(mar=c(5.1,5.1,4.1,2.1))
+plot(M[countSet$FDR>0.05],A[countSet$FDR>0.05],pch=20,
+     main=MethodName,
+     ylab = expression('log'[2]*' Differential ChIP'),
+     xlab = expression("log"[10]~"Mean of Normalized Counts"),cex=0.05,
+     xlim=c(0,15),ylim=ylimits
+     )
+points(M[countSet$FDR<0.05],A[countSet$FDR<0.05],pch=20,col="red",cex=0.05)
+abline(h=0)
+#dev.off()
+}
+
+par(mfrow=c(2,2))
+filename="Rdata/047_countset.Rda"
+load(filename)
+c1<-rowMeans(countSet$db[c(4,6,9)])
+c0<-rowMeans(countSet$db[c(5,7,8)])
+c1<-c1[]
+
+M<-log(c1*c0)
+A<- -log(c1/c0,base=2)
+
+par(mar=c(5.1,5.1,4.1,2.1))
+plot(M[countSet$db$pvalue.wald>0.05],A[countSet$db$pvalue.wald>0.05],pch=20,
+        main="ChIPComp",
+        ylab = expression('log'[2]*' Differential ChIP'),
+        xlab = expression("log"[10]~"Mean of Normalized Counts"),cex=0.05)
+points(M[countSet$db$pvalue.wald<0.05],A[countSet$db$pvalue.wald<0.05],pch=20,col="red",cex=0.05)
+abline(h=0)
+
+plotForComparision(dba_analysis_EDGER, "EdgeR", DBA_EDGER,c(-3,3))
+plotForComparision(dba_analysis_DeSEQ2, "DeSEQ2",DBA_DESEQ2,c(-3,3))
+plotForComparision(dba_analysis_DiffBind, "DiffBind",DBA_DESEQ2,c(-6,2))
+#plotForComparision(jg.dba_analysis, "Brundle",DBA_DESEQ2,c(-6,2))
+
+```
+
+Comparison of ChIP-seq Pipelines.} (A)ChIPComp data was plot from the CountSet object, results show a high number of false positive up-regulated sites.
+(B) EdgeR normalisation is designed for the analysis of transcriptional data.
+In case of large-scale uni-direction changes in binding the assumption of normalisation fail give rise distribution that is artificially symmetric.(C) DeSEQ2 makes use of similar assumptions and results in a similar distortion of data.
+(D) DiffBind utilises normalisation to total library size, and performs significantly better than the other three methods but does not attempt to control for systematic bias in pull-down efficiency of the ChIP.
+
+##Reproducibility plots
+
+```{r, fig.cap="Correlation Plots of Replicate Experiments.
+(A) Scatter plots showing the correlation between the replicates with the lowest correlation value.
+This is provided both the control (top) and treatment (bottom) conditions.
+The plotted condition is highlighted with thick border in tables on the right.
+Colour represents density, blue = lowest, red = highest.
+(B) Tables showing the correlation coefficient for been each replicated.
+"}
 
@@
 
-```{r}
+##Venn diagrams
+
+```{r  out.width = '25%', out.extra='style="float:left"'}
 filename<-"Rdata/040_SLX-14438_dba_human_ER_CTCF.rda"
@@
 ```
-```{r}
+
+```{r, out.width = '25%', out.extra='style="float:left"'}
 
@@
 
-```{r}
+```{r, out.width = '25%', out.extra='style="float:left"'}
 filename<-"Rdata/014_SLX-12998_dba_mouse.rda"
@@
 
+Venn diagrams showing peak overlap between replicates.
+(A) Peak overlap between samples for CTCF internal control.
+(B) Peak overlap for murine chromatin spike-in samples.
+(C) Peak overlap for Drosophila chromatin spike-in.
+As expected, samples treated with fulvestrant showed fewer ER peaks and a higher proportion of peaks unique to individual samples.
+This is indicative in a loss of ER binding leading and the resulting decrease in single to noise ratio.
+As we use a statistical analysis and model the variance of the peaks between replicates, false positives in the treated condition will be removed at the point of differential binding analysis.
+
+##MA Plots of mouse ER normalization
+
 ```{r}
+
+par(mfrow=c(1,3))
 
@@
 mmcounts<-mmconsensus[,-c(1:3)]
-head(hscounts)
-head(mmcounts)
 
@@
 
-```
-
-```{r}
-
 # Residuals of the Drosophila fit
@@
 
-```{r}
-hsconsensus_ER<-readRDS("Rdata/018_SLX-14229_hsconsensus_ER.rds")
-hsconsensus_CTCF<-readRDS("Rdata/018_SLX-14229_hsconsensus_CTCF.rds")
-hscounts_ER<-hsconsensus_ER[,-c(1:3)]
-hscounts_CTCF<-hsconsensus_CTCF[,-c(1:3)]
-
-load("Rdata/012_SLX-14229_aligned.rda")
-aligned<-aligned[grep("SLX-14229.D",names(aligned))]
-aligned<-sapply(aligned,sum)/1E6
-
-#remove inputs/controls to match samples here
-aligned<-aligned[c(1,2,5,7,8,10)]
-
-hsrpm_ER<-hscounts_ER
-for(i in 1:length(hsrpm_ER)){
-  hsrpm_ER[i]<-hscounts_ER[i]/aligned[i]
-}
-M_RPM_ER<-apply(hsrpm_ER,1,function(x){
-  fulvestrant<-mean(x[c(2,4,5)])
-  untreated<-mean(x[c(1,3,6)])
-  fc<-mean(fulvestrant)/mean(untreated)
-  log2fc<-log2(fc)
-  return(log2fc)
-})
-A_RPM_ER<-apply(hsrpm_ER,1,function(x){
-  return(log10(sum(x)))
-})
-plot(A_RPM_ER,M_RPM_ER,cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5,pch=20,ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"), main="RPM aligned reads")
-abline(h=0)
-
-```
-
-```{r}
-
-hsrpm_CTCF<-hscounts_CTCF
-for(i in 1:length(hsrpm_CTCF)){
-  hsrpm_CTCF[i]<-hscounts_CTCF[i]/aligned[i]
-}
-M_RPM_CTCF<-apply(hsrpm_CTCF,1,function(x){
-  fulvestrant<-mean(x[c(2,4,5)])
-  untreated<-mean(x[c(1,3,6)])
-  fc<-mean(fulvestrant)/mean(untreated)
-  log2fc<-log2(fc)
-  return(log2fc)
-})
-A_RPM_CTCF<-apply(hsrpm_CTCF,1,function(x){
-  return(log10(sum(x)))
-})
-
-plot(A_RPM_ER,M_RPM_ER,cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5,pch=20,ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"), main="RPM aligned reads")
-points(A_RPM_CTCF,M_RPM_CTCF,pch=20,col="gray")
-lm1<-lm(M_RPM_CTCF~A_RPM_CTCF)
-abline(lm1$coef,col="blue")
-abline(h=0)
-
-```
-
-```{r}
-
-Mhs_CTCF<-apply(hscounts_CTCF,1,function(x){
-  fulvestrant<-mean(x[c(2,4,5)])
-  untreated<-mean(x[c(1,3,6)])
-  fc<-mean(fulvestrant)/mean(untreated)
-  log2fc<-log2(fc)
-  return(log2fc)
-})
-Ahs_CTCF<-apply(hscounts_CTCF,1,function(x){
-  return(log10(sum(x)))
-})
-
-Mhs_ER<-apply(hscounts_ER,1,function(x){
-  fulvestrant<-mean(x[c(2,4,5)])
-  untreated<-mean(x[c(1,3,6)])
-  fc<-mean(fulvestrant)/mean(untreated)
-  log2fc<-log2(fc)
-  return(log2fc)
-})
-Ahs_ER<-apply(hscounts_ER,1,function(x){
-  return(log10(sum(x)))
-})
-
-lm1<-lm(Mhs_CTCF~Ahs_CTCF)
-intercept<-lm1$coef[1]
-angularcoeff<-lm1$coef[2]
-MhsFit_ER<-Mhs_ER-(Ahs_ER*angularcoeff)-intercept
-MhsFit_CTCF<-Mhs_CTCF-(Ahs_CTCF*angularcoeff)-intercept
-plot(Ahs_ER,MhsFit_ER,cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5,pch=20,ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"), main="Normalised to Human CTCF")
-points(Ahs_CTCF,MhsFit_CTCF,pch=20,col="gray")
-
-abline(h=0)
-legend("topright",legend=c("Human ER",  "Human CTCF"),pch=20,col=c("black","gray"), cex=1)
-
-```
-```{r}
+MA plots showing the addition of Mm derived chromatin spike-in to the ChIP-seq analysis of MCF-7 before and after treatment with fulvestrant.
+(A) MA plot after scaling factor based normalisation shows same characteristic grouping of peaks off axis.
+(B) ER binding in Mm samples shows considerable increase in binding after treatment of the MCF-7 cell line with fulvestrant.
+(C) Attempting to fit a correction factor to the data results in a significant distortion.
+
+##Relative reads aligments in mouse samples
+
+```{r, fig.cap="Distribution of reads for Mm chromatin spike-in normalisation strategy.
+Comparison of murine chromatin between samples showed no systematic bias in the sample preparation.
+Bar plots (left axis) represent the fraction of total aligned reads.
+The dot plot represents the total aligned reads (right axis) for each sample."}
 filename<-"Rdata/012_SLX-12998_nrreads.rda"
@@
 ```
+
+##MA plots of CTCF Parallel-Factor ChIP
+
 ```{r}
+par(mfrow=c(1,3))
+
+hsconsensus_ER<-readRDS("Rdata/018_SLX-14229_hsconsensus_ER.rds")
+hsconsensus_CTCF<-readRDS("Rdata/018_SLX-14229_hsconsensus_CTCF.rds")
+hscounts_ER<-hsconsensus_ER[,-c(1:3)]
+hscounts_CTCF<-hsconsensus_CTCF[,-c(1:3)]
+
+load("Rdata/012_SLX-14229_aligned.rda")
+aligned<-aligned[grep("SLX-14229.D",names(aligned))]
+aligned<-sapply(aligned,sum)/1E6
+
+#remove inputs/controls to match samples here
+aligned<-aligned[c(1,2,5,7,8,10)]
+
+hsrpm_ER<-hscounts_ER
+for(i in 1:length(hsrpm_ER)){
+  hsrpm_ER[i]<-hscounts_ER[i]/aligned[i]
+}
+M_RPM_ER<-apply(hsrpm_ER,1,function(x){
+  fulvestrant<-mean(x[c(2,4,5)])
+  untreated<-mean(x[c(1,3,6)])
+  fc<-mean(fulvestrant)/mean(untreated)
+  log2fc<-log2(fc)
+  return(log2fc)
+})
+A_RPM_ER<-apply(hsrpm_ER,1,function(x){
+  return(log10(sum(x)))
+})
+plot(A_RPM_ER,M_RPM_ER,cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5,pch=20,ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"), main="RPM aligned reads")
+abline(h=0)
+
+hsrpm_CTCF<-hscounts_CTCF
+for(i in 1:length(hsrpm_CTCF)){
+  hsrpm_CTCF[i]<-hscounts_CTCF[i]/aligned[i]
+}
+M_RPM_CTCF<-apply(hsrpm_CTCF,1,function(x){
+  fulvestrant<-mean(x[c(2,4,5)])
+  untreated<-mean(x[c(1,3,6)])
+  fc<-mean(fulvestrant)/mean(untreated)
+  log2fc<-log2(fc)
+  return(log2fc)
+})
+A_RPM_CTCF<-apply(hsrpm_CTCF,1,function(x){
+  return(log10(sum(x)))
+})
+
+plot(A_RPM_ER,M_RPM_ER,cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5,pch=20,ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"), main="RPM aligned reads")
+points(A_RPM_CTCF,M_RPM_CTCF,pch=20,col="gray")
+lm1<-lm(M_RPM_CTCF~A_RPM_CTCF)
+abline(lm1$coef,col="blue")
+abline(h=0)
+
+Mhs_CTCF<-apply(hscounts_CTCF,1,function(x){
+  fulvestrant<-mean(x[c(2,4,5)])
+  untreated<-mean(x[c(1,3,6)])
+  fc<-mean(fulvestrant)/mean(untreated)
+  log2fc<-log2(fc)
+  return(log2fc)
+})
+Ahs_CTCF<-apply(hscounts_CTCF,1,function(x){
+  return(log10(sum(x)))
+})
+
+Mhs_ER<-apply(hscounts_ER,1,function(x){
+  fulvestrant<-mean(x[c(2,4,5)])
+  untreated<-mean(x[c(1,3,6)])
+  fc<-mean(fulvestrant)/mean(untreated)
+  log2fc<-log2(fc)
+  return(log2fc)
+})
+Ahs_ER<-apply(hscounts_ER,1,function(x){
+  return(log10(sum(x)))
+})
+
+lm1<-lm(Mhs_CTCF~Ahs_CTCF)
+intercept<-lm1$coef[1]
+angularcoeff<-lm1$coef[2]
+MhsFit_ER<-Mhs_ER-(Ahs_ER*angularcoeff)-intercept
+MhsFit_CTCF<-Mhs_CTCF-(Ahs_CTCF*angularcoeff)-intercept
+plot(Ahs_ER,MhsFit_ER,cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5,pch=20,ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"), main="Normalised to Human CTCF")
+points(Ahs_CTCF,MhsFit_CTCF,pch=20,col="gray")
+
+abline(h=0)
+legend("topright",legend=c("Human ER",  "Human CTCF"),pch=20,col=c("black","gray"), cex=1)
+
+```
+
+MA plots showing ER binding before and after treatment with fulvestrant including matched CTCF control.} (A) Reads corrected to total aligned reads showed the same off-centre peak density as observed with all that was not-normalised with an internal spike-in control.
+(B) Overlaying the MA plot combining the changes in chromatin binding of ER (black) and CTCF (grey).
+CTCF peaks overlay the off-centre peak density.
+(C) Utilising the CTCF binding events as a ground truth for 0-fold change, a linear fit to the log-fold change is generated (blue line).
+The fit is then also applied to the ER binding events.
+
+##ER and CTCF heatmaps
+
+```{r,results='hide', message=FALSE, warning=FALSE,   out.width = '33%', out.extra='style="float:left"'}
 filename<-"Rdata/015_SLX-14229_dbacounts_human.rda"
@@
 
-```{r}
+Clustering of samples before and after ER and CTCF peak extractions shows the effect of fulvestrant on ER peaks drive clustering of the raw data.} To confirm that the effects seen at the RARa locus were consistent across the genome, we compared the clustering of the CTCF and the ER peaks with respect to the treatment with fulvestrant.
+Initial clustering was weakly correlated with that of the treatment condition (A).
+Clustering specifically to CTCF derived peak data (B) resulted in a loss of grouping by treatment, while clustering specifically ER-derived peak data (C) led to a clearer separation by treatment.
+
+##Normalisation using DESeq2 SizeFactors
+
+```{r, results='hide', message=FALSE, warning=FALSE,  out.width = '50%', out.extra='style="float:left"'}
+
+dbaSummits                <- 200
+jg.controlMinOverlap      <- 5
+jg.controlSampleSheet     <- "samplesheet/samplesheet_SLX8047_dm.csv"
+jg.experimentSampleSheet  <- "samplesheet/samplesheet_SLX8047_hs.csv"
+
+filename<-"Rdata/029_SLX-8047_dba_human_drosophila.rda"
+  load(filename)
+
+jg.experimentPeakset <- jg.dbaGetPeakset(dbaExperiment)
+jg.controlPeakset    <- jg.dbaGetPeakset(dbaControl)
+jg.controlPeaksetDeSeq<-jg.convertPeakset(jg.controlPeakset)
+jg.controlSizeFactors = estimateSizeFactorsForMatrix(jg.controlPeaksetDeSeq)
+jg.conditions <- read.csv(file=jg.controlSampleSheet, header=TRUE, sep=",")['Condition']
+jg.controlDeSeq<-jg.runDeSeq(jg.controlPeaksetDeSeq, jg.conditions,jg.controlSizeFactors)
+jg.controlResultsDeseq   = results(jg.controlDeSeq)
+jg.experimentPeaksetDeSeq<-jg.convertPeakset(jg.experimentPeakset)
+jg.experimentDeSeq<-jg.runDeSeq(jg.experimentPeaksetDeSeq, jg.conditions,jg.controlSizeFactors)
+jg.experimentResultsDeseq   = results(jg.experimentDeSeq)
+
+par(mar=c(5.1,5.1,4.1,2.1))
+jg.plotDeSeq(jg.experimentResultsDeseq,
+             p=0.01,
+             title.main="Fold-change in ER binding",
+             flip=T
+             )
+
+#Repeat not using out control
+jg.experimentDeSeqInternal<-jg.runDeSeq(jg.experimentPeaksetDeSeq, jg.conditions, NULL)
+jg.experimentResultsDeseqInternal   = results(jg.experimentDeSeqInternal)
+
+  jg.plotDeSeq(jg.experimentResultsDeseqInternal,
+               title.main="Fold-change in ER binding (no correction)",
+               p=0.01,
+               flip=T
+               )
+
+```
+Normalisation of ER binding external spike implemented using DESeq2.
+Highlighted data points are considered significant fold-changes with a FDR = 0.01.
+(A) Initial analysis of the ER binding with default parameters shows an equal increase and decrease in ER binding.
+The distribution seen is not reflective of the documented response of ER on treatment with fulvestrant.
+(B) Estimating the DESeq2 size factors from the sample spike-in corrects the distortion in the results.
+
+```{r, results='hide', message=FALSE, warning=FALSE,  out.width = '50%', out.extra='style="float:left"'}
+
+dbaSummits                <- 200
+jg.controlMinOverlap      <- 5
+jg.controlSampleSheet     <- "samplesheet/samplesheet_SLX14229_hs_CTCF_DBA.csv"
+jg.experimentSampleSheet  <- "samplesheet/samplesheet_SLX14229_hs_ER_DBA.csv"
+
+filename<-"Rdata/028_SLX-14229_dba_human_ER_CTCF.rda"
+
+  load(filename)
+
+jg.experimentPeakset <- jg.dbaGetPeakset(dbaExperiment)
+jg.controlPeakset    <- jg.dbaGetPeakset(dbaControl)
+jg.controlPeaksetDeSeq<-jg.convertPeakset(jg.controlPeakset)
+jg.controlSizeFactors = estimateSizeFactorsForMatrix(jg.controlPeaksetDeSeq)
+jg.conditions <- read.csv(file=jg.controlSampleSheet, header=TRUE, sep=",")['Condition']
+jg.controlDeSeq<-jg.runDeSeq(jg.controlPeaksetDeSeq, jg.conditions,jg.controlSizeFactors)
+jg.controlResultsDeseq   = results(jg.controlDeSeq)
+jg.experimentPeaksetDeSeq<-jg.convertPeakset(jg.experimentPeakset)
+jg.experimentDeSeq<-jg.runDeSeq(jg.experimentPeaksetDeSeq, jg.conditions,jg.controlSizeFactors)
+jg.experimentResultsDeseq   = results(jg.experimentDeSeq)
+
+par(mar=c(5.1,5.1,4.1,2.1))
+jg.plotDeSeq(jg.experimentResultsDeseq,
+             p=0.01,
+             title.main="Fold-change in ER binding",
+             flip=T
+             )
+
+#Repeat not using out control
+jg.experimentDeSeqInternal<-jg.runDeSeq(jg.experimentPeaksetDeSeq, jg.conditions, NULL)
+jg.experimentResultsDeseqInternal   = results(jg.experimentDeSeqInternal)
+
+  jg.plotDeSeq(jg.experimentResultsDeseqInternal,
+               title.main="Fold-change in ER binding (no correction)",
+               p=0.01,
+               flip=T
+               )
+
+```
+Normalisation of ER binding internal CTCF control.
+Highlighted data points are considered significant fold-changes with a FDR = 0.01.
+(A) Initial analysis with default DESeq2 parameters gives similar distortion as seen previously.
+(B) Correction using the CTCF peaks to provide an internal control allows for the data to be corrected.
+
+##Comparison of mormalisation DiffBind plots
+
+```{r,results='hide', message=FALSE, warning=FALSE,   out.width = '32%', out.extra='style="float:left"'}
 jg.controlSampleSheet     <- "samplesheet/samplesheet_SLX14438_hs_CTCF_DBA.csv"
@@
 
-dba.plotMA(jg.dba_analysis,bFlip=TRUE, cex.lab=1.5, cex.axis=1.5, cex.main=1.25, cex.sub=1.5)
+#dba.plotMA(jg.dba_analysis,bFlip=TRUE, cex.lab=1.5, cex.axis=1.5, cex.main=1.25, cex.sub=1.5)
 dba_analysis<-dba.analyze(dbaExperiment)
@@
 
-```
-```{r}
 filename<-"Rdata/040_SLX-14438_dba_human_ER_CTCF.rda"
@@
   #Again numbers similar but not exact due to DiffBind
-```
-```{r}
-filename<-"Rdata/032_SLX-14229_dba_human_ER_CTCF_normalised.rda"
-load(filename)
-
-filename<-"Rdata/032_SLX-8047_dba_normalised.rda"
-load(filename)
-
-jg.dba_analysis_SLX8047<-dba.analyze(jg.dba_SLX8047)
-jg.dba_analysis_SLX14229<-dba.analyze(jg.dba_SLX14229)
-
-nearestPeak <- nearest(dba.report(jg.dba_analysis_SLX8047, th=1),dba.report(jg.dba_analysis_SLX14229, th=1))
-SLX8047_fc<-dba.report(jg.dba_analysis_SLX8047,bFlip=TRUE, th=1)$Fold
-SLX14229_fc<-dba.report(jg.dba_analysis_SLX14229,th=1)[nearestPeak]$Fold
-
-smoothScatter(SLX8047_fc,SLX14229_fc,pch=20, col="#00000055",
-              xlab="Fold change normalized using Drosophila/H2av spike-in",
-              ylab="Fold change normalized using CTCF control peaks",
-              main="Comparision of normalisation methods",
-              cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5)
-abline(h=0, col="grey",lty=2)
-abline(v=0, col="grey",lty=2)
-coef<-lm( SLX8047_fc ~ SLX14229_fc + 0 )
-abline(coef, col="navyblue")
-text(3.25,-0.25, expression( "gradient = 0.96, r = 0.557, p-value < 2.2 x" ~ 10^{-16}), cex=1.25)
-
-```
-```{r}
-
-```
-
-```{r}
+
+#Finally plot the same result as earlier so it is the same as in the text to generate our CTCF and linear regression figures
+  dba.plotMA(jg.dba_analysis_noER,bFlip=TRUE,cex.lab=1.5, cex.axis=1.5, cex.main=1.25, cex.sub=1.5)
+
+```
+Comparison of DiffBind results before and after our two methods of normalisation.
+(A) Normalisation to Library Size.
+(B) Applying the corrected size factors from our DESeq2 pipeline generated from CTCF internal control.
+(C) Applying correction using linear regression of CTCF peaks between conditions to normalise data.
+The result is a 10.7% increase in the number of loci detected as significantly changed ER binding.
+
+##Cross normalisation
+
+```{r results='hide', message=FALSE, warning=FALSE, fig.cap="Comparison of fold-change of ER binding after both xenogenic and cross-normalisation.
+Pearson's correlation between the two methods is 0.992 (3sf, p-value tending to 0).
+Deviation of data points from parity is a result the integer nature of read counts, nonetheless this effect is is very small as demonstrated correlation coefficient between the two datasets."}
 filename<-"Rdata/047_SLX-14438_dba_human_ER_CTCF.rda"

### 2017-11-14 049_ChIPComp.R

@@ -42,6 +42,10 @@ countSet=makeCountSet(conf,design,filetype="bam", species="hg19")
 countSet=ChIPComp(countSet)
 countSet$db[c(10:15,17)]
 
+filename="Rdata/047_countset.Rda"
+save(countSet,file=filename)
+
+load(filename)
 c1<-rowMeans(countSet$db[c(4,6,9)])
 c0<-rowMeans(countSet$db[c(5,7,8)])
 c1<-c1[]

### 2017-11-26 052_ChIPQC.R
Case Studies

@@ -7,8 +7,13 @@ qc.samplesheets=c(
     "samplesheet/samplesheet_SLX8047.csv",
     "samplesheet/samplesheet_SLX12998.csv",
     "samplesheet/samplesheet_SLX14229.csv",
-    "samplesheet/samplesheet_SLX14438_hs_DBA.csv")
+    "samplesheet/samplesheet_SLX14438_hs_DBA.csv",
+    "samplesheet/samplesheet_SLX15090_H4K12ac.csv",
+    "samplesheet/samplesheet_SLX15091_ER.csv")
 
+#library(BiocParallel)
+#register(MulticoreParam())
+#register(SerialParam())
 
 for (qc.samplesheet in qc.samplesheets) {
     dba.ChIPQC<-ChIPQC(qc.samplesheet, "hg19", bCount=T)

### 2017-11-26 062_SLX-15090_macs.sh

@@ -5,7 +5,8 @@ control=../blacklist_filtered/SLX-15090.D702_D503.HMTVKBBXX.s_7.bwa.homo_sapiens
 for bam in ../blacklist_filtered/*bam
 do
 	root=`basename $bam .bam`
-	macs2 callpeak -t $bam -c $control -f BAM -n $root -g hs
+	#macs2 callpeak -t $bam -c $control -f BAM -n $root -g hs
+	macs2 callpeak -t $bam -c $control --broad -f BAM -n $root -g hs
 done
 
 

### 2017-11-26 063_SLX-15090_MergeCTCF.sh

@@ -1,0 +1,7 @@
+cd SLX-15090/peaks
+cat SLX-15090.D702_D504.HMTVKBBXX.s_7.bwa.homo_sapiens_peaks.narrowPeak SLX-15090.D702_D501.HMTVKBBXX.s_7.bwa.homo_sapiens_peaks.narrowPeak > CTCF_merged.bed
+grep -v \# CTCF_merged.bed  | grep -v start > CTCF_merged_filtered.bed
+sort -k1,1 -k2,2n CTCF_merged_filtered.bed > CTCF_merged_sorted_filtered.bed
+bedtools merge -i CTCF_merged_sorted_filtered.bed  >CTCF_union.bed
+
+

### 2017-11-26 065_SLX-15090_SplitCTCF.sh

@@ -1,0 +1,17 @@
+cd SLX-15090/peaks/
+mkdir CTCF
+
+for f in *.broadPeak
+do
+	echo $f
+	bedtools intersect -a $f -b CTCF_union.bed > CTCF/$f 
+done
+
+mkdir H4K12ac 
+for f in *.broadPeak
+do
+	echo $f
+	bedtools subtract -b CTCF_union.bed -a $f -A > H4K12ac/$f 
+done
+
+

### 2017-11-26 066_SLX-15090_Brundle.R

@@ -1,0 +1,219 @@
+
+library(devtools)
+
+install_github("andrewholding/Brundle")
+library(Brundle)
+
+setwd("/Volumes/FlyPeakCaseStudy/BrundleDevelopment")
+
+#Tidy naming to ER 
+
+jg.controlSampleSheet     <- "samplesheet/samplesheet_SLX15090_CTCF.csv"
+jg.experimentSampleSheet  <- "samplesheet/samplesheet_SLX15090_H4K12ac.csv"
+jg.treatedCondition       =  "Estrogen"
+jg.untreatedCondition     =  "none"
+
+#####
+#
+# Main Code
+#
+#####
+
+filename<-"Rdata/066_SLX-15090_dba_human_ER_CTCF.rda"
+if(!file.exists(filename)){
+    dbaExperiment <- jg.getDba(jg.experimentSampleSheet, bRemoveDuplicates = TRUE)
+    dbaControl    <- jg.getDba(jg.controlSampleSheet, bRemoveDuplicates = TRUE)
+    save(dbaExperiment,dbaControl,file=filename)
+} else {
+    load(filename)
+}
+
+jg.dba<-Brundle(
+            dbaExperiment,
+            dbaControl,
+            jg.treatedCondition,
+            jg.untreatedCondition,
+            jg.experimentSampleSheet,
+            jg.controlSampleSheet )
+    
+jg.dba_analysis<-dba.analyze(jg.dba)
+png("plots/066__SLX15090_MAplot.png",pointsize=15)
+dba.plotMA(jg.dba_analysis,bFlip=TRUE,th=0.05)
+dev.off()
+
+#Check CTCF 
+ctcf.dba.analysis<-dba.analyze(dbaControl)
+dba.plotMA(ctcf.dba.analysis,bFlip=TRUE) 
+
+
+#IDR for manuscript
+
+
+jg.plotScatter<-function(peaks,samples,yaxis,samplesCombinations,noYLabel=FALSE)
+{
+    peaks$count <- 1
+    peaks_aggregate <- aggregate(count ~ ., peaks, FUN = sum)
+    colormap<-rainbow(255,start=0,end=4/6)
+    colormap<-rev(colormap)
+    peaks_aggregate$Col <- colormap[as.numeric(cut(log(peaks_aggregate$count,2),breaks=255))]
+    peaks_aggregate<-peaks_aggregate[order(peaks_aggregate$count),]
[…]

### 2017-11-26 066_SLX-15091_Brundle.R

@@ -35,8 +35,146 @@ jg.dba<-Brundle(
             jg.controlSampleSheet )
     
 jg.dba_analysis<-dba.analyze(jg.dba)
+png("plots/066_SLX15091_MAplot.png",pointsize=15)
 dba.plotMA(jg.dba_analysis,bFlip=TRUE,th=0.05)
+dev.off()
 
 #Check CTCF 
 ctcf.dba.analysis<-dba.analyze(dbaControl)
 dba.plotMA(ctcf.dba.analysis,bFlip=TRUE) 
+
+dba.report(jg.dba_analysis)
+
+
+#IDR for manuscript
+
+
+jg.plotScatter<-function(peaks,samples,yaxis,samplesCombinations,noYLabel=FALSE)
+{
+    peaks$count <- 1
+    peaks_aggregate <- aggregate(count ~ ., peaks, FUN = sum)
+    colormap<-rainbow(255,start=0,end=4/6)
+    colormap<-rev(colormap)
+    peaks_aggregate$Col <- colormap[as.numeric(cut(log(peaks_aggregate$count,2),breaks=255))]
+    peaks_aggregate<-peaks_aggregate[order(peaks_aggregate$count),]
+    plot(log(peaks_aggregate[,c(1,2)]), col=peaks_aggregate$Col,pch=20, cex=0.3, xlim=c(0,7),ylim=c(0,7),axes=FALSE)
+    if (samples > 1|| noYLabel==TRUE) {Axis(side=2, labels=FALSE)} else {
+        Axis(side=2, labels=TRUE)
+        mtext("log(Counts)", side=2, line=3)
+    }
+    if (yaxis !=TRUE ) {Axis(side=1, labels=FALSE)} else {
+        Axis(side=1, labels=TRUE,ylab="Log(Counts)")
+        mtext("log(Counts)", side=1, line=3)
+    }
+    text(1.5,6.5,paste0(samplesCombinations[,samples][1],"-",samplesCombinations[,samples][2]),cex=1.5)
+    sampleCor<-cor(log(peaks[1]),log(peaks[2]))
+    text(2,6.0,paste0("Correlation = ",signif(sampleCor,3)),cex=1.0)
+    abline(a=0,b=1,col="grey")
+    return(cor(log(peaks_aggregate[,c(1,2)])))
+}
+
+
+jg.plotMargins<-function(a,b)
+{
[…]

### 2017-11-26 067_SLX15091_motifAnalysis.sh

@@ -1,0 +1,3 @@
+cd motifAnalysis
+findMotifsGenome.pl ../bed/ER_45minsGains.bed hg19 ER45minBinding
+

### 2017-11-26 068_Homer_Profile.sh

@@ -1,0 +1,9 @@
+cd SLX-15090/
+makeTagDirectory H4K12ac-E2-ChIP-Seq/ SLX-15090.D702_D502.HMTVKBBXX.s_7.bwa.homo_sapiens.bam SLX-15090.D703_D501.HMTVKBBXX.s_7.bwa.homo_sapiens.bam SLX-15090.D702_D502.HMTVKBBXX.s_7.bwa.homo_sapiens.bam 
+makeTagDirectory H4K12ac-None-ChIP-Seq/  SLX-15090.D701_D501.HMTVKBBXX.s_7.bwa.homo_sapiens.bam  SLX-15090.D701_D504.HMTVKBBXX.s_7.bwa.homo_sapiens.bam  SLX-15090.D703_D502.HMTVKBBXX.s_7.bwa.homo_sapiens.bam
+
+annotatePeaks.pl tss hg19 -size 6000 -hist 25 -d H4K12ac-E2-ChIP-Seq/ H4K12ac-None-ChIP-Seq/ > Homer_Profile.txt
+annotatePeaks.pl ../bed/ERConsensus.bed hg19 -size 6000 -hist 25 -d H4K12ac-E2-ChIP-Seq/ H4K12ac-None-ChIP-Seq/ > Homer_ER_Profile.txt
+annotatePeaks.pl ../motifAnalysis/bed/output.bed hg19 -size 6000 -hist 25 -d H4K12ac-E2-ChIP-Seq/ H4K12ac-None-ChIP-Seq/ > Homer_ERE_Profile.txt
+
+

### 2017-11-28 Manuscript_Figures.Rmd
Updated Markdown

@@
 ```
+
+## Changes in H4K12ac following E2 treatment
+
+```{r,results='hide', message=FALSE, warning=FALSE}
+filename<-"Rdata/066_SLX-15090_dba_human_ER_CTCF.rda"
+load(filename)
+
+#install_github("andrewholding/Brundle")
+library(Brundle)
+
+jg.controlSampleSheet     <- "samplesheet/samplesheet_SLX15090_CTCF.csv"
+jg.experimentSampleSheet  <- "samplesheet/samplesheet_SLX15090_H4K12ac.csv"
+jg.treatedCondition       =  "Estrogen"
+jg.untreatedCondition     =  "none"
+
+jg.dba<-Brundle(
+            dbaExperiment,
+            dbaControl,
+            jg.treatedCondition,
+            jg.untreatedCondition,
+            jg.experimentSampleSheet,
+            jg.controlSampleSheet )
+
+jg.dba_analysis<-dba.analyze(jg.dba)
+dba.plotMA(jg.dba_analysis,bFlip=TRUE,th=0.05)
+
+```
+
+```{r,results='hide', message=FALSE, warning=FALSE}
+
+jg.plotScatter<-function(peaks,samples,yaxis,samplesCombinations,noYLabel=FALSE)
+{
+    peaks$count <- 1
+    peaks_aggregate <- aggregate(count ~ ., peaks, FUN = sum)
+    colormap<-rainbow(255,start=0,end=4/6)
+    colormap<-rev(colormap)
+    peaks_aggregate$Col <- colormap[as.numeric(cut(log(peaks_aggregate$count,2),breaks=255))]
+    peaks_aggregate<-peaks_aggregate[order(peaks_aggregate$count),]
+    plot(log(peaks_aggregate[,c(1,2)]), col=peaks_aggregate$Col,pch=20, cex=0.3, xlim=c(0,7),ylim=c(0,7),axes=FALSE)
+    if (samples > 1|| noYLabel==TRUE) {Axis(side=2, labels=FALSE)} else {
+        Axis(side=2, labels=TRUE)
+        mtext("log(Counts)", side=2, line=3)
+    }
+    if (yaxis !=TRUE ) {Axis(side=1, labels=FALSE)} else {
+        Axis(side=1, labels=TRUE,ylab="Log(Counts)")
+        mtext("log(Counts)", side=1, line=3)
+    }
+    text(1.5,6.5,paste0(samplesCombinations[,samples][1],"-",samplesCombinations[,samples][2]),cex=1.5)
+    sampleCor<-cor(log(peaks[1]),log(peaks[2]))
+    text(2,6.0,paste0("Correlation = ",signif(sampleCor,3)),cex=1.0)
+    abline(a=0,b=1,col="grey")
+    return(cor(log(peaks_aggregate[,c(1,2)])))
+}
+
+jg.plotPanelsPlot<-function(treatmentNames,controlNames,dbaExperiment,noYLabel=FALSE)
+{
+
+    jg.experimentPeakset <- jg.dbaGetPeakset(dbaExperiment)
+    relativeLibrarySize<-jg.getLibrarySize(dbaExperiment,
+                                           jg.experimentPeakset)
+
+    samplesCombinations<-combn(controlNames,2)
+    for (samples in seq(1,length(samplesCombinations)/2))  {
+        peaksLibrarySize<-relativeLibrarySize[samplesCombinations[,samples]]
+        peaks<-jg.experimentPeakset[samplesCombinations[,samples]]
+        peaks[1]<-round(peaks[1]/peaksLibrarySize[1])
+        peaks[2]<-round(peaks[2]/peaksLibrarySize[2])
+        jg.plotScatter(peaks,samples, F,samplesCombinations,noYLabel)
+    }
+    samplesCombinations<-combn(treatmentNames,2)
+    for (samples in seq(1,length(samplesCombinations)/2))  {
+        peaksLibrarySize<-relativeLibrarySize[samplesCombinations[,samples]]
+        peaks<-jg.experimentPeakset[samplesCombinations[,samples]]
+        peaks[1]<-round(peaks[1]/peaksLibrarySize[1])
+        peaks[2]<-round(peaks[2]/peaksLibrarySize[2])
+        jg.plotScatter(peaks,samples,T,samplesCombinations,noYLabel)
+    }
+}
+
+jg.getLibrarySize<-function(dbaExperiment,jg.experimentPeakset) {
+    librarySize<-as.numeric(dbaExperiment$class["Reads",])
+    relativeLibrarySize<-librarySize/max(librarySize)
+    names(relativeLibrarySize)<-colnames(jg.experimentPeakset[c(-1:-3)])
+    return(relativeLibrarySize)
+}
+
+#Worse correlation is 2a-3a (control sample)
+controlNames<-c("2a","3a")
+
+jg.experimentPeakset <- jg.dbaGetPeakset(dbaExperiment)
+relativeLibrarySize<-jg.getLibrarySize(dbaExperiment,
+                                       jg.experimentPeakset)
+peaksLibrarySize<-relativeLibrarySize[controlNames]
+peaks<-jg.experimentPeakset[controlNames]
+peaks[1]<-round(peaks[1]/peaksLibrarySize[1])
+peaks[2]<-round(peaks[2]/peaksLibrarySize[2])
+x<-c("Rep2","Rep3")
+names(x)<-controlNames
+x<-as.matrix(x)
+jg.plotScatter(peaks,controlNames, T,t(x),FALSE)
+Axis(side=2, labels=TRUE,ylab="Log(Counts)")
+mtext("log(Counts)", side=2, line=3)
+
+```
+
+```{r,results='hide', message=FALSE, warning=FALSE}
+
+profile<-read.table("SLX-15090/Homer_Profile.txt",sep="\t")
+profile_ER<-as.matrix(profile[,c(1,2)])
+plot(profile_ER,t="l",col="red", xlab="Distance from TSS",ylab="Read Depth",main="H4K12ac Profile")
+profile_Ctrl<-as.matrix(profile[,c(1,5)])
+lines(profile_Ctrl,col="blue")
+legend("topright",legend = c("Estrogen", "Control"), col=c("red","blue"),pch = c(19,19))
+abline(v=0, lty=3)
+
+```
+
+## Activation of ER in MCF7
+
+```{r,results='hide', message=FALSE, warning=FALSE}
+
+jg.controlSampleSheet     <- "samplesheet/samplesheet_SLX15091_CTCF.csv"
+jg.experimentSampleSheet  <- "samplesheet/samplesheet_SLX15091_ER.csv"
+jg.treatedCondition       =  "Estrogen"
+jg.untreatedCondition     =  "none"
+
+#####
+#
+# Main Code
+#
+#####
+
+filename<-"Rdata/065_SLX-15091_dba_human_ER_CTCF.rda"
+load(filename)
+
+jg.dba<-Brundle(
+            dbaExperiment,
+            dbaControl,
+            jg.treatedCondition,
+            jg.untreatedCondition,
+            jg.experimentSampleSheet,
+            jg.controlSampleSheet )
+
+jg.dba_analysis<-dba.analyze(jg.dba)
+dba.plotMA(jg.dba_analysis,bFlip=TRUE,th=0.05)
+
+```
+
+```{r,results='hide', message=FALSE, warning=FALSE}
+#Check CTCF
+
+controlNames<-c("1a","3b")
+
+jg.experimentPeakset <- jg.dbaGetPeakset(dbaExperiment)
+relativeLibrarySize<-jg.getLibrarySize(dbaExperiment,
+                                       jg.experimentPeakset)
+peaksLibrarySize<-relativeLibrarySize[controlNames]
+peaks<-jg.experimentPeakset[controlNames]
+peaks[1]<-round(peaks[1]/peaksLibrarySize[1])
+peaks[2]<-round(peaks[2]/peaksLibrarySize[2])
+x<-c("Rep1","Rep3")
+names(x)<-controlNames
+x<-as.matrix(x)
+jg.plotScatter(peaks,controlNames, T,t(x),FALSE)
+Axis(side=2, labels=TRUE,ylab="Log(Counts)")
+mtext("log(Counts)", side=2, line=3)
+
+```

### 2017-11-28 066_SLX-15091_Brundle.R

@@ -170,7 +170,7 @@ x["9769",] #RARa
 barplot(2^as.matrix(x["9769",c("Conc_none","Conc_Estrogen")]), main="RARa",xlab="Condition",ylab="Read Depth",names.arg=c("Ctrl","E2"))
 x<-report[report$Chr=='chr21',] 
 x["14816",] #NRIP
-barplot(2^as.matrix(x["14816",c("Conc_none","Conc_Estrogen")]), main="NRIP",xlab="Condition",ylab="Read Depth",names.arg=c("Ctrl","E2"))
+barplot(2^as.matrix(x["14816",c("Conc_none","Conc_Estrogen")]), main="NRIP1",xlab="Condition",ylab="Read Depth",names.arg=c("Ctrl","E2"))
 x<-report[report$Chr=='chr22',] 
 x["15381",] #XBP1
 barplot(2^as.matrix(x["15381",c("Conc_none","Conc_Estrogen")]), main="XBP1",xlab="Condition",ylab="Read Depth",names.arg=c("Ctrl","E2"))

### 2017-11-28 Manuscript_Figures.Rmd
Markdown updated

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, messages=FALSE)
+knitr::opts_chunk$set(echo = FALSE, messages=FALSE, cache=TRUE)
 ```
@@
 
+Comparison of fold-change of H4 acetylation (Lys12) before and after treatment with estradiol.
+MA plot of H4K12ac after normalisation to CTCF binding displays an increase at 2 hours after treatment with estradiol.
+
 ```{r,results='hide', message=FALSE, warning=FALSE}
@@
     peaks_aggregate<-peaks_aggregate[order(peaks_aggregate$count),]
-    plot(log(peaks_aggregate[,c(1,2)]), col=peaks_aggregate$Col,pch=20, cex=0.3, xlim=c(0,7),ylim=c(0,7),axes=FALSE)
+    plot(log(peaks_aggregate[,c(1,2)]), col=peaks_aggregate$Col,pch=20, cex=0.3, xlim=c(0,7),ylim=c(0,7),axes=FALSE, xlab="",ylab="")
     if (samples > 1|| noYLabel==TRUE) {Axis(side=2, labels=FALSE)} else {
@@
 
+Comparison of log(Counts) for binding sites was under taken to confirm reproducibility.
+The data with the lowest correlation is shown and was seen between Replicate 2 and Repelicate 3 in the control condition.
+
 ```{r,results='hide', message=FALSE, warning=FALSE}
@@
 
+H4k12ac occupancy profile before and after treatment with E2 shows a general increase around transcription start sites (TSS).
+
 ## Activation of ER in MCF7
@@
 
+Comparison of fold-change of ER binding before and after treatment with estradiol.
+MA plot of ER binding after normalisation to CTCF binding displays a significant increase in ER binding at 45 minutes after treatment with estradiol.
+
 ```{r,results='hide', message=FALSE, warning=FALSE}
-#Check CTCF
 
@@
 ```
+
+Comparison of log(Counts) for binding sites was under taken to confirm reproducibility.
+The data with the lowest correlation is shown and was seen between Replicate 1 and Replicate 3 in the control condition.

### 2017-11-29 066_SLX-15090_Brundle.R
p-value for SLX15090

@@ -41,10 +41,12 @@ png("plots/066__SLX15090_MAplot.png",pointsize=15)
 dba.plotMA(jg.dba_analysis,bFlip=TRUE,th=0.05)
 dev.off()
 
+report<-dba.report(jg.dba_analysis)
+t.test(report$Conc_Estrogen,report$Conc_none,alternative="greater") #p-value = 0.01834
+
 #Check CTCF 
 ctcf.dba.analysis<-dba.analyze(dbaControl)
 dba.plotMA(ctcf.dba.analysis,bFlip=TRUE) 
-
 
 #IDR for manuscript
 

### 2017-12-05 066_SLX-15090_Brundle.R
Reanalysis of H4K12ac data

@@ -19,7 +19,7 @@ jg.untreatedCondition     =  "none"
 #
 #####
 
-filename<-"Rdata/066_SLX-15090_dba_human_ER_CTCF.rda"
+filename<-"Rdata/066_SLX-15090_dba_human_ER_CTCF_merged.rda"
 if(!file.exists(filename)){
     dbaExperiment <- jg.getDba(jg.experimentSampleSheet, bRemoveDuplicates = TRUE)
     dbaControl    <- jg.getDba(jg.controlSampleSheet, bRemoveDuplicates = TRUE)
@@ -42,7 +42,7 @@ dba.plotMA(jg.dba_analysis,bFlip=TRUE,th=0.05)
 dev.off()
 
 report<-dba.report(jg.dba_analysis)
-t.test(report$Conc_Estrogen,report$Conc_none,alternative="greater") #p-value = 0.01834
+t.test(report$Conc_Estrogen,report$Conc_none,alternative="greater") #p-value = 5.677e-12
 
 #Check CTCF 
 ctcf.dba.analysis<-dba.analyze(dbaControl)
@@ -126,7 +126,7 @@ jg.plotPanelsPlot(treatmentNames,
 )
 
 
-png("plots/066_SLX15091_scatter.png", pointsize = 15)
+png("plots/066_SLX15090_scatter.png", pointsize = 15)
 #Worse correlation is 2a-3a (control sample)
 controlNames<-c("2a","3a")
 jg.plotMargins(1,1)
@@ -150,8 +150,8 @@ dev.off()
 ##Bind at prviously reported sites
 dba.report(jg.dba_analysis)
 countSites<-dba.report(jg.dba_analysis, th=1)
-countSites[countSites$Fold > 0] #3909 sites (Note Fold is flipped this is DOWN)
-countSites[countSites$Fold < 0] #9899 sites
+countSites[countSites$Fold > 0] #4817 sites (Note Fold is flipped this is DOWN)
+countSites[countSites$Fold < 0] #11393 sites
 
 
 report <- dba.report(jg.dba_analysis, th=1,
@@ -159,18 +159,21 @@ report <- dba.report(jg.dba_analysis, th=1,
 scores <- -10*(log10(report$FDR))
 sites  <- cbind(report[,1:3],rownames(report),scores)
 
+#GREB1 chr2;11664454-11665753 FDR=2.71e-04
[…]

### 2017-12-05 069_SLX14438_Stability.R

@@ -1,0 +1,68 @@
+library("Brundle")
+
+jg.controlSampleSheet     <- "samplesheet/samplesheet_SLX14438_hs_CTCF_DBA.csv"
+jg.experimentSampleSheet  <- "samplesheet/samplesheet_SLX14438_hs_ER_DBA.csv"
+jg.treatedCondition       =  "Fulvestrant"
+jg.untreatedCondition     =  "none"
+
+setwd("/Volumes/FlyPeakCaseStudy/BrundleDevelopment")
+filename<-"Rdata/048_SLX-14438_dba_human_ER_CTCF.rda"
+load(filename)
+
+
+
+#Load Sample Ids from control sample sheet.
+jg.sampleIds <- jg.getSampleIds(jg.controlSampleSheet)
+
+## Extract Peak set from DiffBind
+jg.experimentPeakset <- jg.dbaGetPeakset(dbaExperiment)
+jg.controlPeakset    <- jg.dbaGetPeakset(dbaControl)
+
+
+#Get counts for each condition
+jg.controlCountsTreated<-jg.getControlCounts(jg.controlPeakset, 
+                                             jg.controlSampleSheet,
+                                             jg.treatedCondition)
+jg.controlCountsUntreated<-jg.getControlCounts(jg.controlPeakset,
+                                               jg.controlSampleSheet,
+                                               jg.untreatedCondition)
+
+#Get sample names for conditions
+jg.untreatedNames <- names(jg.controlCountsUntreated)
+jg.treatedNames   <- names(jg.controlCountsTreated)
+
+
+jg.coefficient_baseline<-jg.getNormalizationCoefficient(jg.controlCountsTreated,
+                                               jg.controlCountsUntreated)
+rows<-nrow(jg.controlCountsTreated)
+
+jg.coefficient<-matrix(nrow=100,ncol=100)
+
+
+for (n in 100:1)
+{
+    for (i in 1:100)
+    {
+        set.seed(i)
+        sampleRows<-sample(rows, rows*(n/100) )
+        jg.coefficient[n,i]<-jg.getNormalizationCoefficient(jg.controlCountsTreated[sampleRows, ],
+                               jg.controlCountsUntreated[sampleRows, ])
+    }
+    }
+}
+
+png("plots/069_stability.png",pointsize="15")
+plot( rowMeans(jg.coefficient)/jg.coefficient_baseline*100-100,ylim=c(-2,+2),
[…]

### 2017-12-05 SLX-15090/peaks/SLX-15090.D701_D501.HMTVKBBXX.s_7.bwa.homo_sapiens_model.r

@@ -1,10 +1,10 @@
 # R script for Peak Model
 #  -- generated by MACS
-p <- c(0.0150170275326,0.0175084769728,0.0199911660421,0.0225702191897,0.0249390234394,0.0250581644816,0.0250196188503,0.025063420704,0.0250266271469,0.0250756852231,0.0250739331489,0.0251159829285,0.0251878179686,0.0253017027883,0.025380546125,0.0254190917563,0.0255434890209,0.0254804143515,0.0254716539808,0.0255872908746,0.0256170761352,0.0256889111753,0.0257011756943,0.0255540014658,0.0256994236202,0.0257712586603,0.025685407027,0.0257274568066,0.0258868955541,0.0259254411854,0.0260183011153,0.0260253094119,0.0260130448929,0.026167227418,0.0262600873479,0.0262092771976,0.0263949970574,0.0265404192118,0.0264931132097,0.0265071298029,0.0266385353641,0.0267804533702,0.0269749336008,0.0269801898232,0.0270204875286,0.0270432644926,0.0270818101239,0.0270169833804,0.0272500092422,0.0272552654646,0.0272903069476,0.0272377447231,0.0272675299837,0.0272552654646,0.0272535133905,0.0273901751741,0.0272254802041,0.0274532498435,0.0273341088014,0.0273726544326,0.027435729102,0.0275128203646,0.0275320931802,0.0275163245129,0.0274988037714,0.027400687619,0.0277230692625,0.0276302093326,0.0276179448135,0.0277493503747,0.0277458462264,0.027740590004,0.027670507038,0.0278299457856,0.0280314343127,0.0282679643229,0.0283222786215,0.0283030058058,0.0285290233711,0.0285956021888,0.0286534206357,0.0286797017479,0.0288794382009,0.0289950750948,0.0290984474696,0.0290266124295,0.0288426446438,0.0290196041329,0.0288706778302,0.028939008722,0.029103703692,0.0293805314076,0.0294015562974,0.02938578763,0.0292403654757,0.0294365977804,0.0295995406762,0.0296976568286,0.0299552117285,0.0299096578006,0.0298308144639,0.0296994089027,0.0299692283217,0.0299692283217,0.0302022541836,0.0302337915182,0.03039498234,0.0305666856066,0.030712107761,0.0308119759875,0.0308610340637,0.0309819271799,0.0308259925807,0.0309188525106,0.0309784230316,0.030971414735,0.0310362414786,0.0311326055568,0.0311501262983,0.0313621272703,0.031
[…]

### 2017-12-05 SLX-15090/peaks/SLX-15090.D701_D502.HMTVKBBXX.s_7.bwa.homo_sapiens_model.r

@@ -1,10 +1,10 @@
 # R script for Peak Model
 #  -- generated by MACS
-p <- c(0.0141402831563,0.0164624534982,0.0187867387493,0.0211871607329,0.0235283652579,0.0236848685414,0.0236785238137,0.0237673500016,0.0237165921799,0.0237842692755,0.0237398561815,0.0238202227325,0.0238392569156,0.023976726016,0.0241036205702,0.0240761267501,0.0243024220385,0.0243320307678,0.0243806736802,0.02429819222,0.0245350620545,0.0246154286055,0.0247380933412,0.0245921646039,0.0245879347854,0.0246133136963,0.0245985093316,0.0245710155115,0.0246133136963,0.0246831057011,0.0247232889766,0.0248290344384,0.0248671028047,0.0250172613605,0.0250659042729,0.0249707333573,0.0249495842649,0.0249601588111,0.0251779944624,0.0253450722921,0.0252985442889,0.0251082024576,0.0249834228127,0.0250151464512,0.0250257209974,0.0252033733733,0.0252668206504,0.0254444730263,0.025334497746,0.0253154635628,0.0254973457572,0.0256326999483,0.0257490199563,0.0257469050471,0.0257490199563,0.0257130664993,0.0256855726792,0.0257553646841,0.0258611101459,0.0258293865073,0.0258018926873,0.0258632250551,0.0260324177941,0.0260620265234,0.0261275887097,0.0261762316222,0.026326390178,0.0262798621747,0.0262269894438,0.0263919523643,0.0264025269105,0.0265653749217,0.0265399960109,0.0266711203835,0.0266901545667,0.0268381982132,0.0268255087578,0.0267726360269,0.0268064745747,0.026719763296,0.0267409123883,0.0267176483867,0.0268064745747,0.026683809839,0.0266605458373,0.026630937108,0.0267409123883,0.0268551174871,0.0269354840381,0.0270031611337,0.0271829284188,0.0271681240541,0.0272950186083,0.0274874753489,0.0276968513633,0.0277666433681,0.0277624135496,0.0278829633761,0.0278237459175,0.0278448950099,0.027783562642,0.0279316062886,0.0280352368412,0.0279696746548,0.0280817648444,0.0281579015769,0.0282340383094,0.0282382681279,0.0284793677809,0.028623181609,0.0286823990676,0.0285449299672,0.0285237808748,0.0287204674339,0.0288050638033,0.0288389023511,0.0289319583575,0.0289108092652,0.0289467627222,0.0290715423671
[…]

### 2017-12-05 SLX-15090/peaks/SLX-15090.D701_D503.HMTVKBBXX.s_7.bwa.homo_sapiens_model.r

@@ -1,10 +1,10 @@
 # R script for Peak Model
 #  -- generated by MACS
-p <- c(0.0463953235843,0.0540025066093,0.0617431489855,0.0697507100644,0.0770303110452,0.0770667090501,0.0774792197724,0.0777097404701,0.077746138475,0.0776126791237,0.0778431998214,0.0781707818655,0.0782435778754,0.077746138475,0.0783042412169,0.0786439559293,0.0784740985731,0.078267843212,0.0784498332365,0.0788016806172,0.0787774152806,0.0788138132855,0.0788259459538,0.0787774152806,0.0787410172757,0.0788138132855,0.0791049973247,0.079274854681,0.0789351399685,0.0790443339832,0.0793597833591,0.0793597833591,0.0793233853542,0.0796994980715,0.0793112526859,0.0789715379734,0.0787410172757,0.0785954252561,0.079274854681,0.0784740985731,0.0785468945829,0.0780130571776,0.0780979858557,0.078753149944,0.0788866092953,0.0786803539342,0.0786196905927,0.0785711599195,0.0781222511923,0.078110118524,0.0778553324897,0.0781950472021,0.077746138475,0.0773700257577,0.0778553324897,0.0772972297478,0.0776126791237,0.0780130571776,0.0778431998214,0.0776369444603,0.0781222511923,0.078110118524,0.0789108746319,0.0788259459538,0.0782071798704,0.0784740985731,0.0780130571776,0.0780737205191,0.078389169895,0.0784134352316,0.0779645265044,0.0780979858557,0.0775277504456,0.0772850970795,0.0775398831139,0.0779766591727,0.0779766591727,0.0778553324897,0.0774549544358,0.077988791841,0.0776733424652,0.0774792197724,0.0772972297478,0.0772486990746,0.0777825364799,0.0772608317429,0.0767027290011,0.0762902182788,0.0768119230158,0.0767876576792,0.0768240556841,0.0767027290011,0.0771637703965,0.0772972297478,0.0771516377282,0.0771516377282,0.0777704038116,0.0776733424652,0.0774185564309,0.077624811792,0.0782799758803,0.0784134352316,0.077867465158,0.0778917304946,0.0776854751335,0.0783042412169,0.0781586491972,0.0791899260029,0.0787774152806,0.0784498332365,0.0776854751335,0.0777704038116,0.0782557105437,0.0782799758803,0.0787046192708,0.078753149944,0.0791413953297,0.0783527718901,0
[…]

### 2017-12-05 SLX-15090/peaks/SLX-15090.D701_D504.HMTVKBBXX.s_7.bwa.homo_sapiens_model.r

@@ -1,10 +1,10 @@
 # R script for Peak Model
 #  -- generated by MACS
-p <- c(0.01499553949,0.0175009996251,0.0200250704204,0.0224514352495,0.0249452637219,0.0248824527436,0.0250034220352,0.0249755060448,0.0249638743821,0.0250104010327,0.0248964107388,0.0249057160689,0.0250173800303,0.0251778969749,0.0252151182954,0.0253686562424,0.0254477515484,0.0254989308641,0.0255477838472,0.0257362167822,0.0258920810617,0.0258595124063,0.0258176384207,0.0258990600593,0.025994439693,0.0259479130424,0.0259199970521,0.0258781230665,0.0257873960978,0.0257152797894,0.0258269437509,0.0259223233846,0.025994439693,0.0259386077123,0.02595489204,0.0259735027003,0.0260363136786,0.0261851989605,0.0263177999147,0.0263340842425,0.026375958228,0.0265411278376,0.0267109501124,0.0267458451003,0.0269459096979,0.02710410031,0.0272646172546,0.0273506915582,0.0273134702377,0.0273483652256,0.0273437125606,0.0273041649076,0.0273088175726,0.0273646495534,0.0271599322907,0.0272715962521,0.0272041326088,0.0272273959341,0.0273855865461,0.0273181229028,0.0274739871823,0.0275461034907,0.0275019031726,0.0275949564739,0.0277368627582,0.0276577674522,0.0277461680883,0.0277438417558,0.0277926947389,0.0279299483582,0.0278694637124,0.0278462003871,0.027955538016,0.0279136640305,0.0280672019775,0.0281928239341,0.028183518604,0.0284789628353,0.0284394151823,0.028416151857,0.028569689804,0.0285534054763,0.0285208368209,0.0286162164546,0.0285789951341,0.0286883327631,0.0287418384113,0.0287837123968,0.028990755992,0.0292326945752,0.0293141162137,0.0296235184402,0.0297119190764,0.0298049723776,0.0298026460451,0.0298072987101,0.0300562162909,0.0298631306908,0.0297491403969,0.0297212244065,0.0297142454089,0.0296584134282,0.0296398027679,0.0297654247246,0.0299468786619,0.029872436021,0.0299585103246,0.0301981225752,0.0303214181993,0.0303190918668,0.0304656508162,0.0305796411102,0.0307075893993,0.0307517897174,0.0308820643391,0.0310751499391,0.031030949621,0.0310937605994,0.0312217088885,0.0313915311632,0.031
[…]

### 2017-12-05 SLX-15090/peaks/SLX-15090.D702_D501.HMTVKBBXX.s_7.bwa.homo_sapiens_model.r

@@ -1,10 +1,10 @@
 # R script for Peak Model
 #  -- generated by MACS
-p <- c(0.012571747163,0.0146461385799,0.0167162791947,0.0188820628564,0.0210839783359,0.0211201101536,0.02109247994,0.0212518850182,0.0212391326119,0.0211859975859,0.0213050200442,0.0214304187057,0.0215090585443,0.0215069331432,0.0214559235182,0.0216004507891,0.0215898237838,0.0214729267265,0.0214771775286,0.0215749459766,0.0215154347474,0.0214750521276,0.0215111839453,0.0215281871536,0.0215643189713,0.0215536919661,0.0217215986484,0.0218193670964,0.0218916307318,0.0218469973099,0.0219660197683,0.0220871676277,0.0221721836693,0.0223337141485,0.0223719713673,0.0223974761798,0.022408103185,0.0223422157527,0.0224633636121,0.022559006659,0.0225143732371,0.0225462542527,0.022612141685,0.0226142670861,0.0225313764454,0.0226291448933,0.0226950323256,0.0228268071902,0.0228076785809,0.0228820676173,0.0228905692215,0.0228948200236,0.0229564566538,0.022862939008,0.0229692090601,0.0229479550496,0.0228799422163,0.0229373280444,0.0230223440861,0.0229458296486,0.0231987523726,0.0232412603934,0.0231222379351,0.023268890607,0.0232455111955,0.0233581574507,0.0233921638674,0.0234261702841,0.0234665529039,0.0236620897997,0.023526064133,0.0235685721539,0.0236790930081,0.0236982216174,0.0238321218831,0.023827871081,0.0239426427373,0.0240977970133,0.0240340349821,0.0240234079769,0.0240127809716,0.0239256395289,0.0239468935393,0.0240510381904,0.0240637905966,0.0242125686696,0.0243592213415,0.0242593274925,0.0243294657269,0.024384726154,0.0244782437998,0.0246185202686,0.0248119317634,0.0247757999457,0.0250095940603,0.0249288288207,0.0248544397842,0.0250350988728,0.0251179895134,0.025196629352,0.0253028994041,0.025296523201,0.0254155456593,0.0255345681177,0.0255728253364,0.0256982239979,0.0257789892375,0.0258151210552,0.025957522925,0.0259532721229,0.0258831338885,0.0260042817479,0.0260871723886,0.0261233042063,0.0261658122271,0.0261488090188,0.0263337189094,0.0264187349511,0.0264102333469,0.0264676191751,0.0
[…]

### 2017-12-05 SLX-15090/peaks/SLX-15090.D702_D502.HMTVKBBXX.s_7.bwa.homo_sapiens_model.r

@@ -1,10 +1,10 @@
 # R script for Peak Model
 #  -- generated by MACS
-p <- c(0.0144360474646,0.0168767045936,0.0194491977726,0.0219118275766,0.0243305120306,0.0245265758997,0.0244640382863,0.0245417877516,0.0246888356534,0.0248257423206,0.0246634825669,0.0246465805092,0.0245586898093,0.024698976888,0.0248781386994,0.0249457469302,0.0250792731858,0.0250116649551,0.0250387082474,0.0249220840494,0.0251367401819,0.0252668860261,0.0253463256972,0.0253530865202,0.025376749401,0.025493373599,0.0255085854509,0.0255491503893,0.0254815421586,0.0254544988663,0.0254697107182,0.0254849225702,0.0254730911298,0.0254578792779,0.0253564669318,0.0253682983722,0.0254308359856,0.0254883029817,0.0255998565624,0.0258009910488,0.0258331049584,0.0258466266045,0.0258297245468,0.0259057838064,0.0260274786217,0.0259835332717,0.0260579023255,0.026010576564,0.0259936745063,0.0260004353294,0.0260984672639,0.0262573466061,0.0264162259483,0.026404394508,0.0263503079234,0.0264804537675,0.0263807316272,0.0265125676771,0.0265852465251,0.0266900392827,0.0265784857021,0.0265041166483,0.0265429913809,0.0266291918751,0.0267677887481,0.0267711691596,0.0269351191191,0.0269249778845,0.0268472284192,0.0269199072672,0.026899624798,0.0270145587902,0.0268861031518,0.0267982124519,0.0267982124519,0.0268522990365,0.0267593377192,0.0268539892422,0.0269469505595,0.0269030052095,0.0269570917941,0.0268844129461,0.0270247000248,0.0269486407653,0.0270196294075,0.0270179392017,0.027205552042,0.0273576705611,0.0274404906437,0.027555424636,0.0278495204396,0.0280033291645,0.028145306449,0.0283870058738,0.0282670012643,0.0283853156681,0.0282686914701,0.0282027734451,0.0280878394529,0.0281300945971,0.028094600276,0.0280540353375,0.0280557255433,0.0280574157491,0.0283329192893,0.0283532017585,0.0285881403602,0.0287301176447,0.0289768876869,0.0291306964118,0.0291611201156,0.0292811247251,0.0293453525443,0.0295769107345,0.029624236496,0.0296529699941,0.0297695941921,0.0298710065382,0.0298963596247,0.0298862183901,0.
[…]

### 2017-12-05 SLX-15090/peaks/SLX-15090.D702_D503.HMTVKBBXX.s_7.bwa.homo_sapiens_model.r

@@ -1,10 +1,10 @@
 # R script for Peak Model
 #  -- generated by MACS
-p <- c(0.0479003137643,0.0554040334429,0.0630752001437,0.070516127189,0.0780198468676,0.0778419344066,0.0776012293123,0.0771721463181,0.077098888246,0.0765442199852,0.0763349112076,0.0766384089351,0.0769000449072,0.0772349389514,0.0776430910678,0.0783024137174,0.0783128791563,0.0785849805672,0.078961736367,0.0791082525114,0.0788466165393,0.0786582386394,0.0783233445952,0.0782186902064,0.0781558975731,0.0773919205347,0.0773081970236,0.0771093536848,0.0765023582297,0.0765965471796,0.0771721463181,0.0773709896569,0.0775175058013,0.077360524218,0.0769314412238,0.0773291279014,0.0775593675568,0.0776116947512,0.0777268145789,0.0773500587791,0.0772558698292,0.0772558698292,0.0773186624625,0.0777896072122,0.0782710174008,0.0783652063507,0.0781454321342,0.0783338100341,0.0785431188117,0.0786268423228,0.0784384644229,0.0779989159898,0.078072174062,0.0780303123065,0.0779465887954,0.0780826395009,0.0781977593286,0.0787314967116,0.0785849805672,0.0787628930282,0.0787314967116,0.0790035981225,0.0789512709281,0.0795059391889,0.0796315244555,0.0797257134054,0.0795373355055,0.0789303400504,0.0790245290003,0.0791605797058,0.0794536119945,0.0795687318222,0.0799978148164,0.0785640496895,0.0781035703787,0.0778837961621,0.0780826395009,0.0783128791563,0.0784907916173,0.0781245012564,0.0782082247675,0.078961736367,0.078700100395,0.0796315244555,0.0798094369165,0.0798408332331,0.0800606074497,0.0797885060387,0.079673386211,0.0797989714776,0.0792129069002,0.0784175335451,0.0788256856615,0.0786582386394,0.078250086523,0.0783861372285,0.0782814828397,0.0779989159898,0.0774128514124,0.0776535565067,0.0778314689677,0.0780303123065,0.0774651786068,0.0771721463181,0.0774965749235,0.0767849250795,0.0761046715521,0.0766174780574,0.077004699296,0.0776744873845,0.0775070403624,0.0776744873845,0.0776430910678,0.0775175058013,0.0776116947512,0.0778419344066,0.0785745151284,0.07857451
[…]

### 2017-12-05 SLX-15090/peaks/SLX-15090.D702_D504.HMTVKBBXX.s_7.bwa.homo_sapiens_model.r

@@ -1,10 +1,10 @@
 # R script for Peak Model
 #  -- generated by MACS
-p <- c(0.0126599337853,0.01481094361,0.0169454072052,0.0190840073578,0.021288792428,0.0214315036568,0.0214480498862,0.021404616034,0.021526644476,0.0215225079186,0.0216693557047,0.0217748379173,0.021789315868,0.0217851793107,0.0217148578356,0.0215514638201,0.021603170787,0.0216383315246,0.021654877754,0.021789315868,0.0219072077526,0.0218596373431,0.0218927298019,0.0220685334895,0.0220995576696,0.0221223087351,0.022283634472,0.0222939758653,0.022225722669,0.0222443371771,0.0223084538161,0.0223456828323,0.0225628520934,0.022540101028,0.0226476515192,0.0227283143876,0.0227034950435,0.0228172503708,0.0229516884848,0.0230488975826,0.0228979132392,0.0228979132392,0.0227882944693,0.0228069089774,0.0228441379936,0.0228069089774,0.0228358648789,0.0227117681582,0.0227096998795,0.0225380327493,0.0226021493883,0.0225690569295,0.0225566472574,0.0225959445523,0.0225835348802,0.0226972902075,0.0227469288957,0.0228751621737,0.0228689573377,0.0230406244679,0.0231336970084,0.0232143598768,0.0232764082372,0.0233198420894,0.0233405248762,0.0233363883188,0.0234232560233,0.0235990597109,0.023758317169,0.0237645220051,0.0237004053661,0.0238451848735,0.024002374053,0.0240313299545,0.0241161293803,0.0241823142979,0.024136812167,0.0240768320854,0.0241057879869,0.0242216115928,0.0243519131495,0.024322957248,0.0243002061826,0.0243829373297,0.0244201663459,0.0244636001981,0.0244842829849,0.0244987609356,0.0245049657716,0.0243850056083,0.0242753868384,0.0242940013465,0.0243415717561,0.0242629771664,0.0243581179855,0.0244036201164,0.0246559501151,0.0246290624923,0.024765568885,0.0249599870807,0.0251337224896,0.0252061122433,0.0252702288823,0.0254315546192,0.0254998078155,0.0256011534707,0.0256735432245,0.0257480012569,0.0257335233061,0.0257252501914,0.0258514151907,0.0259920581408,0.0259858533048,0.0260003312555,0.0259465560099,0.0260665161732,0.0261823397791,0.0264057138763,0.0264181235484,0.0265339471543,0.0264739
[…]

### 2017-12-05 SLX-15090/peaks/SLX-15090.D703_D501.HMTVKBBXX.s_7.bwa.homo_sapiens_model.r

@@ -1,10 +1,10 @@
 # R script for Peak Model
 #  -- generated by MACS
-p <- c(0.013967379942,0.0163178171676,0.0185446698962,0.0208064993693,0.0232222265179,0.0233947784571,0.0235253583029,0.0235370172177,0.0234763908607,0.0235300218689,0.023646611017,0.0236279567533,0.0238588032665,0.0240173645079,0.0240220280738,0.0238541397006,0.0237468776843,0.0237422141184,0.0238424807858,0.0238075040413,0.0238238265221,0.0238727939643,0.0236699288466,0.0236862513273,0.0237212280717,0.0238844528791,0.0240080373761,0.0240663319501,0.0241269583071,0.024131621873,0.0241199629582,0.0241386172219,0.0244137676114,0.024439417224,0.0244440807899,0.0243135009441,0.0244300900922,0.0243578048203,0.0242971784633,0.0243951133477,0.0246049738143,0.024684254435,0.0246912497839,0.0248194978468,0.0248614699401,0.0251552745933,0.0252788590903,0.0254397521147,0.025479392425,0.025479392425,0.0254071071532,0.0253558079281,0.0253091722688,0.0251925831207,0.0251832559889,0.0251925831207,0.0251692652911,0.0252858544392,0.025418766068,0.0253511443621,0.0252928497881,0.0254257614169,0.0255027102547,0.0255143691695,0.0256566079301,0.0257592063805,0.0256869211086,0.0256239629687,0.0255190327354,0.025668266845,0.0257918513419,0.0258897862263,0.0259247629708,0.0259364218856,0.0260040435915,0.0258384870012,0.0258011784738,0.0257638699464,0.0258151691716,0.0257615381634,0.0257918513419,0.025847814133,0.0258594730478,0.0262605397173,0.0263188342913,0.0263584746017,0.0265916528979,0.026768868403,0.0268224994111,0.0269484156911,0.0270370234436,0.0270183691799,0.0270626730562,0.0269111071637,0.026827162977,0.027076663754,0.0269717335207,0.0269833924355,0.0269787288696,0.0270207009629,0.0270976498006,0.0271116404984,0.0272375567783,0.0271815939872,0.027379795539,0.0273028467013,0.0274614079427,0.0276642730604,0.0279254327521,0.0281469521335,0.0282588777157,0.0283941211275,0.0285316963222,0.0286016498111,0.0285596777177,0.028606313377,0.0287742017502,0.0288418234561,0.0288791319835,0.0289420901235,0.02
[…]

### 2017-12-05 SLX-15090/peaks/SLX-15090.D703_D502.HMTVKBBXX.s_7.bwa.homo_sapiens_model.r

@@ -1,10 +1,10 @@
 # R script for Peak Model
 #  -- generated by MACS
-p <- c(0.0151416223184,0.0177535454435,0.0202767008414,0.0228267555506,0.0254736477804,0.0253660505353,0.025487097436,0.025277282808,0.0252611432213,0.0253687404664,0.0254171592267,0.0252719029458,0.0252853526014,0.0252907324637,0.0252423137034,0.0253337713617,0.025339151224,0.0255543457142,0.0254225390889,0.025508616885,0.025530136334,0.0256484933037,0.0258179589647,0.0258798273807,0.0259309360721,0.025971285039,0.0260385333172,0.0260627426974,0.0263398056035,0.0262913868432,0.0264716122288,0.0265576900249,0.0264581625731,0.0264500927798,0.0264151236751,0.0265092712646,0.0267944039641,0.0268777918291,0.0267675046528,0.0269531099007,0.0269477300384,0.0267997838264,0.0268885515536,0.0268105435509,0.0269746293497,0.027001528661,0.0269477300384,0.0269638696252,0.0271360252174,0.0271414050796,0.027127955424,0.0273431499142,0.0273969485367,0.0276632517184,0.0276659416495,0.0276013833025,0.0276444222005,0.027590623578,0.0277197402721,0.0278488569662,0.0278918958643,0.0280828809743,0.0281259198724,0.027991423316,0.0279645240047,0.028201237944,0.0281581990459,0.0282550365665,0.0282146875996,0.0280586715942,0.0280667413876,0.027969903867,0.0281689587704,0.0282388969798,0.0284191223653,0.0283841532607,0.028518649817,0.0284271921587,0.0283841532607,0.0286370067867,0.0286961852715,0.0288468214146,0.0288414415524,0.029064705836,0.0291319541142,0.0292261017037,0.029465505574,0.0297371886179,0.0298044368961,0.029758708067,0.0298985844856,0.0300196313864,0.0301030192513,0.0301083991136,0.0300357709731,0.0299201039346,0.0296403510973,0.0296833899954,0.0298447858631,0.0299012744167,0.0300061817307,0.0300088716619,0.0299389334525,0.0299227938658,0.0301998567719,0.0305011290582,0.0308723395539,0.0307647423088,0.0307297732041,0.0307297732041,0.0308212308625,0.0310418052149,0.0312731392919,0.0314937136444,0.0316201404074,0.0316981484101,0.0316309001319,0.0320693589058,0.0322926231894,0.032642314236,0.0328010
[…]

### 2017-12-05 Manuscript_Figures.Rmd
Reanalysis of H4K12ac data
Output for GEO

@@
 ```{r,results='hide', message=FALSE, warning=FALSE}
-filename<-"Rdata/066_SLX-15090_dba_human_ER_CTCF.rda"
+filename<-"Rdata/066_SLX-15090_dba_human_ER_CTCF_merged.rda"
 load(filename)
@@
 dba.plotMA(jg.dba_analysis,bFlip=TRUE,th=0.05)
+
+write.csv(dba.report(jg.dba_analysis,th=1),file="csv/H4_normalised_results.csv")
 
@@
 
+write.csv(dba.report(jg.dba_analysis,th=1),file="csv/E2_normalised_results.csv")
+
 ```

### 2017-12-05 052_ChIPQC.R
update ChIP-QC

@@ -8,8 +8,8 @@ qc.samplesheets=c(
     "samplesheet/samplesheet_SLX12998.csv",
     "samplesheet/samplesheet_SLX14229.csv",
     "samplesheet/samplesheet_SLX14438_hs_DBA.csv",
-    "samplesheet/samplesheet_SLX15090_H4K12ac.csv",
-    "samplesheet/samplesheet_SLX15091_ER.csv")
+    "samplesheet/samplesheet_SLX15090.csv",
+    "samplesheet/samplesheet_SLX15091.csv")
 
 #library(BiocParallel)
 #register(MulticoreParam())

### 2017-12-05 069_SLX14438_Stability.R
Updated Figure

@@ -58,11 +58,11 @@ plot( rowMeans(jg.coefficient)/jg.coefficient_baseline*100-100,ylim=c(-2,+2),
       type="l",
       xlab="Percent of CTCF sites",
       ylab="Percent Error in Normalisation Coefficient",
-      main="Stablity of Normalisation Coefficient")
-lines( apply(t(jg.coefficient), 2, max) /jg.coefficient_baseline*100-100,col="grey",lty=2) #max=1.956436
-lines( apply(t(jg.coefficient), 2, min)/jg.coefficient_baseline*100-100, col="grey",lty=2) #min=-1.970216
-lines( apply(t(jg.coefficient), 2, quantile)['25%',]/jg.coefficient_baseline*100-100 ,col="grey")#min=-0.4968219
-lines( apply(t(jg.coefficient), 2, quantile)['75%',]/jg.coefficient_baseline*100-100, col="grey")#max=-0.5254374
+      main="Stablity of Normalisation Coefficient",lwd=2)
+lines( apply(t(jg.coefficient), 2, max) /jg.coefficient_baseline*100-100,col="black",lty=2,lwd=2) #max=1.956436
+lines( apply(t(jg.coefficient), 2, min)/jg.coefficient_baseline*100-100, col="black",lty=2,lwd=2) #min=-1.970216
+lines( apply(t(jg.coefficient), 2, quantile)['25%',]/jg.coefficient_baseline*100-100 ,col="black", lwd=2)#min=-0.4968219
+lines( apply(t(jg.coefficient), 2, quantile)['75%',]/jg.coefficient_baseline*100-100, col="black",lwd=2)#max=-0.5254374
 legend("topleft", legend=c("Average", "25%/75% quantile","Max/min error"),
-       col=c("blue", "grey","grey"), lty=1:1:2, cex=0.8)
+       col=c("blue", "black","black"), lty=1:1:2, cex=0.8)
 dev.off()

### 2017-12-05 package/brundle.R

@@ -311,6 +311,14 @@ jg.normalize<-function(
     return(jg.dba)
 }
 
+jg.getCorrectionFactorDBA<-function(dbaExperiment,jg.treatedCondition,jg.untreatedCondition)
+{
+
+    untreated<-as.numeric(row.names(dbaExperiment$samples)[dbaExperiment$samples$Condition==jg.untreatedCondition])
+    treated<-as.numeric(row.names(dbaExperiment$samples)[dbaExperiment$samples$Condition==jg.treatedCondition])
+    jg.correctionfactor<-sum(as.numeric(dbaExperiment$class['Reads',c(treated)]))/sum(as.numeric(dbaExperiment$class['Reads',c(untreated)]))
+    return(jg.correctionfactor)
+}
 
 #Check Functions
 checkUsage(jg.countAlignedMReads)

### 2018-01-13 Manuscript_Figures.Rmd
Updated Figure
Update Markdown to reflect new Brundle function
Reworked Rmd for Manuscript

@@
 
-## Comparison of simple normalisation methods
-
-```{r results='hide', message=FALSE, warning=FALSE}
-library(DiffBind)
-
-```
-
-```{r, fig.cap="Comparison of simple normalisation strategies employed.
-MA plots showing the changes in ER binding after 48 hours treatment with 100 nM fulvestrant.
-Three simple normalisation methods were applied to this data and compared to the raw count data.
-(A) Raw counts.
-(B) Reads Per Million (RPM) reads in peaks.
-(C) RPM aligned reads.
-(D) RPM total reads.
-Note that the highlighted peaks remain above zero under all three standard normalisations."}
-
-par(mfrow=c(2,2))
-
-hsconsensus<-readRDS("Rdata/015_SLX-12998_hsconsensus.rds")
-mmconsensus<-readRDS("Rdata/015_SLX-12998_mmconsensus.rds")
-hscounts<-hsconsensus[,-c(1:3)]
-mmcounts<-mmconsensus[,-c(1:3)]
-### MA RPM in peaks
-hsrpm<-apply(hscounts,2,function(x){
-  1E6*x/sum(x)
-})
-mmrpm<-apply(mmcounts,2,function(x){
-  1E6*x/sum(x)
-})
-M<-apply(hsrpm,1,function(x){
-  fulvestrant<-mean(x[c(1,2,4,5)])
-  untreated<-mean(x[c(3,6,7,8)])
-  fc<-mean(fulvestrant)/mean(untreated)
-  log2fc<-log2(fc)
-  return(log2fc)
-})
-A<-apply(hsrpm,1,function(x){
-  return(log10(sum(x)))
-})
-
-plot(A,M,cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5,pch=20,ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"),main="RPM reads in peaks",ylim=c(-6.25,4.25))
-abline(h=0)
-
-load("Rdata/012_SLX-12998_aligned.rda")
-aligned<-aligned[grep("SLX-12998.D",names(aligned))]
-aligned<-sapply(aligned,sum)/1E6
-aligned<-aligned[c(4,7,8,9,2,3,6,5)]
-
-hsrpm<-hscounts
-for(i in 1:length(hsrpm)){
-  hsrpm[i]<-hscounts[i]/aligned[i]
-}
-mmrpm<-mmcounts
-for(i in 1:length(mmrpm)){
-  mmrpm[i]<-mmcounts[i]/aligned[i]
-}
-M<-apply(hsrpm,1,function(x){
-  fulvestrant<-mean(x[c(1,2,4,5)])
-  untreated<-mean(x[c(3,6,7,8)])
-  fc<-mean(fulvestrant)/mean(untreated)
-  log2fc<-log2(fc)
-  return(log2fc)
-})
-A<-apply(hsrpm,1,function(x){
-  return(log10(sum(x)))
-})
-
-plot(A,M,cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5,cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5,pch=20,ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"),main="RPM aligned reads",ylim=c(-6.25,4.25))
-abline(h=0)
-
-load("Rdata/012_SLX-12998_nrreads.rda")
-nrreads<-nrreads[grep("SLX-12998.D",names(nrreads))]
-nrreads<-nrreads/1E6
-nrreads<-nrreads[c(4,7,8,9,2,3,6,5)]
-hsrpm<-hscounts
-for(i in 1:length(hsrpm)){
-  hsrpm[i]<-hscounts[i]/nrreads[i]
-}
-mmrpm<-mmcounts
-for(i in 1:length(mmrpm)){
-  mmrpm[i]<-mmcounts[i]/nrreads[i]
-}
-M<-apply(hsrpm,1,function(x){
-  fulvestrant<-mean(x[c(1,2,4,5)])
-  untreated<-mean(x[c(3,6,7,8)])
-  fc<-mean(fulvestrant)/mean(untreated)
-  log2fc<-log2(fc)
-  return(log2fc)
-})
-A<-apply(hsrpm,1,function(x){
-  return(log10(sum(x)))
-})
-
-plot(A,M,cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5,pch=20,ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"),main="RPM total reads",ylim=c(-6.25,4.25))
-abline(h=0)
-
-### MA Counts human
-# a is treated, b is not treated
-M<-apply(hscounts,1,function(x){
-  fulvestrant<-mean(x[c(1,2,4,5)])
-  untreated<-mean(x[c(3,6,7,8)])
-  fc<-mean(fulvestrant)/mean(untreated)
-  log2fc<-log2(fc)
-  return(log2fc)
-})
-A<-apply(hscounts,1,function(x){
-  return(log10(sum(x)))
-})
-
-par(mar=c(5.1,5.1,4.1,2.1))
-plot(A,M,cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5,pch=20,ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"),main="Raw counts in peaks", ylim=c(-6.25,4.25))
-
-abline(h=0)
-
-```
-
-## MA Plot of H2Av normalization
-```{r, fig.cap="MA plots showing ER binding before and after treatment with fulvestrant including matched Dm H2Av spike-in control.} (A) Reads corrected to total aligned reads showed the same off-centre peak density as observed in Figure 1.
+## MA plot of H2Av normalization
+```{r H2AvNormalisationPlots, fig.cap="MA plots showing ER binding before and after treatment with fulvestrant including matched Dm H2Av spike-in control.} (A) Reads corrected to total aligned reads showed the same off-centre peak density as observed in Figure 1.
 Putative unchanged ER binding sites are within the red triangle.
@@
 
-Figure 3 can be viewed interactively on the [USCS track](http://genome.ucsc.edu/cgi-bin/hgTracks?hgS_doOtherUser=submit&hgS_otherUserName=andrewholding&hgS_otherUserSessionName=Brundle).
-
-## MA Plot of CTCF Peaks
-
-```{r, results='hide', message=FALSE, warning=FALSE, fig.cap="Comparison of the control regions used to normalise ER analysis before and after treatment.
+Figure 2 can be viewed interactively on the [USCS track](http://genome.ucsc.edu/cgi-bin/hgTracks?hgS_doOtherUser=submit&hgS_otherUserName=andrewholding&hgS_otherUserSessionName=Brundle).
+
+##H2av with DiffBind
+
+```{r H2avDiffBind, results='hide', message=FALSE, warning=FALSE,   out.width = '49%', out.extra='style="float:left"'}
+jg.controlSampleSheet     <- "samplesheet/samplesheet_SLX8047_dm.csv"
+jg.experimentSampleSheet  <- "samplesheet/samplesheet_SLX8047_hs.csv"
+filename<-"Rdata/029_SLX-8047_dba_human_drosophila.rda"
+
+load(filename)
+
+source('package/brundle.R')
+
+## Extract Peak set from DiffBind
+jg.experimentPeakset <- jg.dbaGetPeakset(dbaExperiment)
+jg.controlPeakset    <- jg.dbaGetPeakset(dbaControl)
+
+#Convert Peakset to DeSeq Workflow
+jg.controlPeaksetDeSeq<-jg.convertPeakset(jg.controlPeakset)
+
+#Establish size factors directly from Control data
+jg.controlSizeFactors = estimateSizeFactorsForMatrix(jg.controlPeaksetDeSeq)
+
+#Get conditions dataframe for DeSeq
+jg.conditions <- read.csv(file=jg.controlSampleSheet, header=TRUE, sep=",")['Condition']
+
+#Run DeSeq on control
+jg.controlDeSeq<-jg.runDeSeq(jg.controlPeaksetDeSeq, jg.conditions,jg.controlSizeFactors)
+jg.controlResultsDeseq   = results(jg.controlDeSeq)
+
+#Repeat for experimental conditions
+
+#Convert experiment Peakset to DeSeq Workflow
+jg.experimentPeaksetDeSeq<-jg.convertPeakset(jg.experimentPeakset)
+
+#Run DeSeq on experiment
+jg.experimentDeSeq<-jg.runDeSeq(jg.experimentPeaksetDeSeq, jg.conditions,jg.controlSizeFactors)
+jg.experimentResultsDeseq   = results(jg.experimentDeSeq)
+
+jg.experimentDeSeqInternal<-jg.runDeSeq(jg.experimentPeaksetDeSeq, jg.conditions, NULL)
+jg.experimentResultsDeseqInternal   = results(jg.experimentDeSeqInternal)
+
+dbaExperimentCorrected<-jg.correctDBASizeFactors(dbaExperiment,
+                                                 jg.controlSizeFactors
+                                                 )
+
+dbaExperimentAnalysis<-dba.analyze(dbaExperiment)
+  dba.plotMA(dbaExperimentAnalysis)
+
+```
+
+```{r H2avDiffBindNormalised, results='hide', message=FALSE, warning=FALSE, out.width = '49%', out.extra='style="float:left"'}
+  dbaExperimentAnalysisCorrected<-dba.analyze(dbaExperimentCorrected)
+  dba.plotMA(dbaExperimentAnalysisCorrected)
+```
+
+Comparison of DiffBind output before and after applying the corrected size factors from our pipeline generated from Drosophila spike-in control.
+(A) Analysis of ER binding before and after treatment with fulvestrant demonstrates that DiffBind’s default normalisation strategy is more effective than the DESeq2 default, but demonstrates a bias between samples.
+(B) Applying the correct size factors from our DESeq2 pipeline reduces the bias in the analysis (Data: SLX-8047).
+
+```{r DeSeqCTCFControlPeaks, results='hide', message=FALSE, warning=FALSE, fig.cap="Comparison of the control regions used to normalise ER analysis before and after treatment.
 Dots highlighted in red are significant (FDR = 0.01).
@@
 
-source('package/brundle.R')
+#This is head as otherwise next figure doesn't work.
+
 filename<-"Rdata/040_SLX-14438_dba_human_ER_CTCF.rda"
@@
 
-jg.correctionFactor<-jg.getCorrectionFactor(jg.experimentSampleSheet,
-                                            jg.treatedNames,
-                                            jg.untreatedNames
-)
+#jg.correctionFactor<-jg.getCorrectionFactor(jg.experimentSampleSheet,
+#                                            jg.treatedNames,
+#                                            jg.untreatedNames
+#)
+
+#Manuscript figure uses the above function, but to enable running of the code without the bam files we 
+#estimate the Factor from diffbind object.
+
+untreated <-
+            as.numeric(row.names(dbaExperiment$samples)[dbaExperiment$samples$Condition ==
+                                                            jg.untreatedCondition])
+        treated <-
+            as.numeric(row.names(dbaExperiment$samples)[dbaExperiment$samples$Condition ==
+                                                            jg.treatedCondition])
+        jg.correctionFactor <-
+            sum(as.numeric(dbaExperiment$class['Reads', c(treated)])) / sum(as.numeric(dbaExperiment$class['Reads', c(untreated)]))
 
@@
 
-jg.plotDeSeq(jg.controlResultsDeseq,
-             p=0.01, 
-             title.main="Fold-change in CTCF binding",
-             flip=T
-)
-
-```
-
-##H2av With DiffBind
-
-```{r results='hide', message=FALSE, warning=FALSE,   out.width = '49%', out.extra='style="float:left"'}
-jg.controlSampleSheet     <- "samplesheet/samplesheet_SLX8047_dm.csv"
-jg.experimentSampleSheet  <- "samplesheet/samplesheet_SLX8047_hs.csv"
-filename<-"Rdata/029_SLX-8047_dba_human_drosophila.rda"
-
-load(filename)
-
-## Extract Peak set from DiffBind
-jg.experimentPeakset <- jg.dbaGetPeakset(dbaExperiment)
-jg.controlPeakset    <- jg.dbaGetPeakset(dbaControl)
-
-#Convert Peakset to DeSeq Workflow
-jg.controlPeaksetDeSeq<-jg.convertPeakset(jg.controlPeakset)
-
-#Establish size factors directly from Control data
-jg.controlSizeFactors = estimateSizeFactorsForMatrix(jg.controlPeaksetDeSeq)
-
-#Get conditions dataframe for DeSeq
-jg.conditions <- read.csv(file=jg.controlSampleSheet, header=TRUE, sep=",")['Condition']
-
-#Run DeSeq on control
-jg.controlDeSeq<-jg.runDeSeq(jg.controlPeaksetDeSeq, jg.conditions,jg.controlSizeFactors)
-jg.controlResultsDeseq   = results(jg.controlDeSeq)
-
-#Repeat for experimental conditions
-
-#Convert experiment Peakset to DeSeq Workflow
-jg.experimentPeaksetDeSeq<-jg.convertPeakset(jg.experimentPeakset)
-
-#Run DeSeq on experiment
-jg.experimentDeSeq<-jg.runDeSeq(jg.experimentPeaksetDeSeq, jg.conditions,jg.controlSizeFactors)
-jg.experimentResultsDeseq   = results(jg.experimentDeSeq)
-
-jg.experimentDeSeqInternal<-jg.runDeSeq(jg.experimentPeaksetDeSeq, jg.conditions, NULL)
-jg.experimentResultsDeseqInternal   = results(jg.experimentDeSeqInternal)
-
-dbaExperimentCorrected<-jg.correctDBASizeFactors(dbaExperiment,
-                                                 jg.controlSizeFactors
-                                                 )
-
-dbaExperimentAnalysis<-dba.analyze(dbaExperiment)
-  dba.plotMA(dbaExperimentAnalysis)
-
-```
-
-```{r results='hide', message=FALSE, warning=FALSE, out.width = '49%', out.extra='style="float:left"'}
-  dbaExperimentAnalysisCorrected<-dba.analyze(dbaExperimentCorrected)
-  dba.plotMA(dbaExperimentAnalysisCorrected)
-```
-
-Comparison of DiffBind output before and after applying the corrected size factors from our pipeline generated from Drosophila spike-in control.
-(A) Analysis of ER binding before and after treatment with fulvestrant demonstrates that DiffBind’s default normalisation strategy is more effective than the DESeq2 default, but demonstrates a bias between samples.
-(B) Applying the correct size factors from our DESeq2 pipeline reduces the bias in the analysis (Data: SLX-8047).
+#Store this and use it later.
+
+CTCFbindingPlot<-jg.controlResultsDeseq
+
+```
 
@@
 
-```{r results='hide', message='false', warning='false', fig.cap="Comparison of mean counts in CTCF peaks before and after treatment.
+```{r LinearModel, results='hide', message='false', warning='false', fig.cap="Comparison of mean counts in CTCF peaks before and after treatment.
 If the samples have no systematic bias before and after treatment then the linear fit would be expected to have a gradient of 1.
@@
 
+##Normalisation factors are consistent over a wide range in number of control binding sites
+
+```{r CTCFStablility, echo=FALSE,warning=FALSE,message=FALSE,error=FALSE}
+library("Brundle")
+
+jg.controlSampleSheet     <- "samplesheet/samplesheet_SLX14438_hs_CTCF_DBA.csv"
+jg.experimentSampleSheet  <- "samplesheet/samplesheet_SLX14438_hs_ER_DBA.csv"
+jg.treatedCondition       =  "Fulvestrant"
+jg.untreatedCondition     =  "none"
+
+filename<-"Rdata/048_SLX-14438_dba_human_ER_CTCF.rda"
+load(filename)
+
+#Load Sample Ids from control sample sheet.
+jg.sampleIds <- jg.getSampleIds(jg.controlSampleSheet)
+
+## Extract Peak set from DiffBind
+jg.experimentPeakset <- jg.dbaGetPeakset(dbaExperiment)
+jg.controlPeakset    <- jg.dbaGetPeakset(dbaControl)
+
+#Get counts for each condition
+jg.controlCountsTreated<-jg.getControlCounts(jg.controlPeakset, 
+                                             jg.controlSampleSheet,
+                                             jg.treatedCondition)
+jg.controlCountsUntreated<-jg.getControlCounts(jg.controlPeakset,
+                                               jg.controlSampleSheet,
+                                               jg.untreatedCondition)
+
+#Get sample names for conditions
+jg.untreatedNames <- names(jg.controlCountsUntreated)
+jg.treatedNames   <- names(jg.controlCountsTreated)
+
+jg.coefficient_baseline<-jg.getNormalizationCoefficient(jg.controlCountsTreated,
+                                               jg.controlCountsUntreated)
+rows<-nrow(jg.controlCountsTreated)
+
+jg.coefficient<-matrix(nrow=100,ncol=100)
+
+for (n in 100:1)
+{
+    for (i in 1:100)
+    {
+        set.seed(i)
+        sampleRows<-sample(rows, rows*(n/100) )
+        jg.coefficient[n,i]<-jg.getNormalizationCoefficient(jg.controlCountsTreated[sampleRows, ],
+                               jg.controlCountsUntreated[sampleRows, ])
+    }
+}
+
+plot( rowMeans(jg.coefficient)/jg.coefficient_baseline*100-100,ylim=c(-2,+2),
+      xlim=c(100,0),
+      pch=20,col="blue",
+      type="l",
+      xlab="Percent of CTCF sites",
+      ylab="Percent Error in Normalisation Coefficient",
+      main="Stablity of Normalisation Coefficient",lwd=2)
+lines( apply(t(jg.coefficient), 2, max) /jg.coefficient_baseline*100-100,col="black",lty=2,lwd=2) #max=1.956436
+lines( apply(t(jg.coefficient), 2, min)/jg.coefficient_baseline*100-100, col="black",lty=2,lwd=2) #min=-1.970216
+lines( apply(t(jg.coefficient), 2, quantile)['25%',]/jg.coefficient_baseline*100-100 ,col="black", lwd=2)#min=-0.4968219
+lines( apply(t(jg.coefficient), 2, quantile)['75%',]/jg.coefficient_baseline*100-100, col="black",lwd=2)#max=-0.5254374
+legend("topleft", legend=c("Average", "25%/75% quantile","Max/min error"),
+       col=c("blue", "black","black"), lty=1:1:2, cex=0.8)
+
+```
+
+Stability of CTCF derived normalisation coefficient.
+Stability of the CTCF derived normalisation coefficient was analysed by sub-sampling CTCF peaks before undertaking the calculation (between 1-100% of total sites) at random.
+This analysis was repeated 100 times to model the variability of the result.
+
 ##Comparison of CTCF and H2Av normalisation methods
 
-```{r results='hide', echo=FALSE,warning=FALSE,message=FALSE,error=FALSE} 
+```{r CTCF.H2avComparision, results='hide', echo=FALSE,warning=FALSE,message=FALSE,error=FALSE} 
 jg.controlSampleSheet     <- "samplesheet/samplesheet_SLX8047_dm.csv"
@@
 
-```{r out.width = '50%', out.extra='style="float:left"'}
+```{r CTCF.H2avOverlay, out.width = '50%', out.extra='style="float:left"'}
 a + as.layer(b) 
@@
 
-```{r out.width = '50%', out.extra='style="float:left"', fig.cap="Comparison of normalisation methods using consensus peak set.
+```{r CTCF.H2avXYScatter, out.width = '50%', out.extra='style="float:left"', fig.cap="Comparison of normalisation methods using consensus peak set.
 (A) The analysis for the CTCF normalised (blue) and H2Av normalised (green) dataset using an ER consensus peak set of 10,000 peaks were formatted as an MA plot and overlaid.
@@
 
+## Comparison of simple normalisation methods
+
+```{r LoadDiffBind, results='hide', message=FALSE, warning=FALSE}
+library(DiffBind)
+
+```
+
+```{r ClassicMethods, fig.cap="Comparison of simple normalisation strategies employed.
+MA plots showing the changes in ER binding after 48 hours treatment with 100 nM fulvestrant.
+Three simple normalisation methods were applied to this data and compared to the raw count data.
+(A) Raw counts.
+(B) Reads Per Million (RPM) reads in peaks.
+(C) RPM aligned reads.
+(D) RPM total reads.
+Note that the highlighted peaks remain above zero under all three standard normalisations."}
+
+par(mfrow=c(2,2))
+
+hsconsensus<-readRDS("Rdata/015_SLX-12998_hsconsensus.rds")
+mmconsensus<-readRDS("Rdata/015_SLX-12998_mmconsensus.rds")
+hscounts<-hsconsensus[,-c(1:3)]
+mmcounts<-mmconsensus[,-c(1:3)]
+### MA RPM in peaks
+hsrpm<-apply(hscounts,2,function(x){
+  1E6*x/sum(x)
+})
+mmrpm<-apply(mmcounts,2,function(x){
+  1E6*x/sum(x)
+})
+M<-apply(hsrpm,1,function(x){
+  fulvestrant<-mean(x[c(1,2,4,5)])
+  untreated<-mean(x[c(3,6,7,8)])
+  fc<-mean(fulvestrant)/mean(untreated)
+  log2fc<-log2(fc)
+  return(log2fc)
+})
+A<-apply(hsrpm,1,function(x){
+  return(log10(sum(x)))
+})
+
+plot(A,M,cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5,pch=20,ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"),main="RPM reads in peaks",ylim=c(-6.25,4.25))
+abline(h=0)
+
+load("Rdata/012_SLX-12998_aligned.rda")
+aligned<-aligned[grep("SLX-12998.D",names(aligned))]
+aligned<-sapply(aligned,sum)/1E6
+aligned<-aligned[c(4,7,8,9,2,3,6,5)]
+
+hsrpm<-hscounts
+for(i in 1:length(hsrpm)){
+  hsrpm[i]<-hscounts[i]/aligned[i]
+}
+mmrpm<-mmcounts
+for(i in 1:length(mmrpm)){
+  mmrpm[i]<-mmcounts[i]/aligned[i]
+}
+M<-apply(hsrpm,1,function(x){
+  fulvestrant<-mean(x[c(1,2,4,5)])
+  untreated<-mean(x[c(3,6,7,8)])
+  fc<-mean(fulvestrant)/mean(untreated)
+  log2fc<-log2(fc)
+  return(log2fc)
+})
+A<-apply(hsrpm,1,function(x){
+  return(log10(sum(x)))
+})
+
+plot(A,M,cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5,cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5,pch=20,ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"),main="RPM aligned reads",ylim=c(-6.25,4.25))
+abline(h=0)
+
+load("Rdata/012_SLX-12998_nrreads.rda")
+nrreads<-nrreads[grep("SLX-12998.D",names(nrreads))]
+nrreads<-nrreads/1E6
+nrreads<-nrreads[c(4,7,8,9,2,3,6,5)]
+hsrpm<-hscounts
+for(i in 1:length(hsrpm)){
+  hsrpm[i]<-hscounts[i]/nrreads[i]
+}
+mmrpm<-mmcounts
+for(i in 1:length(mmrpm)){
+  mmrpm[i]<-mmcounts[i]/nrreads[i]
+}
+M<-apply(hsrpm,1,function(x){
+  fulvestrant<-mean(x[c(1,2,4,5)])
+  untreated<-mean(x[c(3,6,7,8)])
+  fc<-mean(fulvestrant)/mean(untreated)
+  log2fc<-log2(fc)
+  return(log2fc)
+})
+A<-apply(hsrpm,1,function(x){
+  return(log10(sum(x)))
+})
+
+plot(A,M,cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5,pch=20,ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"),main="RPM total reads",ylim=c(-6.25,4.25))
+abline(h=0)
+
+### MA Counts human
+# a is treated, b is not treated
+M<-apply(hscounts,1,function(x){
+  fulvestrant<-mean(x[c(1,2,4,5)])
+  untreated<-mean(x[c(3,6,7,8)])
+  fc<-mean(fulvestrant)/mean(untreated)
+  log2fc<-log2(fc)
+  return(log2fc)
+})
+A<-apply(hscounts,1,function(x){
+  return(log10(sum(x)))
+})
+
+par(mar=c(5.1,5.1,4.1,2.1))
+plot(A,M,cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5,pch=20,ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"),main="Raw counts in peaks", ylim=c(-6.25,4.25))
+
+abline(h=0)
+
+```
+
 ##Method comparision
 
-```{r, results='hide', message=FALSE, warning=FALSE}
+```{r AnalysisPackageComparison, results='hide', message=FALSE, warning=FALSE}
 
@@
                                                jg.controlCountsUntreated)
-jg.correctionFactor<-jg.getCorrectionFactor(jg.experimentSampleSheet,
-                                            jg.treatedNames,
-                                            jg.untreatedNames)
+
+#Publication uses the counts from the Bam file, but to save space here we get them from the DiffBind object
+
+#jg.correctionFactor<-jg.getCorrectionFactor(jg.experimentSampleSheet,
+#                                            jg.treatedNames,
+#                                            jg.untreatedNames
+#                                            )
+
+untreated <-
+            as.numeric(row.names(dbaExperiment$samples)[dbaExperiment$samples$Condition ==
+                                                            jg.untreatedCondition])
+        treated <-
+            as.numeric(row.names(dbaExperiment$samples)[dbaExperiment$samples$Condition ==
+                                                            jg.treatedCondition])
+        jg.correctionFactor <-
+            sum(as.numeric(dbaExperiment$class['Reads', c(treated)])) / sum(as.numeric(dbaExperiment$class['Reads', c(untreated)]))
 
@@
 
-```{r, fig.cap="Correlation Plots of Replicate Experiments.
+```{r ReproducbilityXYScatter, fig.cap="Correlation Plots of Replicate Experiments.
 (A) Scatter plots showing the correlation between the replicates with the lowest correlation value.
@@
 
-##Venn diagrams
-
-```{r  out.width = '25%', out.extra='style="float:left"'}
-filename<-"Rdata/040_SLX-14438_dba_human_ER_CTCF.rda"
-load(filename)
-
-dba.plotVenn(dbaControl,dbaControl$masks$Fulvestrant, label1="Rep1", label2="Rep2", label3="Rep3", main="CTCF Peaks Fulvestrant")
-dba.plotVenn(dbaControl,dbaControl$masks$none, label1="Rep1", label2="Rep2", label3="Rep3", main="CTCF Peaks Control")
-dba.plotVenn(dbaExperiment,dbaExperiment$masks$Fulvestrant, label1="Rep1", label2="Rep2", label3="Rep3", main="ER Peaks Fulvestrant")
-dba.plotVenn(dbaExperiment,dbaExperiment$masks$none,label1="Rep1", label2="Rep2", label3="Rep3", main="ER Peaks Control")
-```
-
-```{r, out.width = '25%', out.extra='style="float:left"'}
-
-filename<-"Rdata/026_SLX-8047_dba_drosophila.rda"
-load(filename)
-filename<-"Rdata/026_SLX-8047_dba_human.rda"
-load(filename)
-
-dba.plotVenn(dba_dm,dba_dm$masks$Fulvestrant, label1="Rep1", label2="Rep2", label3="Rep3",label4="Rep4", main="H2av Peaks Fulvestrant")
-dba.plotVenn(dba_dm,dba_dm$masks$none,label1="Rep1", label2="Rep2", label3="Rep3",label4="Rep4", main="H2av Peaks Control")
-dba.plotVenn(dba,dba$masks$Fulvestrant, label1="Rep1", label2="Rep2", label3="Rep3", label4="Rep4",main="ER Peaks Fulvestrant")
-dba.plotVenn(dba,dba$masks$none,label1="Rep1", label2="Rep2", label3="Rep3",label4="Rep4", main="ER Peaks Control")
-```
-
-```{r, out.width = '25%', out.extra='style="float:left"'}
-filename<-"Rdata/014_SLX-12998_dba_mouse.rda"
-load(filename)
-
-dba.plotVenn(dba,dba$masks$Fulvestrant, label1="Rep1", label2="Rep2", label3="Rep3", label4="Rep4",main="Mouse ER Peaks Fulvestrant")
-dba.plotVenn(dba,dba$masks$none,label1="Rep1", label2="Rep2", label3="Rep3",label4="Rep4", main="Mouse ER Peaks Control")
-
-filename<-"Rdata/015_SLX-12998_dba_human.rda"
-load(filename)
-
-dba.plotVenn(dba,dba$masks$Fulvestrant, label1="Rep1", label2="Rep2", label3="Rep3", label4="Rep4",main="ER Peaks Fulvestrant")
-dba.plotVenn(dba,dba$masks$none,label1="Rep1", label2="Rep2", label3="Rep3",label4="Rep4", main="ER Peaks Control")
-```
-
-Venn diagrams showing peak overlap between replicates.
-(A) Peak overlap between samples for CTCF internal control.
-(B) Peak overlap for murine chromatin spike-in samples.
-(C) Peak overlap for Drosophila chromatin spike-in.
-As expected, samples treated with fulvestrant showed fewer ER peaks and a higher proportion of peaks unique to individual samples.
-This is indicative in a loss of ER binding leading and the resulting decrease in single to noise ratio.
-As we use a statistical analysis and model the variance of the peaks between replicates, false positives in the treated condition will be removed at the point of differential binding analysis.
-
 ##MA Plots of mouse ER normalization
 
-```{r}
+```{r MAPlotMouseSpikeIn}
 
@@
 
-```{r, fig.cap="Distribution of reads for Mm chromatin spike-in normalisation strategy.
+```{r RelativeReadAligmentMm, fig.cap="Distribution of reads for Mm chromatin spike-in normalisation strategy.
 Comparison of murine chromatin between samples showed no systematic bias in the sample preparation.
@@
 
-```{r}
+```{r CTCFNormalisationPlots}
 par(mfrow=c(1,3))
@@
 
-```{r,results='hide', message=FALSE, warning=FALSE,   out.width = '33%', out.extra='style="float:left"'}
+```{r ERCTCFHeatmap,results='hide', message=FALSE, warning=FALSE,   out.width = '33%', out.extra='style="float:left"'}
 filename<-"Rdata/015_SLX-14229_dbacounts_human.rda"
@@
 
+##Comparison of control regions
+
+```{r H2avControlRegions, results='hide', message=FALSE, warning=FALSE,  out.width = '50%', out.extra='style="float:left"'}
+
+filename<-"Rdata/029_SLX-8047_dba_human_drosophila.rda"
+  load(filename)
+
+ jg.controlSampleSheet     <- "samplesheet/samplesheet_SLX8047_dm.csv"
+
+  ## Extract Peak set from DiffBind
+jg.controlPeakset    <- jg.dbaGetPeakset(dbaControl)
+
+#Convert Peakset to DeSeq Workflow
+jg.controlPeaksetDeSeq<-jg.convertPeakset(jg.controlPeakset)
+
+#Establish size factors directly from Control data
+jg.controlSizeFactors = estimateSizeFactorsForMatrix(jg.controlPeaksetDeSeq)
+
+#Get conditions dataframe for DeSeq
+jg.conditions <- read.csv(file=jg.controlSampleSheet, header=TRUE, sep=",")['Condition']
+
+#Run DeSeq on control
+#jg.controlDeSeq<-jg.runDeSeq(jg.controlPeaksetDeSeq, jg.conditions,jg.controlSizeFactors)
+#jg.controlResultsDeseq   = results(jg.controlDeSeq)
+
+#Some defaults in DeSeq changed after preparing figures, the next 3 steps overides these setting to get the same figures as in the manuscript.
+
+DESeqData<-DESeqDataSetFromMatrix(jg.controlPeaksetDeSeq, jg.conditions, ~Condition)
+jg.controlDeSeq<-DESeq(DESeqData,betaPrior = TRUE)
+jg.controlResultsDeseq   = results(jg.controlDeSeq, MLE=TRUE)
+
+ jg.plotDeSeq(jg.controlResultsDeseq,
+               p=0.01, 
+               title.main="Fold-change in H2av binding",
+               flip=T
+               )
+```
+
+```{r CTCFControlRegions, results='hide', message=FALSE, warning=FALSE,  out.width = '50%', out.extra='style="float:left"'}
+
+jg.plotDeSeq(CTCFbindingPlot,
+             p=0.01, 
+             title.main="Fold-change in CTCF binding",
+             flip=T
+)
+
+```
+
+Comparison of the control regions used to normalise ER analysis before and after treatment.
+Dots highlighted in red are significant (FDR = 0.01).
+(A) H2Av occupancy of the Drosophila genome shows no significant changes before and after treatment.
+(B) The CTCF peaks used for normalisation show no significant change in the number reads before and after treatment.
+
 ##Normalisation using DESeq2 SizeFactors
 
-```{r, results='hide', message=FALSE, warning=FALSE,  out.width = '50%', out.extra='style="float:left"'}
+```{r DeSEQDmNormalisation, results='hide', message=FALSE, warning=FALSE,  out.width = '50%', out.extra='style="float:left"'}
 
@@
 
-```{r, results='hide', message=FALSE, warning=FALSE,  out.width = '50%', out.extra='style="float:left"'}
+```{r DeSEQCTCFNormalisation, results='hide', message=FALSE, warning=FALSE,  out.width = '50%', out.extra='style="float:left"'}
 
@@
 
-##Comparison of mormalisation DiffBind plots
-
-```{r,results='hide', message=FALSE, warning=FALSE,   out.width = '32%', out.extra='style="float:left"'}
+##Comparison of normalisation DiffBind plots
+
+```{r DiffBindNormalisation,results='hide', message=FALSE, warning=FALSE,   out.width = '32%', out.extra='style="float:left"'}
 jg.controlSampleSheet     <- "samplesheet/samplesheet_SLX14438_hs_CTCF_DBA.csv"
@@
                                                jg.controlCountsUntreated)
-jg.correctionFactor<-jg.getCorrectionFactor(jg.experimentSampleSheet,
-                                            jg.treatedNames,
-                                            jg.untreatedNames)
+
+#Estiamte correction factor from Diffbind Object as Bams may not be avalible
+
+untreated <-
+            as.numeric(row.names(dbaExperiment$samples)[dbaExperiment$samples$Condition ==
+                                                            jg.untreatedCondition])
+        treated <-
+            as.numeric(row.names(dbaExperiment$samples)[dbaExperiment$samples$Condition ==
+                                                            jg.treatedCondition])
+        jg.correctionFactor <-
+            sum(as.numeric(dbaExperiment$class['Reads', c(treated)])) / sum(as.numeric(dbaExperiment$class['Reads', c(untreated)]))
+
+#jg.correctionFactor<-jg.getCorrectionFactor(jg.experimentSampleSheet,
+#                                            jg.treatedNames,
+#                                            jg.untreatedNames)
 
@@
 
-```{r results='hide', message=FALSE, warning=FALSE, fig.cap="Comparison of fold-change of ER binding after both xenogenic and cross-normalisation.
-Pearson's correlation between the two methods is 0.992 (3sf, p-value tending to 0).
-Deviation of data points from parity is a result the integer nature of read counts, nonetheless this effect is is very small as demonstrated correlation coefficient between the two datasets."}
+```{r CrossNormalisationXYScatter, results='hide', message=FALSE, warning=FALSE,  out.width = '50%', out.extra='style="float:left"'}
 filename<-"Rdata/047_SLX-14438_dba_human_ER_CTCF.rda"
@@
 
+jg.treatedNamesPF<-jg.treatedNames
+jg.untreatedNamesPF<-jg.untreatedNames
+
 #jg.dba is now normalised a list of ER peaks and their expected ratio
@@
 #Get sample names for conditions
-jg.untreatedNames <- names(jg.ERCountsTreated)
-jg.treatedNames   <- names(jg.ERCountsUntreated)
+jg.treatedNames <- names(jg.ERCountsTreated)
+jg.untreatedNames   <- names(jg.ERCountsUntreated)
 
@@
 #Get sample names for conditions
-jg.untreatedNames <- names(jg.EROnlyCountsTreated)
-jg.treatedNames   <- names(jg.EROnlyCountsUntreated)
+jg.untreatedNames <- names(jg.EROnlyCountsUntreated)
+jg.treatedNames   <- names(jg.EROnlyCountsTreated)
 
@@
 
-g.correctionFactor<-jg.getCorrectionFactor(jg.EROnlyExperimentSampleSheet,
-                                            jg.treatedNames,
-                                            jg.untreatedNames
-)
+untreated <-
+            as.numeric(row.names(dbaERonlyExperiment$samples)[dbaERonlyExperiment$samples$Condition ==
+                                                            jg.untreatedCondition])
+        treated <-
+            as.numeric(row.names(dbaERonlyExperiment$samples)[dbaERonlyExperiment$samples$Condition ==
+                                                            jg.treatedCondition])
+        jg.correctionFactor <-
+            sum(as.numeric(dbaERonlyExperiment$class['Reads', c(treated)])) / sum(as.numeric(dbaERonlyExperiment$class['Reads', c(untreated)]))
+
+#jg.correctionFactor<-jg.getCorrectionFactor(jg.EROnlyExperimentSampleSheet,
+#                                            jg.treatedNames,
+#                                            jg.untreatedNames
+#)
 
@@
 
-## Changes in H4K12ac following E2 treatment
-
-```{r,results='hide', message=FALSE, warning=FALSE}
-filename<-"Rdata/066_SLX-15090_dba_human_ER_CTCF_merged.rda"
-load(filename)
-
-#install_github("andrewholding/Brundle")
+```{r CrossNormalisationBoxPlot, message=FALSE, warning=FALSE, out.width = '50%', out.extra='style="float:left"', fig.cap="Comparison of fold-change of ER binding after both xenogenic and cross-normalisation.
+(A) Scatter plot of fold-change as established at individual sites by each method.
+Pearson's correlation between the two methods is 0.992 (3sf, p-value tending to 0).
+Deviation of data points from parity is a result the integer nature of read counts, nonetheless this effect is is very small as demonstrated correlation coefficient between the two datasets.
+(B) Box-plot showing the fold-change of ER binding before and after treatment at ∼550 ER sites proximal to CTCF binding.
+The mean and maximum fold-change is reduced at these sites by Parallel-Factor ChiP, but the effect is marginal."}
+
 library(Brundle)
 
-jg.controlSampleSheet     <- "samplesheet/samplesheet_SLX15090_CTCF.csv"
-jg.experimentSampleSheet  <- "samplesheet/samplesheet_SLX15090_H4K12ac.csv"
+erChIP<-makeGRangesFromDataFrame(jg.EROnlyExperimentPeakset,keep.extra.columns=TRUE)
+pfChIP<-makeGRangesFromDataFrame(jg.normalisedExperimentPeakset,keep.extra.columns=TRUE)
+ctcfBED<-read.delim("SLX-14438_merged/peaks/CTCF_union.bed",header = F)
+colnames(ctcfBED)<-c( "CHR" ,"START","END")
+jg.controlPeaksetWide<-ctcfBED
+jg.controlPeaksetWide[2]<-ctcfBED[2]-1000
+jg.controlPeaksetWide[3]<-ctcfBED[3]+1000
+ctcfChIP<-makeGRangesFromDataFrame(jg.controlPeaksetWide)
+
+#jg.plotMA(jg.normalisedExperimentPeakset,jg.normalisedExperimentPeakset,jg.untreatedNamesPF,jg.treatedNamesPF,1)
+#jg.plotMA(jg.EROnlyExperimentPeakset,jg.EROnlyExperimentPeakset,jg.treatedNames,jg.untreatedNames,1/jg.ERNormalisationRatio)
+
+erChIPctcfOverlap<-subsetByOverlaps(erChIP,ctcfChIP)
+pfChIPctcfOverlap<-subsetByOverlaps(pfChIP,ctcfChIP)
+
+erChIPpfOverlap<-subsetByOverlaps(erChIPctcfOverlap,pfChIPctcfOverlap)
+pfChIPerOverlap<-subsetByOverlaps(pfChIPctcfOverlap,erChIPctcfOverlap)
+
+erChIPuntreated<-rowMeans(as.matrix(elementMetadata(erChIPpfOverlap)[,jg.untreatedNames]))
+erChIPtreated<-rowMeans(as.matrix(elementMetadata(erChIPpfOverlap)[,jg.treatedNames]))
+#erChIPtreated/erChIPuntreated
+
+pfChIPuntreated<-rowMeans(as.matrix(elementMetadata(pfChIPerOverlap)[,jg.untreatedNamesPF]))
+pfChIPtreated<-rowMeans(as.matrix(elementMetadata(pfChIPerOverlap)[,jg.treatedNamesPF]))
+#pfChIPtreated/pfChIPuntreated
+
+boxplot(log(erChIPtreated/erChIPuntreated),
+        log(pfChIPtreated/pfChIPuntreated),
+        xlab="Experiment",
+        ylab="Log(Fold-Change)",
+        main=paste("Fold-Change of ER binding","Proximal to CTCF",collapse="\n"),
+        names=c(" Cross-Normalised"," Parrallel Factor"),
+        pch=20)
+
+```
+
+## Activation of ER in MCF7
+
+```{r ERNormalisedMAPlot,results='hide', message=FALSE, warning=FALSE}
+
+jg.controlSampleSheet     <- "samplesheet/samplesheet_SLX15091_CTCF.csv"
+jg.experimentSampleSheet  <- "samplesheet/samplesheet_SLX15091_ER.csv"
 jg.treatedCondition       =  "Estrogen"
@@
 
-jg.dba<-Brundle(
+#####
+#
+# Main Code
+#
+#####
+
+filename<-"Rdata/065_SLX-15091_dba_human_ER_CTCF.rda"
+load(filename)
+
+jg.experimentPeaksetNormalised<-Brundle(
             dbaExperiment,
@@
             jg.experimentSampleSheet,
-            jg.controlSampleSheet )
+            jg.controlSampleSheet,
+            jg.noBAMs = TRUE ) #As previous examples, BAMs were used in the manuscript.
+
+jg.dba <- DiffBind:::pv.resetCounts(dbaExperiment, jg.experimentPeaksetNormalised)
 
@@
 
+write.csv(dba.report(jg.dba_analysis,th=1),file="csv/E2_normalised_results.csv")
+
+```
+
+Comparison of fold-change of ER binding before and after treatment with estradiol.
+MA plot of ER binding after normalisation to CTCF binding displays a significant increase in ER binding at 45 minutes after treatment with estradiol.
+
+```{r ERChIPReproducibility,results='hide', message=FALSE, warning=FALSE}
+
+controlNames<-c("1a","3b")
+
+jg.experimentPeakset <- jg.dbaGetPeakset(dbaExperiment)
+relativeLibrarySize<-jg.getLibrarySize(dbaExperiment,
+                                       jg.experimentPeakset)
+peaksLibrarySize<-relativeLibrarySize[controlNames]
+peaks<-jg.experimentPeakset[controlNames]
+peaks[1]<-round(peaks[1]/peaksLibrarySize[1])
+peaks[2]<-round(peaks[2]/peaksLibrarySize[2])
+x<-c("Rep1","Rep3")
+names(x)<-controlNames
+x<-as.matrix(x)
+jg.plotScatter(peaks,controlNames, T,t(x),FALSE)
+Axis(side=2, labels=TRUE,ylab="Log(Counts)")
+mtext("log(Counts)", side=2, line=3)
+
+```
+
+Comparison of log(Counts) for binding sites was under taken to confirm reproducibility.
+The data with the lowest correlation is shown and was seen between Replicate 1 and Replicate 3 in the control condition.
+
+## Comparision of ER binding from public datasets
+
+Comparison ER binding from public datasets.Common peaks detected for ER Ross-Innes CS, et al.
+2010; Welboren WJ, et al., 2009; Ceschin DG, et al.
+2011 and our data (FDR = 0.01).
+Venn diagram was generated with ChIPSeqAnno.
+```{r ERVennOverlap, message=FALSE, warning=FALSE}
+library(ChIPpeakAnno)
+library(GenomicRanges)
+
+m1 = read.table("bed/ERoverlap/Carrol-GenesDev.bed", sep="\t")
+m2 = read.table("bed/ERoverlap/ER_45minsGains_001_hg18.bed", sep="\t")
+m3 = read.table("bed/ERoverlap/LinER.txt", sep="\t")
+m4 = read.table("bed/ERoverlap/Welboren.bed", sep="\t")
+m5 = read.table("bed/ERoverlap/NA_peaks.narrowPeak", sep="\t") #Called with macs2 against EtOH control
+m5<-m5[c(1:3)]
+
+m1.r = BED2RangedData(m1)
+m2.r = BED2RangedData(m2)
+m3.r = BED2RangedData(m3)
+m4.r = BED2RangedData(m4)
+m5.r = BED2RangedData(m5)
+#mkd = makeVennDiagram(list(m1.r,m2.r,m3.r,m4.r,m5.r),
+#                      NameOfPeaks=c("Ross-Innes 2010", "Guertin 2017", "Lin 2007", "Welboren 2009","Ceschin 2011"), totalTest=200000)
+
+mkd = makeVennDiagram(list(m1.r,m2.r,m4.r,m5.r),
+                      NameOfPeaks=c("Ross-Innes 2010", "Guertin 2017", "Welboren 2009","Ceschin 2011"))
+
+```
+
+## Changes in H4K12ac following E2 treatment
+
+```{r H4K12acNormalised,results='hide', message=FALSE, warning=FALSE}
+filename<-"Rdata/066_SLX-15090_dba_human_ER_CTCF_merged.rda"
+load(filename)
+
+#install_github("andrewholding/Brundle")
+library(Brundle)
+
+jg.controlSampleSheet     <- "samplesheet/samplesheet_SLX15090_CTCF.csv"
+jg.experimentSampleSheet  <- "samplesheet/samplesheet_SLX15090_H4K12ac.csv"
+jg.treatedCondition       =  "Estrogen"
+jg.untreatedCondition     =  "none"
+
+jg.experimentPeaksetNormalised<-Brundle(
+            dbaExperiment,
+            dbaControl,
+            jg.treatedCondition,
+            jg.untreatedCondition,
+            jg.experimentSampleSheet,
+            jg.controlSampleSheet,
+            jg.noBAMs=TRUE
+            ) #BAMS we used in the manuscript data, but removed to keep file size down.
+
+jg.dba <- DiffBind:::pv.resetCounts(dbaExperiment, jg.experimentPeaksetNormalised)
+
+jg.dba_analysis<-dba.analyze(jg.dba)
+dba.plotMA(jg.dba_analysis,bFlip=TRUE,th=0.05)
+
 write.csv(dba.report(jg.dba_analysis,th=1),file="csv/H4_normalised_results.csv")
@@
 
-```{r,results='hide', message=FALSE, warning=FALSE}
+```{r H4K12acReproducibility,results='hide', message=FALSE, warning=FALSE}
 
@@
 
-```{r,results='hide', message=FALSE, warning=FALSE}
-
-profile<-read.table("SLX-15090/Homer_Profile.txt",sep="\t")
+```{r H4K12acProfile,results='hide', message=FALSE, warning=FALSE}
+
+profile<-read.table("SLX-15090/Homer_ER_Profile.txt",sep="\t")
 profile_ER<-as.matrix(profile[,c(1,2)])
-plot(profile_ER,t="l",col="red", xlab="Distance from TSS",ylab="Read Depth",main="H4K12ac Profile")
+plot(profile_ER,t="l",col="red", xlab="Distance from ER Binding site",ylab="Read Depth",main="H4K12ac Profile")
 profile_Ctrl<-as.matrix(profile[,c(1,5)])
@@
 H4k12ac occupancy profile before and after treatment with E2 shows a general increase around transcription start sites (TSS).
-
-## Activation of ER in MCF7
-
-```{r,results='hide', message=FALSE, warning=FALSE}
-
-jg.controlSampleSheet     <- "samplesheet/samplesheet_SLX15091_CTCF.csv"
-jg.experimentSampleSheet  <- "samplesheet/samplesheet_SLX15091_ER.csv"
-jg.treatedCondition       =  "Estrogen"
-jg.untreatedCondition     =  "none"
-
-#####
-#
-# Main Code
-#
-#####
-
-filename<-"Rdata/065_SLX-15091_dba_human_ER_CTCF.rda"
-load(filename)
-
-jg.dba<-Brundle(
-            dbaExperiment,
-            dbaControl,
-            jg.treatedCondition,
-            jg.untreatedCondition,
-            jg.experimentSampleSheet,
-            jg.controlSampleSheet )
-
-jg.dba_analysis<-dba.analyze(jg.dba)
-dba.plotMA(jg.dba_analysis,bFlip=TRUE,th=0.05)
-
-write.csv(dba.report(jg.dba_analysis,th=1),file="csv/E2_normalised_results.csv")
-
-```
-
-Comparison of fold-change of ER binding before and after treatment with estradiol.
-MA plot of ER binding after normalisation to CTCF binding displays a significant increase in ER binding at 45 minutes after treatment with estradiol.
-
-```{r,results='hide', message=FALSE, warning=FALSE}
-
-controlNames<-c("1a","3b")
-
-jg.experimentPeakset <- jg.dbaGetPeakset(dbaExperiment)
-relativeLibrarySize<-jg.getLibrarySize(dbaExperiment,
-                                       jg.experimentPeakset)
-peaksLibrarySize<-relativeLibrarySize[controlNames]
-peaks<-jg.experimentPeakset[controlNames]
-peaks[1]<-round(peaks[1]/peaksLibrarySize[1])
-peaks[2]<-round(peaks[2]/peaksLibrarySize[2])
-x<-c("Rep1","Rep3")
-names(x)<-controlNames
-x<-as.matrix(x)
-jg.plotScatter(peaks,controlNames, T,t(x),FALSE)
-Axis(side=2, labels=TRUE,ylab="Log(Counts)")
-mtext("log(Counts)", side=2, line=3)
-
-```
-
-Comparison of log(Counts) for binding sites was under taken to confirm reproducibility.
-The data with the lowest correlation is shown and was seen between Replicate 1 and Replicate 3 in the control condition.

### 2018-01-13 073_SLX-14229_SLX-8047_ER_normalisation_highlightCTCF.R

@@ -189,11 +189,11 @@ Mhs<-log2fc
 
 Ahs<-log10(rowSums(jg.EROnlyExperimentPeakset[4:11]))
 
-pdf("plots/pdf/074_ER-ER_normalised.pdf", points=15)
+#pdf("plots/pdf/074_ER-ER_normalised.pdf", points=15)
 par(mar=c(5.1,5.1,4.1,2.1))
 plot(Ahs,Mhs,cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5,pch=20,ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"),main="Counts normalized",ylim=c(-6.25,2))
 abline(h=0)
-dev.off()
+#dev.off()
 
 
 #NEW CODE
@@ -223,7 +223,7 @@ pfChIPuntreated<-rowMeans(as.matrix(elementMetadata(pfChIPerOverlap)[,jg.untreat
 pfChIPtreated<-rowMeans(as.matrix(elementMetadata(pfChIPerOverlap)[,jg.treatedNamesPF]))
 pfChIPtreated/pfChIPuntreated
 
-png("plots/073_CTCFproximal.png",pointsize=15)
+#png("plots/073_CTCFproximal.png",pointsize=15)
 boxplot(log(erChIPtreated/erChIPuntreated),
         log(pfChIPtreated/pfChIPuntreated),
         xlab="Experiment",
@@ -232,4 +232,4 @@ boxplot(log(erChIPtreated/erChIPuntreated),
         names=c(" Cross-Normalised"," Parrallel Factor"),
         pch=20)
 
-dev.off()
+#dev.off()

### 2018-02-20 Manuscript_Figures.Rmd
Added PDX Analysis

--- 2017-12-05 Manuscript_Figures.Rmd
@@
 
-##Comparison of mormalisation DiffBind plots
+##Comparison of normalisation DiffBind plots
 
@@
 
-jg.dba<-Brundle(
+jg.experimentPeaksetNormalised<-Brundle(
             dbaExperiment,
@@
             jg.controlSampleSheet )
+
+jg.dba <- DiffBind:::pv.resetCounts(dbaExperiment, jg.experimentPeaksetNormalised)
 
@@
 The data with the lowest correlation is shown and was seen between Replicate 1 and Replicate 3 in the control condition.
+
+#PDX Analysis
+
+```{r, results='hide', message=FALSE, warning=FALSE}
+
+pdx<-readRDS(file="Rdata/pdx.rds")
+ctcf<-readRDS(file="Rdata/ctcf.rds")
+
+pdx<- dba.count(pdx,peaks = NULL, score = DBA_SCORE_READS)
+ctcf<- dba.count(ctcf,peaks = NULL, score = DBA_SCORE_READS)
+
+peakset.ctcf <- dba.peakset(ctcf, bRetrieve = TRUE, DataType = DBA_DATA_FRAME)
+peakset.er <- dba.peakset(pdx, bRetrieve = TRUE, DataType = DBA_DATA_FRAME)
+
+normalisation.factor<-colSums(peakset.ctcf[4:8])/median(colSums(peakset.ctcf[4:8]))
+
+normalised.er<-cbind(peakset.er[1:3],peakset.er[4]/normalisation.factor[1],
+                                     peakset.er[5]/normalisation.factor[2],
+                                     peakset.er[6]/normalisation.factor[3],
+                                     peakset.er[7]/normalisation.factor[4],
+                                     peakset.er[8]/normalisation.factor[5])
+
+colSums(normalised.er[4:8])
+colSums(peakset.er[4:8])
+
+normalised.ctcf<-cbind(peakset.ctcf[1:3],
+                       peakset.ctcf[4]/normalisation.factor[1],
+                       peakset.ctcf[5]/normalisation.factor[2],
+                       peakset.ctcf[6]/normalisation.factor[3],
+                       peakset.ctcf[7]/normalisation.factor[4],
+                       peakset.ctcf[8]/normalisation.factor[5])
+
+colSums(normalised.ctcf[4:8])
+
+rpm<-as.numeric(pdx$class["Reads",])/10^6
+
+addToDf<-function(data,factor,method,locus) {
+        data.frame(counts=as.numeric(t(data)),factor=rep(factor,4),method=rep(method,4),rep=c("PDX01","PDX03","PDX04","PDX05"),locus=rep(locus,4))
+}
+
+library(ggpubr)
+
+relativeRPM<-(rpm[c(1,3:5)]/median(rpm[c(1,3:5)]))
+
+df<-addToDf(peakset.ctcf[13416,c(4,6:8)],"CTCF","Raw","RARa")
+df<-rbind(df,addToDf(peakset.er[639,c(4,6:8)],"ER","Raw","RARa"))
+df<-rbind(df,addToDf(peakset.ctcf[13416,c(4,6:8)]/relativeRPM,"CTCF","RPM","RARa"))
+df<-rbind(df,addToDf(peakset.er[639,c(4,6:8)]/relativeRPM,"ER","RPM","RARa"))
+df<-rbind(df,addToDf(normalised.ctcf[13416,c(4,6:8)],"CTCF","pfChIP","RARa"))
+df<-rbind(df,addToDf(normalised.er[639,c(4,6:8)],"ER","pfChIP","RARa"))
+
+ctcfSite<-"21382"
+erSite<-"891"
+name<-"GREB1"
+df<-rbind(df,addToDf(peakset.ctcf[ctcfSite,c(4,6:8)],"CTCF","Raw",name))
+df<-rbind(df,addToDf(peakset.er[erSite,c(4,6:8)],"ER","Raw",name))
+df<-rbind(df,addToDf(normalised.ctcf[ctcfSite,c(4,6:8)],"CTCF","pfChIP",name))
+df<-rbind(df,addToDf(normalised.er[erSite,c(4,6:8)],"ER","pfChIP",name))
+df<-rbind(df,addToDf(peakset.ctcf[ctcfSite,c(4,6:8)]/relativeRPM,"CTCF","RPM",name))
+df<-rbind(df,addToDf(peakset.er[erSite,c(4,6:8)]/relativeRPM,"ER","RPM",name))
+
+ctcfSite<-"20453"
+erSite<-"1064"
+name<-"CLIC6"
+df<-rbind(df,addToDf(peakset.ctcf[ctcfSite,c(4,6:8)],"CTCF","Raw",name))
+df<-rbind(df,addToDf(peakset.er[erSite,c(4,6:8)],"ER","Raw",name))
+df<-rbind(df,addToDf(normalised.ctcf[ctcfSite,c(4,6:8)],"CTCF","pfChIP",name))
+df<-rbind(df,addToDf(normalised.er[erSite,c(4,6:8)],"ER","pfChIP",name))
+df<-rbind(df,addToDf(peakset.ctcf[ctcfSite,c(4,6:8)]/relativeRPM,"CTCF","RPM",name))
+df<-rbind(df,addToDf(peakset.er[erSite,c(4,6:8)]/relativeRPM,"ER","RPM",name))
+
+ggbarplot(df, x="rep", y = "counts",
+          fill = "method",    
+          color = "white",            # Set bar border colors to white
+          palette =c("#FC4E07", "#E7B800", "#00AFBB"),
+          sort.by.groups = FALSE,     # Don't sort inside each group
+          x.text.angle = 90,          # Rotate vertically x axis texts
+          ylab = "Corrected Counts",
+          xlab = "PDX",
+          legend.title = "Comparison",
+          facet.by=c("locus","factor"),
+          position=position_dodge(width=0.7)
+)
+
+ggboxplot(df, x="method", y = "counts",
+          palette =c("#FC4E07", "#E7B800", "#00AFBB"),
+          ylab = "Corrected Counts",
+          xlab = "PDX",
+          color="method",
+          facet.by=c("locus","factor"),
+          legend.title = "Comparison"
+)
+
+```
+
+```{r, results='hide', message=FALSE, warning=FALSE}
+#Homer profile
+
+ctcf<-read.table("SLX-15439/blacklist_filtered/Homer_CTCF_Profile.txt", sep="\t", header=TRUE)
+er<-read.table("SLX-15439/blacklist_filtered/Homer_ER_Profile.txt", sep="\t", header=TRUE)
+
+par(mfrow=c(2,2))
+profile_CTCF<-as.matrix(ctcf[,c(1,2,5,8,11)])
+plot(profile_CTCF,t="l",lwd=3,col="#00AFBB", xlab="Distance from Summit",ylab="Read Depth",main="CTCF Profile")
+lines(profile_CTCF[,c(1,3)],lwd=3,col="#E7B800")
+lines(profile_CTCF[,c(1,4)],lwd=3,col="#FC4E07")
+lines(profile_CTCF[,c(1,5)],lwd=3,col="#FC00FE")
+legend("topright",legend = c("PDX01", "PDX03", "PDX04", "PDX05"), col=c("#00AFBB","#E7B800", "#FC4E07","#FC00FE"),pch = c(19,19))
+
+maxCTCF<-colMaxs(profile_CTCF)
+minCTCF<-colMins(profile_CTCF)
+
+profile_CTCF[,2:5]<-profile_CTCF[,2:5]-minCTCF[2:5]
+
+colSums(profile_CTCF[,2:5]/maxCTCF[2:5])
+normalised_CTCF<-t((t(profile_CTCF[,2:5])/maxCTCF[2:5]))
+normalised_CTCF<-cbind(profile_CTCF[,1],normalised_CTCF)
+
+plot(normalised_CTCF,lwd=3,t="l",col="#00AFBB", xlab="Distance from Summit",ylab="Read Depth",main="Normalised CTCF Profile")
+lines(normalised_CTCF[,c(1,3)],lwd=3,col="#E7B800")
+lines(normalised_CTCF[,c(1,4)],lwd=3,col="#FC4E07")
+lines(normalised_CTCF[,c(1,5)],lwd=3,col="#FC00FE")
+legend("topright",legend = c("PDX01", "PDX03", "PDX04", "PDX05"), col=c("#00AFBB","#E7B800", "#FC4E07","#FC00FE"),pch = c(19,19))
+
+profile_ER<-as.matrix(er[,c(1,2,5,8,11)])
+plot(profile_ER,t="l",lwd=3,col="#00AFBB", xlab="Distance from Summit",ylab="Read Depth",main="ER Profile")
+lines(profile_ER[,c(1,3)],lwd=3,col="#E7B800")
+lines(profile_ER[,c(1,4)],lwd=3,col="#FC4E07")
+lines(profile_ER[,c(1,5)],lwd=3,col="#FC00FE")
+legend("topright",legend = c("PDX01", "PDX03", "PDX04", "PDX05"), col=c("#00AFBB","#E7B800", "#FC4E07","#FC00FE"),pch = c(19,19))
+
+normalised_ER<-t((t(profile_ER[,2:5])/maxCTCF[2:5]))
+normalised_ER<-cbind(profile_ER[,1],normalised_ER)
+
+plot(normalised_ER,t="l",lwd=3,col="#00AFBB", xlab="Distance from Summit",ylab="Read Depth",main="Normalised ER Profile")
+lines(normalised_ER[,c(1,3)],lwd=3,col="#E7B800")
+lines(normalised_ER[,c(1,4)],lwd=3,col="#FC4E07")
+lines(normalised_ER[,c(1,5)],lwd=3,col="#FC00FE")
+legend("topright",legend = c("PDX01", "PDX03", "PDX04", "PDX05"), col=c("#00AFBB","#E7B800", "#FC4E07","#FC00FE"),pch = c(19,19))
+
+```

### 2018-02-20 074_SLX-15439_Download_reads.sh

@@ -1,0 +1,3 @@
+#!/bin/sh
+java -jar ../java/clarity-tools.jar -l   SLX-15439
+

### 2018-02-20 075_SLX-15439_blacklist.sh

@@ -1,0 +1,24 @@
+
+bl=../blacklists/hg19-blacklist.bed
+
+
+cd SLX-15439
+mkdir ./blacklist_filtered
+for f in *.bam
+do
+	echo $f
+	bedtools intersect -v -abam $f -b $bl  > ./blacklist_filtered/$f
+done
+
+### Re-index
+
+cd ./blacklist_filtered
+for f in *.bam
+do
+samtools index $f
+done
+
+cd ..
+
+
+

### 2018-02-20 076_SLX-15439_peakCalling.sh

@@ -1,0 +1,36 @@
+### MACS peak caller
+
+### Run macs on the blacklisted data
+cd ./SLX-15439
+mkdir ./peaks
+cd peaks
+
+
+control=D702_D504
+bam=D701_D501
+macs2 callpeak -t ../blacklist_filtered/SLX-15439.$bam.HNKNYBBXX.s_6.bwa.homo_sapiens.bam -c ../blacklist_filtered/SLX-15439.$control.HNKNYBBXX.s_6.bwa.homo_sapiens.bam -f BAM -n $bam.bam -g hs &
+
+
+control=D701_D504
+bam=D703_D502
+macs2 callpeak -t ../blacklist_filtered/SLX-15439.$bam.HNKNYBBXX.s_6.bwa.homo_sapiens.bam -c ../blacklist_filtered/SLX-15439.$control.HNKNYBBXX.s_6.bwa.homo_sapiens.bam -f BAM -n $bam.bam -g hs &
+
+
+control=D702_D502
+bam=D701_D503
+macs2 callpeak -t ../blacklist_filtered/SLX-15439.$bam.HNKNYBBXX.s_6.bwa.homo_sapiens.bam -c ../blacklist_filtered/SLX-15439.$control.HNKNYBBXX.s_6.bwa.homo_sapiens.bam -f BAM -n $bam.bam -g hs &
+
+control=D701_D502
+bam=D702_D501
+macs2 callpeak -t ../blacklist_filtered/SLX-15439.$bam.HNKNYBBXX.s_6.bwa.homo_sapiens.bam -c ../blacklist_filtered/SLX-15439.$control.HNKNYBBXX.s_6.bwa.homo_sapiens.bam -f BAM -n $bam.bam -g hs &
+
+control=D703_D501
+bam=D702_D503
+macs2 callpeak -t ../blacklist_filtered/SLX-15439.$bam.HNKNYBBXX.s_6.bwa.homo_sapiens.bam -c ../blacklist_filtered/SLX-15439.$control.HNKNYBBXX.s_6.bwa.homo_sapiens.bam -f BAM -n $bam.bam -g hs &
+
+
+
+
+
+
+

### 2018-02-20 077_SLX-15439_splitCTCF.sh

@@ -1,0 +1,18 @@
+cd SLX-15439/peaks/
+mkdir CTCF
+
+#BEDs for CTCF from SLX-15091
+
+for f in *.narrowPeak
+do
+	echo $f
+	bedtools intersect -a $f -b ../../txt/CTCF_ERConsensusremoved.bed > CTCF/$f 
+done
+
+mkdir ER
+for f in *.narrowPeak
+do
+	echo $f
+	bedtools subtract -b ../../txt/CTCF_union.bed -a $f -A > ER/$f 
+done
+

### 2018-02-20 078_SLX-15439_LoadCounts.R

@@ -1,0 +1,32 @@
+library(Brundle)
+
+#"SLX-15439","D701-D501","ATTACTCG-TATAGCCT","PDX01-AB555B-x2" ER 8, PR 2, FOXA1 8
+#"SLX-15439","D703-D502","CGCTCATT-ATAGAGGC","PDX02-VHI0244o2-x4" ER 8, PR 7, FOXA1 8
+#"SLX-15439","D701-D503","ATTACTCG-CCTATCCT","PDX03-AB580-x1"  - ER 8, PR/FOXA1 unknown
+#"SLX-15439","D702-D501","TCCGGAGA-TATAGCCT","PDX04-STG195-x4" - ER/PR/FOXA1/8
+#"SLX-15439","D702-D503","TCCGGAGA-CCTATCCT","PDX05-V0980U-x2"  ER 5, PR 0, FOXA
+
+
+pdx<-dba(sampleSheet="samplesheet/SLX-15439.csv")
+pdx<-dba.count(pdx, bParallel=FALSE,bRemoveDuplicates=TRUE)
+dba.plotPCA(pdx)
+dba.plotHeatmap(pdx)
+saveRDS(file="Rdata/pdx.rds",pdx)
+
+
+ctcf<-dba(sampleSheet="samplesheet/SLX-15439_CTCF.csv")
+ctcf<-dba.count(ctcf, bParallel=FALSE,minOverlap=3,bRemoveDuplicates=TRUE)
+dba.plotPCA(ctcf)
+dba.plotHeatmap(ctcf)
+saveRDS(file="Rdata/ctcf.rds",ctcf)
+
+pdx<-readRDS(file="Rdata/pdx.rds")
+ctcf<-readRDS(file="Rdata/ctcf.rds")
+
+library(ChIPQC)
+samplesheet<-read.csv("samplesheet/SLX-15439_CTCF.csv")
+exp = ChIPQC(ctcf)
+ChIPQCreport(exp,reportFolder="ChIPQCreport/SLX-15439.csv")
+
+write.table(file="txt/PDXconsensus.txt",as.data.frame(dba.peakset(pdx,bRetrieve=TRUE))[,1:3],quote=FALSE, row.names = FALSE, col.names=FALSE, sep="\t")
+

### 2018-02-20 079_SLX-15439_supp_fig.R

@@ -1,0 +1,96 @@
+library(Brundle)
+
+
+pdx<-readRDS(file="Rdata/pdx.rds")
+ctcf<-readRDS(file="Rdata/ctcf.rds")
+
+pdx<- dba.count(pdx,peaks = NULL, score = DBA_SCORE_READS)
+ctcf<- dba.count(ctcf,peaks = NULL, score = DBA_SCORE_READS)
+
+peakset.ctcf <- dba.peakset(ctcf, bRetrieve = TRUE, DataType = DBA_DATA_FRAME)
+peakset.er <- dba.peakset(pdx, bRetrieve = TRUE, DataType = DBA_DATA_FRAME)
+
+normalisation.factor<-colSums(peakset.ctcf[4:8])/median(colSums(peakset.ctcf[4:8]))
+
+
+
+normalised.er<-cbind(peakset.er[1:3],peakset.er[4]/normalisation.factor[1],
+                                     peakset.er[5]/normalisation.factor[2],
+                                     peakset.er[6]/normalisation.factor[3],
+                                     peakset.er[7]/normalisation.factor[4],
+                                     peakset.er[8]/normalisation.factor[5])
+
+colSums(normalised.er[4:8])
+colSums(peakset.er[4:8])
+
+normalised.ctcf<-cbind(peakset.ctcf[1:3],
+                       peakset.ctcf[4]/normalisation.factor[1],
+                       peakset.ctcf[5]/normalisation.factor[2],
+                       peakset.ctcf[6]/normalisation.factor[3],
+                       peakset.ctcf[7]/normalisation.factor[4],
+                       peakset.ctcf[8]/normalisation.factor[5])
+
+colSums(normalised.ctcf[4:8])
+
+rpm<-as.numeric(pdx$class["Reads",])/10^6
+
+
+addToDf<-function(data,factor,method,locus) {
+        data.frame(counts=as.numeric(t(data)),factor=rep(factor,4),method=rep(method,4),rep=c("PDX01","PDX03","PDX04","PDX05"),locus=rep(locus,4))
+}
+
+library(ggpubr)
+
+relativeRPM<-(rpm[c(1,3:5)]/median(rpm[c(1,3:5)]))
+
[…]

### 2018-02-20 080_SLX-15439_peak_profile.sh

@@ -1,0 +1,8 @@
+cd SLX-15439/blacklist_filtered/
+makeTagDirectory PDX01/ SLX-15439.D701_D501.HNKNYBBXX.s_6.bwa.homo_sapiens.bam
+makeTagDirectory PDX03/ SLX-15439.D701_D503.HNKNYBBXX.s_6.bwa.homo_sapiens.bam
+makeTagDirectory PDX04/ SLX-15439.D702_D501.HNKNYBBXX.s_6.bwa.homo_sapiens.bam
+makeTagDirectory PDX05/ SLX-15439.D702_D503.HNKNYBBXX.s_6.bwa.homo_sapiens.bam
+
+annotatePeaks.pl ../../txt//CTCF_ERConsensusremoved.bed hg19 -size 6000 -hist 25 -d PDX01 PDX03 PDX04 PDX05 > Homer_CTCF_Profile.txt
+annotatePeaks.pl ../../txt/PDXconsensus.txt hg19 -size 6000 -hist 25 -d PDX01 PDX03 PDX04 PDX05 > Homer_ER_Profile.txt

### 2018-02-20 081_SLX-15439_homer.R

@@ -1,0 +1,53 @@
+#Homer profile
+
+ctcf<-read.table("SLX-15439/blacklist_filtered/Homer_CTCF_Profile.txt", sep="\t", header=TRUE)
+er<-read.table("SLX-15439/blacklist_filtered/Homer_ER_Profile.txt", sep="\t", header=TRUE)
+
+
+par(mfrow=c(2,2))
+profile_CTCF<-as.matrix(ctcf[,c(1,2,5,8,11)])
+plot(profile_CTCF,t="l",lwd=3,col="#00AFBB", xlab="Distance from Summit",ylab="Read Depth",main="CTCF Profile")
+lines(profile_CTCF[,c(1,3)],lwd=3,col="#E7B800")
+lines(profile_CTCF[,c(1,4)],lwd=3,col="#FC4E07")
+lines(profile_CTCF[,c(1,5)],lwd=3,col="#FC00FE")
+legend("topright",legend = c("PDX01", "PDX03", "PDX04", "PDX05"), col=c("#00AFBB","#E7B800", "#FC4E07","#FC00FE"),pch = c(19,19))
+
+maxCTCF<-colMaxs(profile_CTCF)
+minCTCF<-colMins(profile_CTCF)
+
+profile_CTCF[,2:5]<-profile_CTCF[,2:5]-minCTCF[2:5]
+
+colSums(profile_CTCF[,2:5]/maxCTCF[2:5])
+normalised_CTCF<-t((t(profile_CTCF[,2:5])/maxCTCF[2:5]))
+normalised_CTCF<-cbind(profile_CTCF[,1],normalised_CTCF)
+
+
+
+plot(normalised_CTCF,lwd=3,t="l",col="#00AFBB", xlab="Distance from Summit",ylab="Read Depth",main="Normalised CTCF Profile")
+lines(normalised_CTCF[,c(1,3)],lwd=3,col="#E7B800",)
+lines(normalised_CTCF[,c(1,4)],lwd=3,col="#FC4E07")
+lines(normalised_CTCF[,c(1,5)],lwd=3,col="#FC00FE")
+legend("topright",legend = c("PDX01", "PDX03", "PDX04", "PDX05"), col=c("#00AFBB","#E7B800", "#FC4E07","#FC00FE"),pch = c(19,19))
+
+
+
+
+profile_ER<-as.matrix(er[,c(1,2,5,8,11)])
[…]

### 2018-02-20 SLX-15439/peaks/D701_D501.bam_model.r

@@ -1,0 +1,17 @@
+# R script for Peak Model
+#  -- generated by MACS
+p <- c(0.0169790987323,0.0198171286684,0.0226522714936,0.0255538178677,0.0285823971179,0.02866034911,0.0288018175402,0.0288480113133,0.0291598192819,0.029462965918,0.0296621765646,0.029751677,0.0299451134249,0.0300981302984,0.0300490494144,0.0299826458656,0.0299451134249,0.0301443240715,0.030216501842,0.0301905178446,0.0301587596256,0.030184743623,0.0301991791771,0.0303204378315,0.0304041640453,0.0304878902591,0.0305023258132,0.0303666316046,0.0303060022774,0.0303031151666,0.0304907773699,0.0307419560112,0.0307708271194,0.0307592786761,0.0309122955496,0.031082635088,0.031290507067,0.0314435239405,0.0313511363942,0.0315070403785,0.0314608466054,0.0314550723838,0.0315907665923,0.0315994279248,0.0315561212625,0.0313886688349,0.0313540235051,0.0313626848375,0.0315907665923,0.0314521852729,0.0315647825949,0.0315099274893,0.0313828946133,0.0313424750618,0.0313511363942,0.0314550723838,0.0313049426211,0.0312731844021,0.0311663613018,0.0313944430565,0.0314204270539,0.0316485088087,0.0316196377005,0.0316196377005,0.0317091381359,0.0317120252467,0.0319401070015,0.0320411558802,0.0321710758671,0.0320729140992,0.032061365656,0.0318275096795,0.0319949621071,0.0321017852074,0.0320758012101,0.0321970598645,0.0320382687694,0.0319545425556,0.0318621550094,0.0319776394422,0.0320758012101,0.0320065105504,0.0320209461045,0.0320267203261,0.0317524447982,0.0316716056953,0.0317062510251,0.0316658314736,0.0317784287956,0.0316658314736,0.0315503470408,0.0316167505897,0.0315705568166,0.0316485088087,0.032012284772,0.0320382687694,0.0321537532022,0.032188398532,0.032205721197,0.032269237635,0.0323876091786,0.0325030936114,0.0326359007091,0.0325406260521,0.0323443025163,0.0324309158409,0.0323702865137,0.0324482385058,0.0323933834003,0.0325694971603,0.0326561104849,0.0326532233741,0.032526190498,0.0325204162763,0.0327196269229,0.0327196269229,0.0326474491524,0.0327860304718,0.0327398366987,0.03269941714
[…]

### 2018-02-20 SLX-15439/peaks/D701_D503.bam_model.r

@@ -1,0 +1,17 @@
+# R script for Peak Model
+#  -- generated by MACS
+p <- c(0.0145708579908,0.0168068890283,0.0191712169287,0.0215355448291,0.0239640211609,0.0240556617772,0.0241931227017,0.0240281695923,0.0238815446063,0.0237990680516,0.0235608024492,0.0239090367911,0.0239640211609,0.0238907086679,0.0242572711331,0.0240006774074,0.0243030914412,0.0245413570436,0.0248987554471,0.0247796226459,0.0250637085564,0.0249537398169,0.024852935139,0.0250820366797,0.0247154742145,0.025100364803,0.0249629038785,0.0249170835704,0.0249812320018,0.0254027788367,0.0252836460356,0.0253386304053,0.0256502085007,0.0255494038228,0.0256593725624,0.0254852553914,0.0256318803775,0.0256043881926,0.0254577632065,0.0253844507135,0.0250178882483,0.0250545444948,0.0249629038785,0.024752130461,0.024926247632,0.0249720679402,0.024926247632,0.024852935139,0.0250820366797,0.0248987554471,0.0250453804332,0.0253294663437,0.0254669272681,0.0256960288089,0.0256318803775,0.0256502085007,0.0255402397612,0.0256502085007,0.0253661225902,0.025347794467,0.0255585678844,0.0250912007413,0.0246238335982,0.0246879820296,0.0244955367354,0.0245596851668,0.0245138648587,0.0244863726738,0.0247429663994,0.0250545444948,0.025100364803,0.0251553491727,0.0253844507135,0.0252561538507,0.0253661225902,0.0251553491727,0.0253752866518,0.0253569585286,0.0254302710216,0.0252561538507,0.0250178882483,0.0250637085564,0.025173677296,0.0250912007413,0.0252744819739,0.0253569585286,0.0249445757553,0.0255035835147,0.0253386304053,0.0249720679402,0.0251645132344,0.024752130461,0.0246238335982,0.0248162788925,0.0247063101529,0.0245780132901,0.0250270523099,0.024358075811,0.0241564664551,0.0243672398726,0.0242755992563,0.0247154742145,0.0246238335982,0.0245780132901,0.0244680445505,0.0244038961191,0.0241014820854,0.0243855679959,0.0246329976599,0.0246513257831,0.0248071148308,0.0250637085564,0.0252286616658,0.0253752866518,0.0254577632065,0.0255035835147,0.0255310756995,0.0255035835147,0.0251370210495,
[…]

### 2018-02-20 SLX-15439/peaks/D702_D501.bam_model.r

@@ -1,0 +1,17 @@
+# R script for Peak Model
+#  -- generated by MACS
+p <- c(0.0147293179978,0.0170416423293,0.0193727660456,0.0216756906847,0.0238470196301,0.0239034177846,0.0242324070187,0.0241102110175,0.0243922017896,0.0243076045579,0.024448599944,0.0245989950225,0.024495598406,0.0245331971757,0.0245895953301,0.0242700057883,0.0238282202453,0.0238376199377,0.0237060242441,0.0237906214757,0.0234804316264,0.0232736383935,0.0234240334719,0.023518030396,0.0237248236289,0.023959815939,0.0242606060959,0.0242230073263,0.0242418067111,0.0246553931769,0.0246365937921,0.0246929919465,0.024796388563,0.0250125814883,0.0246741925617,0.0245519965605,0.0241760088643,0.0243640027124,0.0245895953301,0.0241102110175,0.0241572094795,0.02435460302,0.0242136076339,0.0239034177846,0.0243264039428,0.0241008113251,0.0242606060959,0.0239410165542,0.0234992310112,0.0232830380859,0.0232642387011,0.0231514423922,0.0231984408542,0.0231890411618,0.0231044439302,0.0234146337795,0.0237060242441,0.0238752187073,0.0240444131706,0.0242230073263,0.0241854085567,0.0241384100947,0.0241854085567,0.0243264039428,0.023912817477,0.0237718220909,0.023659025782,0.0235744285504,0.0233676353175,0.0233582356251,0.0231420426998,0.0230104470061,0.0228882510049,0.0226156599251,0.0230668451606,0.0231326430074,0.0230856445454,0.0230856445454,0.0230574454682,0.0230386460833,0.02361202732,0.0239034177846,0.0238846183997,0.0239504162466,0.0238094208605,0.02361202732,0.0234240334719,0.0236402263972,0.0237718220909,0.0237906214757,0.0233958343947,0.0231984408542,0.0232454393162,0.0233582356251,0.0232266399314,0.0234240334719,0.023471031934,0.0232266399314,0.0233112371631,0.0233018374707,0.0234334331643,0.023471031934,0.0235744285504,0.0236872248593,0.0235838282428,0.0233864347023,0.0236402263972,0.0237624223985,0.0236778251668,0.0235368297808,0.0234898313188,0.0236966245517,0.0237342233213,0.0234616322415,0.0239222171694,0.0239974147086,0.0236026276276,0.0235462294732,0.0237342233213,0.02
[…]

### 2018-02-20 SLX-15439/peaks/D702_D503.bam_model.r

@@ -1,0 +1,17 @@
+# R script for Peak Model
+#  -- generated by MACS
+p <- c(0.01431100023,0.0167211306208,0.01930146796,0.0218341473537,0.0242102363548,0.0241217287416,0.0239310969593,0.0242510860225,0.0247208572003,0.0247344737562,0.0247685151459,0.024870639315,0.0246323495871,0.0244757591945,0.0246731992547,0.0245846916415,0.0247004323665,0.0243940598592,0.0243804433033,0.0242715108563,0.024421292971,0.0246663909768,0.0247821317018,0.0249727634841,0.0247208572003,0.0251702035444,0.0255378505531,0.0254016849943,0.0251838201002,0.0254221098281,0.0252042449341,0.0247344737562,0.0247276654783,0.0247344737562,0.0249182972606,0.0246595826989,0.0244553343607,0.0247753234238,0.0247208572003,0.0246731992547,0.0249591469282,0.0251838201002,0.0252995608252,0.0253336022149,0.0254016849943,0.0252723277135,0.0253131773811,0.0253131773811,0.0256059333325,0.0257216740575,0.0260144300089,0.026007621731,0.0260825127883,0.026000813453,0.025891881006,0.0259667720633,0.0258169899487,0.0259735803413,0.0257284823354,0.0256263581663,0.0252314780458,0.025102120765,0.0249182972606,0.0246527744209,0.0245370336959,0.0243736350254,0.0243327853577,0.0239242886813,0.0241149204637,0.0237745065667,0.0235974913402,0.0236451492858,0.0236791906755,0.0237608900108,0.0238085479564,0.023849397624,0.023740465177,0.0238221645122,0.0235702582285,0.0237064237873,0.023624724452,0.0235498333946,0.0234340926696,0.0235838747843,0.0235021754491,0.0233796264461,0.0234409009476,0.0233251602226,0.0232094194976,0.0231889946638,0.023277502277,0.0231753781079,0.0232979271108,0.0229166635462,0.0228894304344,0.0229030469903,0.0228349642109,0.0229234718241,0.0229711297697,0.0228894304344,0.022943896658,0.0228826221565,0.022596674483,0.0229507049359,0.0226375241507,0.0227804979874,0.0226783738183,0.0228690056006,0.0229507049359,0.0228145393771,0.0229643214918,0.0231958029417,0.0233387767785,0.0232026112197,0.0234204761138,0.0233319685005,0.0234681340593,0.0231889946638,0.0230051711594,0.02
[…]

### 2018-02-20 SLX-15439/peaks/D703_D502.bam_model.r

@@ -1,0 +1,17 @@
+# R script for Peak Model
+#  -- generated by MACS
+p <- c(0.00234351198706,0.00234351198706,0.00312468264942,0.00390585331177,0.00468702397413,0.00312468264942,0.00312468264942,0.00390585331177,0.00468702397413,0.00468702397413,0.00703053596119,0.00937404794826,0.00937404794826,0.00937404794826,0.0085928772859,0.0085928772859,0.0085928772859,0.00781170662355,0.00703053596119,0.00703053596119,0.00546819463648,0.00312468264942,0.00234351198706,0.00156234132471,0.00234351198706,0.00312468264942,0.00390585331177,0.00468702397413,0.00624936529884,0.00624936529884,0.00624936529884,0.00624936529884,0.00624936529884,0.00624936529884,0.00624936529884,0.00624936529884,0.00546819463648,0.00468702397413,0.00390585331177,0.00312468264942,0.00234351198706,0.00390585331177,0.00390585331177,0.00390585331177,0.00390585331177,0.00312468264942,0.00312468264942,0.00312468264942,0.00234351198706,0.00234351198706,0.00234351198706,0.00156234132471,0.00234351198706,0.00234351198706,0.00234351198706,0.00234351198706,0.00234351198706,0.00234351198706,0.00312468264942,0.00312468264942,0.00390585331177,0.00468702397413,0.00390585331177,0.00390585331177,0.00312468264942,0.00312468264942,0.00312468264942,0.00312468264942,0.00234351198706,0.00234351198706,0.00156234132471,0.000781170662355,0.00234351198706,0.00312468264942,0.00312468264942,0.00312468264942,0.00312468264942,0.00312468264942,0.00312468264942,0.00390585331177,0.00390585331177,0.00390585331177,0.00234351198706,0.00156234132471,0.00234351198706,0.00234351198706,0.00234351198706,0.00312468264942,0.00312468264942,0.00390585331177,0.00390585331177,0.00390585331177,0.00703053596119,0.00703053596119,0.00624936529884,0.00624936529884,0.00624936529884,0.00546819463648,0.00546819463648,0.00468702397413,0.00546819463648,0.00468702397413,0.00156234132471,0.00156234132471,0.00156234132471,0.00234351198706,0.00312468264942,0.00312468264942,0.00312468264942,0.00234351198706,0.00312468264942,0.00312468264942,0.00312468264942,0.00312468264942,0.00390585331177,0.00468702397413,0.003905853
[…]

### 2018-02-20 Manuscript_Figures.Rmd
Corrected headings

--- 2018-01-13 Manuscript_Figures.Rmd
@@
 H4k12ac occupancy profile before and after treatment with E2 shows a general increase around transcription start sites (TSS).
+
+## Activation of ER in MCF7
+
+```{r,results='hide', message=FALSE, warning=FALSE}
+
+jg.controlSampleSheet     <- "samplesheet/samplesheet_SLX15091_CTCF.csv"
+jg.experimentSampleSheet  <- "samplesheet/samplesheet_SLX15091_ER.csv"
+jg.treatedCondition       =  "Estrogen"
+jg.untreatedCondition     =  "none"
+
+#####
+#
+# Main Code
+#
+#####
+
+filename<-"Rdata/065_SLX-15091_dba_human_ER_CTCF.rda"
+load(filename)
+
+jg.experimentPeaksetNormalised<-Brundle(
+            dbaExperiment,
+            dbaControl,
+            jg.treatedCondition,
+            jg.untreatedCondition,
+            jg.experimentSampleSheet,
+            jg.controlSampleSheet )
+
+jg.dba <- DiffBind:::pv.resetCounts(dbaExperiment, jg.experimentPeaksetNormalised)
+
+jg.dba_analysis<-dba.analyze(jg.dba)
+dba.plotMA(jg.dba_analysis,bFlip=TRUE,th=0.05)
+
+write.csv(dba.report(jg.dba_analysis,th=1),file="csv/E2_normalised_results.csv")
+
+```
+
+Comparison of fold-change of ER binding before and after treatment with estradiol.
+MA plot of ER binding after normalisation to CTCF binding displays a significant increase in ER binding at 45 minutes after treatment with estradiol.
+
+```{r,results='hide', message=FALSE, warning=FALSE}
+
+controlNames<-c("1a","3b")
+
+jg.experimentPeakset <- jg.dbaGetPeakset(dbaExperiment)
+relativeLibrarySize<-jg.getLibrarySize(dbaExperiment,
+                                       jg.experimentPeakset)
+peaksLibrarySize<-relativeLibrarySize[controlNames]
+peaks<-jg.experimentPeakset[controlNames]
+peaks[1]<-round(peaks[1]/peaksLibrarySize[1])
+peaks[2]<-round(peaks[2]/peaksLibrarySize[2])
+x<-c("Rep1","Rep3")
+names(x)<-controlNames
+x<-as.matrix(x)
+jg.plotScatter(peaks,controlNames, T,t(x),FALSE)
+Axis(side=2, labels=TRUE,ylab="Log(Counts)")
+mtext("log(Counts)", side=2, line=3)
+
+```
+
+Comparison of log(Counts) for binding sites was under taken to confirm reproducibility.
+The data with the lowest correlation is shown and was seen between Replicate 1 and Replicate 3 in the control condition.
+
+##PDX Analysis
+
+```{r, results='hide', message=FALSE, warning=FALSE}
+
+pdx<-readRDS(file="Rdata/pdx.rds")
+ctcf<-readRDS(file="Rdata/ctcf.rds")
+
+pdx<- dba.count(pdx,peaks = NULL, score = DBA_SCORE_READS)
+ctcf<- dba.count(ctcf,peaks = NULL, score = DBA_SCORE_READS)
+
+peakset.ctcf <- dba.peakset(ctcf, bRetrieve = TRUE, DataType = DBA_DATA_FRAME)
+peakset.er <- dba.peakset(pdx, bRetrieve = TRUE, DataType = DBA_DATA_FRAME)
+
+normalisation.factor<-colSums(peakset.ctcf[4:8])/median(colSums(peakset.ctcf[4:8]))
+
+normalised.er<-cbind(peakset.er[1:3],peakset.er[4]/normalisation.factor[1],
+                                     peakset.er[5]/normalisation.factor[2],
+                                     peakset.er[6]/normalisation.factor[3],
+                                     peakset.er[7]/normalisation.factor[4],
+                                     peakset.er[8]/normalisation.factor[5])
+
+colSums(normalised.er[4:8])
+colSums(peakset.er[4:8])
+
+normalised.ctcf<-cbind(peakset.ctcf[1:3],
+                       peakset.ctcf[4]/normalisation.factor[1],
+                       peakset.ctcf[5]/normalisation.factor[2],
+                       peakset.ctcf[6]/normalisation.factor[3],
+                       peakset.ctcf[7]/normalisation.factor[4],
+                       peakset.ctcf[8]/normalisation.factor[5])
+
+colSums(normalised.ctcf[4:8])
+
+rpm<-as.numeric(pdx$class["Reads",])/10^6
+
+addToDf<-function(data,factor,method,locus) {
+        data.frame(counts=as.numeric(t(data)),factor=rep(factor,4),method=rep(method,4),rep=c("PDX01","PDX03","PDX04","PDX05"),locus=rep(locus,4))
+}
+
+library(ggpubr)
+
+relativeRPM<-(rpm[c(1,3:5)]/median(rpm[c(1,3:5)]))
+
+df<-addToDf(peakset.ctcf[13416,c(4,6:8)],"CTCF","Raw","RARa")
+df<-rbind(df,addToDf(peakset.er[639,c(4,6:8)],"ER","Raw","RARa"))
+df<-rbind(df,addToDf(peakset.ctcf[13416,c(4,6:8)]/relativeRPM,"CTCF","RPM","RARa"))
+df<-rbind(df,addToDf(peakset.er[639,c(4,6:8)]/relativeRPM,"ER","RPM","RARa"))
+df<-rbind(df,addToDf(normalised.ctcf[13416,c(4,6:8)],"CTCF","pfChIP","RARa"))
+df<-rbind(df,addToDf(normalised.er[639,c(4,6:8)],"ER","pfChIP","RARa"))
+
+ctcfSite<-"21382"
+erSite<-"891"
+name<-"GREB1"
+df<-rbind(df,addToDf(peakset.ctcf[ctcfSite,c(4,6:8)],"CTCF","Raw",name))
+df<-rbind(df,addToDf(peakset.er[erSite,c(4,6:8)],"ER","Raw",name))
+df<-rbind(df,addToDf(normalised.ctcf[ctcfSite,c(4,6:8)],"CTCF","pfChIP",name))
+df<-rbind(df,addToDf(normalised.er[erSite,c(4,6:8)],"ER","pfChIP",name))
+df<-rbind(df,addToDf(peakset.ctcf[ctcfSite,c(4,6:8)]/relativeRPM,"CTCF","RPM",name))
+df<-rbind(df,addToDf(peakset.er[erSite,c(4,6:8)]/relativeRPM,"ER","RPM",name))
+
+ctcfSite<-"20453"
+erSite<-"1064"
+name<-"CLIC6"
+df<-rbind(df,addToDf(peakset.ctcf[ctcfSite,c(4,6:8)],"CTCF","Raw",name))
+df<-rbind(df,addToDf(peakset.er[erSite,c(4,6:8)],"ER","Raw",name))
+df<-rbind(df,addToDf(normalised.ctcf[ctcfSite,c(4,6:8)],"CTCF","pfChIP",name))
+df<-rbind(df,addToDf(normalised.er[erSite,c(4,6:8)],"ER","pfChIP",name))
+df<-rbind(df,addToDf(peakset.ctcf[ctcfSite,c(4,6:8)]/relativeRPM,"CTCF","RPM",name))
+df<-rbind(df,addToDf(peakset.er[erSite,c(4,6:8)]/relativeRPM,"ER","RPM",name))
+
+ggbarplot(df, x="rep", y = "counts",
+          fill = "method",    
+          color = "white",            # Set bar border colors to white
+          palette =c("#FC4E07", "#E7B800", "#00AFBB"),
+          sort.by.groups = FALSE,     # Don't sort inside each group
+          x.text.angle = 90,          # Rotate vertically x axis texts
+          ylab = "Corrected Counts",
+          xlab = "PDX",
+          legend.title = "Comparison",
+          facet.by=c("locus","factor"),
+          position=position_dodge(width=0.7)
+)
+
+ggboxplot(df, x="method", y = "counts",
+          palette =c("#FC4E07", "#E7B800", "#00AFBB"),
+          ylab = "Corrected Counts",
+          xlab = "PDX",
+          color="method",
+          facet.by=c("locus","factor"),
+          legend.title = "Comparison"
+)
+
+```
+
+```{r, results='hide', message=FALSE, warning=FALSE}
+#Homer profile
+
+ctcf<-read.table("SLX-15439/blacklist_filtered/Homer_CTCF_Profile.txt", sep="\t", header=TRUE)
+er<-read.table("SLX-15439/blacklist_filtered/Homer_ER_Profile.txt", sep="\t", header=TRUE)
+
+par(mfrow=c(2,2))
+profile_CTCF<-as.matrix(ctcf[,c(1,2,5,8,11)])
+plot(profile_CTCF,t="l",lwd=3,col="#00AFBB", xlab="Distance from Summit",ylab="Read Depth",main="CTCF Profile")
+lines(profile_CTCF[,c(1,3)],lwd=3,col="#E7B800")
+lines(profile_CTCF[,c(1,4)],lwd=3,col="#FC4E07")
+lines(profile_CTCF[,c(1,5)],lwd=3,col="#FC00FE")
+legend("topright",legend = c("PDX01", "PDX03", "PDX04", "PDX05"), col=c("#00AFBB","#E7B800", "#FC4E07","#FC00FE"),pch = c(19,19))
+
+maxCTCF<-colMaxs(profile_CTCF)
+minCTCF<-colMins(profile_CTCF)
+
+profile_CTCF[,2:5]<-profile_CTCF[,2:5]-minCTCF[2:5]
+
+colSums(profile_CTCF[,2:5]/maxCTCF[2:5])
+normalised_CTCF<-t((t(profile_CTCF[,2:5])/maxCTCF[2:5]))
+normalised_CTCF<-cbind(profile_CTCF[,1],normalised_CTCF)
+
+plot(normalised_CTCF,lwd=3,t="l",col="#00AFBB", xlab="Distance from Summit",ylab="Read Depth",main="Normalised CTCF Profile")
+lines(normalised_CTCF[,c(1,3)],lwd=3,col="#E7B800")
+lines(normalised_CTCF[,c(1,4)],lwd=3,col="#FC4E07")
+lines(normalised_CTCF[,c(1,5)],lwd=3,col="#FC00FE")
+legend("topright",legend = c("PDX01", "PDX03", "PDX04", "PDX05"), col=c("#00AFBB","#E7B800", "#FC4E07","#FC00FE"),pch = c(19,19))
+
+profile_ER<-as.matrix(er[,c(1,2,5,8,11)])
+plot(profile_ER,t="l",lwd=3,col="#00AFBB", xlab="Distance from Summit",ylab="Read Depth",main="ER Profile")
+lines(profile_ER[,c(1,3)],lwd=3,col="#E7B800")
+lines(profile_ER[,c(1,4)],lwd=3,col="#FC4E07")
+lines(profile_ER[,c(1,5)],lwd=3,col="#FC00FE")
+legend("topright",legend = c("PDX01", "PDX03", "PDX04", "PDX05"), col=c("#00AFBB","#E7B800", "#FC4E07","#FC00FE"),pch = c(19,19))
+
+normalised_ER<-t((t(profile_ER[,2:5])/maxCTCF[2:5]))
+normalised_ER<-cbind(profile_ER[,1],normalised_ER)
+
+plot(normalised_ER,t="l",lwd=3,col="#00AFBB", xlab="Distance from Summit",ylab="Read Depth",main="Normalised ER Profile")
+lines(normalised_ER[,c(1,3)],lwd=3,col="#E7B800")
+lines(normalised_ER[,c(1,4)],lwd=3,col="#FC4E07")
+lines(normalised_ER[,c(1,5)],lwd=3,col="#FC00FE")
+legend("topright",legend = c("PDX01", "PDX03", "PDX04", "PDX05"), col=c("#00AFBB","#E7B800", "#FC4E07","#FC00FE"),pch = c(19,19))
+
+```
